### 2025-01-22 style/proceedings.tex
initial commit

\documentclass{sigchi}

% Arabic page numbers for submission.  Remove this line to eliminate
% page numbers for the camera ready copy
% \pagenumbering{arabic}

% Load basic packages
\usepackage{balance}       % to better equalize the last page
\usepackage{graphics}      % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
\usepackage{txfonts}
\usepackage{mathptmx}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{color}
\usepackage{booktabs}
\usepackage{textcomp}

% Some optional stuff you might like/need.
\usepackage{microtype}        % Improved Tracking and Kerning
% \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
\usepackage{ccicons}          % Cite your images correctly!
% \usepackage[utf8]{inputenc} % for a UTF8 editor only

% If you want to use todo notes, marginpars etc. during creation of
% your draft document, you have to enable the "chi_draft" option for
% the document class. To do this, change the very first line to:
% "\documentclass[chi_draft]{sigchi}". You can then place todo notes
% by using the "\todo{...}"  command. Make sure to disable the draft
% option again before submitting your final document.
\usepackage{todonotes}

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{SIGCHI Conference Proceedings Format}
\def\plainauthor{First Author, Second Author, Third Author,
  Fourth Author, Fifth Author, Sixth Author}
\def\emptyauthor{}
\def\plainkeywords{Authors' choice; of terms; separated; by
  semicolons; include commas, within terms only; this section is required.}
\def\plaingeneralterms{Documentation, Standardization}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
% Use \plainauthor for final version.
%  pdfauthor={\plainauthor},
  pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  pdfdisplaydoctitle=true, % For Accessibility
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

% create a shortcut to typeset table headings
% \newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{\plaintitle}

\numberofauthors{2}
\author{%
  \alignauthor{First Author\\
    \email{e-mail address}}\\
  \alignauthor{Second Author\\
    \email{e-mail address}}\\
}

\maketitle

\section{Introduction}

Författarnamnen kommer så småningom att flyttas över till ett särskilt försättsblad inför slutlig publicering i DiVA. Det kan dock vara bra att ha med författarnamnen redan under tiden som rapporten utarbetas. Det kommer även att behövs skrivas ett abstract, som även den kommer att läggas på ett försättsblad. Själva den 10-sidiga exjobbsrapporten ska dock formateras enligt denna template.

Som ni ser så får man jobba lite med att placera sina figurer, eftersom \LaTeX enligt här mallen vill lägga figurerna längst upp eller längst ner på en sida.

We ask that authors follow some
simple guidelines. You should format your paper exactly like this
document. The easiest way to do this is to replace the content with
your own material.  This document describes how to prepare your
submissions using \LaTeX.

\section{Page Size and Columns}
On each page your material should fit within a rectangle of 7 $\times$
9.15 inches (18 $\times$ 23.2 cm), centered on a US Letter page (8.5
$\times$ 11 inches), beginning 0.85 inches (1.9 cm) from the top of
the page, with a 0.3 inches (0.85 cm) space between two 3.35 inches
(8.4 cm) columns. Right margins should be justified, not
ragged. Please be sure your document and PDF are US letter and not A4.

\section{Typeset Text}
The styles contained in this document have been modified from the
default styles to reflect ACM formatting conventions. For example,
content paragraphs like this one are formatted using the Normal style.

\LaTeX\ sometimes will create overfull lines that extend into columns.
To attempt to combat this, the \texttt{.cls} file has a command,
\texttt{{\textbackslash}sloppy}, that essentially asks \LaTeX\ to
prefer underfull lines with extra whitespace.  For more details on
this, and info on how to control it more finely, check out
{\url{http://www.economics.utoronto.ca/osborne/latex/PMAKEUP.HTM}}.

\subsection{Title and Authors}

Your paper's title, authors and affiliations should run across the
full width of the page in a single column 17.8 cm (7 in.) wide.  The
title should be in Helvetica or Arial 18-point bold.  Authors' names
should be in Times New Roman or Times Roman 12-point bold, and
affiliations in 12-point regular.

See \texttt{{\textbackslash}author} section of this template for
instructions on how to format the authors. For more than three
authors, you may have to place some address information in a footnote,
or in a named section at the end of your paper. Names may optionally
be placed in a single centered row instead of at the top of each
column. Leave one 10-point line of white space below the last line of
affiliations.

\subsection{Normal or Body Text}

Please use a 10-point Times New Roman or Times Roman font or, if this
is unavailable, another proportional font with serifs, as close as
possible in appearance to Times Roman 10-point. Other than Helvetica
or Arial headings, please use sans-serif or non-proportional fonts
only for special purposes, such as source code text.

\subsection{First Page Copyright Notice}
This template include a sample ACM copyright notice at the bottom of
page 1, column 1.  Upon acceptance, you will be provided with the
appropriate copyright statement and unique DOI string for publication.
Accepted papers will be distributed in the conference
publications. They will also be placed in the ACM Digital Library,
where they will remain accessible to thousands of researchers and
practitioners worldwide. See
\url{http://acm.org/publications/policies/copyright_policy} for the
ACM's copyright and permissions policy.

\subsection{Subsequent Pages}

On pages beyond the first, start at the top of the page and continue
in double-column format.  The two columns on the last page should be
of equal length.

\begin{figure}
\centering
  \includegraphics[width=0.9\columnwidth]{figures/sigchi-logo}
  \caption{Insert a caption below each figure. Do not alter the
    Caption style.  One-line captions should be centered; multi-line
    should be justified. }~\label{fig:figure1}
\end{figure}

\subsection{References and Citations}

Use a numbered list of references at the end of the article, ordered
alphabetically by last name of first author, and referenced by numbers
in
brackets~\cite{acm_categories,ethics,Klemmer:2002:WSC:503376.503378}.
Your references should be published materials accessible to the
public. Internal technical reports may be cited only if they are
easily accessible (i.e., you provide the address for obtaining the
report within your citation) and may be obtained by any reader for a
nominal fee. Proprietary information may not be cited. Private
communications should be acknowledged in the main text, not referenced
(e.g., ``[Borriello, personal communication]'').

References should be in ACM citation format:
\url{http://acm.org/publications/submissions/latex_style}. This
includes citations to internet
resources~\cite{acm_categories,cavender:writing,CHINOSAUR:venue,psy:gangnam}
according to ACM format, although it is often appropriate to include
URLs directly in the text, as above.

% Use a numbered list of references at the end of the article, ordered
% alphabetically by first author, and referenced by numbers in
% brackets~\cite{ethics, Klemmer:2002:WSC:503376.503378,
%   Mather:2000:MUT, Zellweger:2001:FAO:504216.504224}. For papers from
% conference proceedings, include the title of the paper and an
% abbreviated name of the conference (e.g., for Interact 2003
% proceedings, use \textit{Proc. Interact 2003}). Do not include the
% location of the conference or the exact date; do include the page
% numbers if available. See the examples of citations at the end of this
% document. Within this template file, use the \texttt{References} style
% for the text of your citation.

% Your references should be published materials accessible to the
% public.  Internal technical reports may be cited only if they are
% easily accessible (i.e., you provide the address for obtaining the
% report within your citation) and may be obtained by any reader for a
% nominal fee.  Proprietary information may not be cited. Private
% communications should be acknowledged in the main text, not referenced
% (e.g., ``[Robertson, personal communication]'').

\begin{table}
  \centering
  \begin{tabular}{l r r r}
    % \toprule
    & & \multicolumn{2}{c}{\small{\textbf{Test Conditions}}} \\
    \cmidrule(r){3-4}
    {\small\textit{Name}}
    & {\small \textit{First}}
      & {\small \textit{Second}}
    & {\small \textit{Final}} \\
    \midrule
    Marsden & 223.0 & 44 & 432,321 \\
    Nass & 22.2 & 16 & 234,333 \\
    Borriello & 22.9 & 11 & 93,123 \\
    Karat & 34.9 & 2200 & 103,322 \\
    % \bottomrule
  \end{tabular}
  \caption{Table captions should be placed below the table. We
    recommend table lines be 1 point, 25\% black. Minimize use of
    table grid lines.}~\label{tab:table1}
\end{table}

\section{Sections}

The heading of a section should be in Helvetica or Arial 9-point bold,
all in capitals. Sections should \textit{not} be numbered.

\subsection{Subsections}

Headings of subsections should be in Helvetica or Arial 9-point bold
with initial letters capitalized.  For sub-sections and
sub-subsections, a word like \emph{the} or \emph{of} is not
capitalized unless it is the first word of the heading.

\subsubsection{Sub-subsections}

Headings for sub-subsections should be in Helvetica or Arial 9-point
italic with initial letters capitalized.  Standard
\texttt{{\textbackslash}section}, \texttt{{\textbackslash}subsection},
and \texttt{{\textbackslash}subsubsection} commands will work fine in
this template.

\section{Figures/Captions}

Place figures and tables at the top or bottom of the appropriate
column or columns, on the same page as the relevant text (see
Figure~\ref{fig:figure1}). A figure or table may extend across both
columns to a maximum width of 17.78 cm (7 in.).

\begin{figure*}
  \centering
  \includegraphics[width=1.75\columnwidth]{figures/map}
  \caption{In this image, the map maximizes use of space. You can make
    figures as wide as you need, up to a maximum of the full width of
    both columns. Note that \LaTeX\ tends to render large figures on a
    dedicated page. Image: \ccbynd~ayman on
    Flickr.}~\label{fig:figure2}
\end{figure*}

Captions should be Times New Roman or Times Roman 9-point bold.  They
should be numbered (e.g., ``Table~\ref{tab:table1}'' or
``Figure~\ref{fig:figure1}''), centered (if one line) otherwise justified, and placed beneath the figure
or table.  Please note that the words ``Figure'' and ``Table'' should
be spelled out (e.g., ``Figure'' rather than ``Fig.'') wherever they
occur. Figures, like Figure~\ref{fig:figure2}, may span columns and
all figures should also include alt text for improved accessibility.
Papers and notes may use color figures, which are included in the page
limit; the figures must be usable when printed in black-and-white in
the proceedings.

The paper may be accompanied by a short video figure (we recommend staying within five
minutes in length). However, the paper should stand on its own without
the video figure, as the video may not be available to everyone who
reads the paper.

\subsection{Inserting Images}
When possible, include a vector formatted graphic (i.e. PDF or EPS).
When including bitmaps,  use an image editing tool to resize the image
at the appropriate printing resolution (usually 300 dpi).

\section{Quotations}
Quotations may be italicized when \textit{``placed inline''}.

\begin{quote}
Longer quotes, when placed in their own paragraph, need not be
italicized or in quotation marks when indented.  
\end{quote}

\section{Language, Style, and Content}

The written and spoken language of SIGCHI is English. Spelling and
punctuation may use any dialect of English (e.g., British, Canadian,
US, etc.) provided this is done consis- tently. Hyphenation is
optional. To ensure suitability for an international audience, please
pay attention to the following:

\begin{itemize}
\item Write in a straightforward style.
\item Try to avoid long or complex sentence structures.
\item Use common and basic vocabulary (e.g., use the word ``unusual'' rather than the word ``arcane''.
\item Briefly define or explain all technical terms that may be
  unfamiliar to readers.
\item Explain all acronyms the first time they are used in your
  text---e.g., ``Digital Signal Processing (DSP)''.
\item Explain local references (e.g., not everyone knows all city
  names in a particular country).
\item Explain ``insider'' comments. Ensure that your whole audience
  understands any reference whose meaning you do not describe (e.g.,
  do not assume that everyone has used a Macintosh or a particular
  application).
\item Explain colloquial language and puns. Understanding phrases like
  ``red herring'' may require a local knowledge of English.  Humor and
  irony are difficult to translate.
\item Use unambiguous forms for culturally localized concepts, such as
  times, dates, currencies, and numbers (e.g., ``1--5--97'' or
  ``5/1/97'' may mean 5 January or 1 May, and ``seven o'clock'' may
  mean 7:00 am or 19:00). For currencies, indicate equivalences:
  ``Participants were paid {\fontfamily{txr}\selectfont \textwon}
  25,000, or roughly US \$22.''
\item Be careful with the use of gender-specific pronouns (he, she)
  and other gendered words (chairman, manpower, man-months). Use
  inclusive language that is gender-neutral (e.g., she or he, they,
  s/he, chair, staff, staff-hours, person-years). See the
  \textit{Guidelines for Bias-Free Writing} for further advice and
  examples regarding gender and other personal
  attributes~\cite{Schwartz:1995:GBF}. Be particularly aware of
  considerations around writing about people with disabilities.
\item If possible, use the full (extended) alphabetic character set
  for names of persons, institutions, and places (e.g.,
  Gr{\o}nb{\ae}k, Lafreni\'ere, S\'anchez, Nguy{\~{\^{e}}}n,
  Universit{\"a}t, Wei{\ss}enbach, Z{\"u}llighoven, \r{A}rhus, etc.).
  These characters are already included in most versions and variants
  of Times, Helvetica, and Arial fonts.
\end{itemize}

\section{Accessibility}
The Executive Council of SIGCHI has committed to making SIGCHI
conferences more inclusive for researchers, practitioners, and
educators with disabilities. As a part of this goal, the all authors
are asked to work on improving the accessibility of their
submissions. Specifically, we encourage authors to carry out the
following five steps:
\begin{enumerate}
\item Add alternative text to all figures
\item Mark table headings
\item Add tags to the PDF
\item Verify the default language
\item Set the tab order to ``Use Document Structure''
\end{enumerate}
For more information and links to instructions and resources, please
see: \url{http://chi2016.acm.org/accessibility}.  The
\texttt{{\textbackslash}hyperref} package allows you to create well tagged PDF files,
please see the preamble of this template for an example.

\section{Page Numbering, Headers and Footers}
Your final submission should not contain footer or header information
at the top or bottom of each page. Specifically, your final submission
should not include page numbers. Initial submissions may include page
numbers, but these must be removed for camera-ready. Page numbers will
be added to the PDF when the proceedings are assembled.

\section{Producing and Testing PDF Files}

We recommend that you produce a PDF version of your submission well
before the final deadline.  Your PDF file must be ACM DL
Compliant. The requirements for an ACM Compliant PDF are available at:
{\url{http://www.scomminc.com/pp/acmsig/ACM-DL-pdfs-requirements.htm}}.

Test your PDF file by viewing or printing it with the same software we
will use when we receive it, Adobe Acrobat Reader Version 10. This is
widely available at no cost. Note that most
reviewers will use a North American/European version of Acrobat
reader, so please check your PDF accordingly.

\section{Conclusion}

It is important that you write for the SIGCHI audience. Please read
previous years' proceedings to understand the writing style and
conventions that successful authors have used. It is particularly
important that you state clearly what you have done, not merely what
you plan to do, and explain how your work is different from previously
published work, i.e., the unique contribution that your work makes to
the field. Please consider what the reader will learn from your
submission, and how they will find your work useful. If you write with
these questions in mind, your work is more likely to be successful,
both in being accepted into the conference, and in influencing the
work of our field.

\section{Acknowledgments}

Sample text: We thank all the volunteers, and all publications support
and staff, who wrote and provided helpful comments on previous
versions of this document. Authors 1, 2, and 3 gratefully acknowledge
the grant from NSF (\#1234--2012--ABC). \textit{This whole paragraph is
  just an example.}

% Balancing columns in a ref list is a bit of a pain because you
% either use a hack like flushend or balance, or manually insert
% a column break.  http://www.tex.ac.uk/cgi-bin/texfaq2html?label=balance
% multicols doesn't work because we're already in two-column mode,
% and flushend isn't awesome, so I choose balance.  See this
% for more info: http://cs.brown.edu/system/software/latex/doc/balance.pdf
%
% Note that in a perfect world balance wants to be in the first
% column of the last page.
%
% If balance doesn't work for you, you can remove that and
% hard-code a column break into the bbl file right before you
% submit:
%
% http://stackoverflow.com/questions/2149854/how-to-manually-equalize-columns-
% in-an-ieee-paper-if-using-bibtex
%
% Or, just remove \balance and give up on balancing the last page.
%
\balance{}

\section{References Format}
Your references should be published materials accessible to the
public. Internal technical reports may be cited only if they are
easily accessible and may be obtained by any reader for a nominal
fee. Proprietary information may not be cited. Private communications
should be acknowledged in the main text, not referenced (e.g.,
[Golovchinsky, personal communication]). References must be the same
font size as other body text. References should be in alphabetical
order by last name of first author. Use a numbered list of references
at the end of the article, ordered alphabetically by last name of
first author, and referenced by numbers in brackets. For papers from
conference proceedings, include the title of the paper and the name of
the conference. Do not include the location of the conference or the
exact date; do include the page numbers if available.

References should be in ACM citation format:
\url{http://www.acm.org/publications/submissions/latex_style}.  This
includes citations to Internet
resources~\cite{CHINOSAUR:venue,cavender:writing,psy:gangnam}
according to ACM format, although it is often appropriate to include
URLs directly in the text, as above. Example reference formatting for
individual journal articles~\cite{ethics}, articles in conference
proceedings~\cite{Klemmer:2002:WSC:503376.503378},
books~\cite{Schwartz:1995:GBF}, theses~\cite{sutherland:sketchpad},
book chapters~\cite{winner:politics}, an entire journal
issue~\cite{kaye:puc},
websites~\cite{acm_categories,cavender:writing},
tweets~\cite{CHINOSAUR:venue}, patents~\cite{heilig:sensorama}, 
games~\cite{supermetroid:snes}, and
online videos~\cite{psy:gangnam} is given here.  See the examples of
citations at the end of this document and in the accompanying
\texttt{BibTeX} document. This formatting is a edited version of the
format automatically generated by the ACM Digital Library
(\url{http://dl.acm.org}) as ``ACM Ref.'' DOI and/or URL links are
optional but encouraged as are full first names. Note that the
Hyperlink style used throughout this document uses blue links;
however, URLs in the references section may optionally appear in
black.

% BALANCE COLUMNS
\balance{}

% REFERENCES FORMAT
% References must be the same font size as other body text.
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{sample}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2025-02-19 docs/style/proceedings.tex
revised template compilation

\documentclass{sigchi}

% Arabic page numbers for submission.  Remove this line to eliminate
% page numbers for the camera ready copy
% \pagenumbering{arabic}

% Load basic packages
\usepackage{balance}       % to better equalize the last page
\usepackage{graphics}      % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
\usepackage{txfonts}
\usepackage{mathptmx}
\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{color}
\usepackage{booktabs}
\usepackage{textcomp}

% Some optional stuff you might like/need.
\usepackage{microtype}        % Improved Tracking and Kerning
% \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
\usepackage{ccicons}          % Cite your images correctly!
% \usepackage[utf8]{inputenc} % for a UTF8 editor only

% If you want to use todo notes, marginpars etc. during creation of
% your draft document, you have to enable the "chi_draft" option for
% the document class. To do this, change the very first line to:
% "\documentclass[chi_draft]{sigchi}". You can then place todo notes
% by using the "\todo{...}"  command. Make sure to disable the draft
% option again before submitting your final document.
\usepackage{todonotes}

% Paper metadata (use plain text, for PDF inclusion and later
% re-using, if desired).  Use \emtpyauthor when submitting for review
% so you remain anonymous.
\def\plaintitle{SIGCHI Conference Proceedings Format}
\def\plainauthor{First Author, Second Author, Third Author,
  Fourth Author, Fifth Author, Sixth Author}
\def\emptyauthor{}
\def\plainkeywords{Authors' choice; of terms; separated; by
  semicolons; include commas, within terms only; this section is required.}
\def\plaingeneralterms{Documentation, Standardization}

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{
    \def\UrlFont{\sf}
  }{
    \def\UrlFont{\small\bf\ttfamily}
  }}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, to give it a
% fighting chance of not being over-written, since its job is to
% redefine many LaTeX commands.
\definecolor{linkColor}{RGB}{6,125,233}
\hypersetup{%
  pdftitle={\plaintitle},
% Use \plainauthor for final version.
%  pdfauthor={\plainauthor},
  pdfauthor={\emptyauthor},
  pdfkeywords={\plainkeywords},
  pdfdisplaydoctitle=true, % For Accessibility
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

% create a shortcut to typeset table headings
% \newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{\plaintitle}

\numberofauthors{2}
\author{%
	\alignauthor{Love Arreborn\\
		\email{lovar063@student.liu.se}}\\
	\alignauthor{Nadim Lakrouz\\
		\email{nadla777@student.liu.se}}\\
}

\maketitle

\section{1 INTRODUCTION}\label{introduction}

In cryptography, there are many applications for randomly generated
numbers. However, the process of producing these random numbers tends to
be pseudo-random, e.g.~utilizing the current states of various modules
{[}@randomness{]}. These numbers do not generate true randomness, and in
order to heighten security other methods of generating these are
required. Current random number generators (\emph{RNG}) are usually
implemented in code, using certain states of the host machine as a
starting point before running a predetermined algorithm
{[}@randomness{]}. This pseudo-random number generation (\emph{PRNG})
comes with the drawback that the result is always deterministic,
provided that the initial state is known.

True random numbers, then, cannot be produced solely through code. These
systems require some input that is neither replicable nor reproducible.
One proposed solution for this is quantum random number generation
(\emph{QRNG}) {[}@QRNG{]}. By reading quantum fluctuations from any
given source, for instance an optical signal, the inherent natural
unpredictability of said source can be harnessed in order to produce a
random number from a state that is nigh impossible to reproduce
accurately.

In this project, we will be writing firmware for one such solution,
which reads quantum variations from an optical signal. Further details
about how this signal is produced will be introduced in section 2 and
builds on the work of Clason {[}@Clason2023{]}. This optical signal will
be converted to a stream of random, raw bits via an Analog to Digital
Converter (\emph{ADC}). In turn, these random bits will be processed via
Toeplitz-hashing {[}@toeplitz{]} in order to process these bits into
random numbers. Some processing has to be done on the microcontrollers
themselves in order to ensure that the data is workable, and
Toeplitz-hashing is a tried and tested method to accomplish this. These
random numbers will then be output from the microcontroller to the host
computer via USB. This thesis will aim to answer one key research
question: How can sampled vacuum fluctuations be processed efficiently
in order to output QRNG?

In producing this firmware, several key considerations have to be made
in order for this system to be usable in a production environment. The
vision for the end product is a simple USB-stick that can be connected
to a host, and produce true random numbers from ambient quantum
fluctuations. While this system could very well be implemented on
physically larger hardware, thus avoiding the constraints that limited
hardware introduces, Clason proposes a simpler and cheaper way to
achieve QRNG {[}@Clason2023{]}. In keeping with this, utilizing
microcontrollers for processing the raw bits extracted further helps to
keep the costs low and the solution reasonably complex. Further details
regarding the microcontrollers used in our work will be outlined in
section 3. However, due to this portability constraint, our
implementation needs to work quickly and efficiently on resource
constrained hardware. As such, our main question is broken down into two
concrete research areas:

\textbf{Research area 1 (\emph{RA1})}: How can Toeplitz-hashing be
implemented as effectively as possible on resource constrained hardware
in order to process raw bits into a workable random number?

Toeplitz-hashing been optimized quite well, and previous research can be
utilized for this. However, there are still considerations when
implementing the firmware for the microcontroller in order to optimize
the code. Our goal is to attempt several implementations in order to
find the most optimal implementation with the least amount of
CPU-cycles.

\textbf{Research area 2 (\emph{RA2})}: How can we ensure that the output
of random numbers is not limited by our firmware, but rather only
limited by the USB transfer speed of the microcontroller or the ADC?

There will unequivocally be a bottleneck for the processing speed. For
instance, the speed at which the ADC can process the optical signal into
raw bits as well as the speed that the USB output can transfer processed
random number to the host computer will be limiting factors. Further
details on the limitations of the ADC will be outlined in section 3. The
slowest of these bottlenecks will inevitably be the limiting factor for
any implementation. Our research aims to ensure that our implementation
of Toeplitz-hashing does not become the limiting factor, but rather
processing data fast enough to match or exceed the speed of the
hardware.

\subsection{2 RELATED WORKS}\label{related-works}

This work is a practical continuation of the work of Clason
{[}@Clason2023{]}. In this work, the author aimed to study quantum shot
noise originating from photodiodes, and in so doing built a device which
read from an optical source, outputting analog voltage from the data
``seen'' by the diodes. A prototype was constructed, in which an LED is
read by a photodiode soldered millimeters apart.

The idea of an optical QRNG (\emph{OQRNG}) is not a novel one. The basis
of the theory is that intrinsically random properties of a quantum
process. Stefanov et. al.~proposes using the random choice of a photon
between two output signals to generate a random stream of bits, however
the theory behind it can be applied to other quantum processes as well
{[}@StefanovOptical{]}. This particular theorem has been implemented by
Wayne et. al.~to create a quantum number generator {[}@Wayne{]}. While
this article proves the efficacy of OQRNG, it utilizes a slightly
different method.

Our work revolves around the measurement of shot noise of vacuum states
rather than measuring arrival times of photons. Essentially, this is
another quantum process with the same inherently random properties as
described by Stefanov et. al., but instead using shot noise. As
described by Niemczuk {[}@shotnoise{]}, shot noise is minor fluctuations
in an elecritcal current, which is inherently random. Reading this
property, then, gives us an intrisically random source from which to
generate a random output, which in turn can be processed into a random
number.

Implementations of this theory exist, however with significant
drawbacks. Shen et. al.~presents an implementation using a fairly
complex setup, in which a continuous-wave fiber laser is the optical
source {[}@contender1{]}. They conclude that sampling the shot noise is,
indeed, suitable for OQRNG. However, the implementation requires
expensive and complex hardware, and the sheer size of the system
prohibits it from being portable and easily reproducible in small-scale
tests.

A more recent implementation of OQRNG in a smaller scale has been
presented by Singh et. al. {[}@singh{]}. This particular implementation
uses a bespoke circuit board where all components are present on a
single board -- e.g., this experimental setup contains an integrated
ADC, post-processor, entropy controller and entropy generator. While
this article cements the viability of OQRNG using shot noise (despite
the article not being confirmed as peer reviewed), the bespoke nature of
the circuit board makes this experiment difficult to reproduce. As our
thesis will use commercially available ADCs and microcontrollers, the
only bespoke component is the shot noise generator itself. Furthermore,
the Toeplitz-hashing is not run on the microcontroller itself in these
experiments -- instead, the hashing of these raw bits is done on the
receiving computer as this bespoke circuit board featured a relatively
weak processor.

In summary, previous research has proven that OQRNG can generate true
randomness, and more specifically, Shen et. al. {[}@contender1{]} and
Singh et. al. {[}@singh{]} both implement OQRNG through readings of shot
noise. However, there are limitations in both of these works. Either the
system that generates the shot noise is large and complex
{[}@contender1{]} or the system is built on bespoke hardware with
limitations in processing power which prevents a fully integrated system
{[}@singh{]}. Our work aims to bridge this gap by using commercially
available hardware (other than the bespoke shot noise generator
{[}@Clason2023{]}) and focuses on implementing Toeplitz-hashing directly
on the microcontroller.

\pagebreak
% BALANCE COLUMNS
\balance{}

% REFERENCES FORMAT
% References must be the same font size as other body text.
\bibliographystyle{SIGCHI-Reference-Format}
\bibliography{sample}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2025-02-19 docs/fix_citations.py

@@ -1,0 +1,14 @@
+import sys
+import re
+
+def fix_citations(file_path):
+    with open(file_path, "r", encoding="utf-8") as f:
+        content = f.read()
+
+    content = re.sub(r'\{\[}@([^{}]+?)\{\]\}', r'\\cite{\1}', content)
+
+    with open(file_path, "w", encoding="utf-8") as f:
+        f.write(content)
+
+if __name__ == "__main__":
+    fix_citations(sys.argv[1])

### 2025-03-10 docs/style/article-body.tex
some more intro and small details, one more reference added

\section{1 INTRODUCTION}\label{introduction}

In computer science, there are many applications for randomly generated numbers. From generating keys for cryptography

However, the process of producing these random numbers tends to be pseudo-random, e.g.~utilizing the current states of various modules \cite{randomness}. These numbers do not generate true randomness, and in order to heighten security other methods of generating these are required. Current random number generators (\emph{RNG}) are usually implemented in code, using certain states of the host machine as a starting point before running a predetermined algorithm \cite{randomness}. This pseudo-random number generation (\emph{PRNG}) comes with the drawback that the result is always deterministic, provided that the initial state is known.

Imagine, then, if a malicious attacker somehow manages to ascertain the state a computer was in when it generated a random number, for instance to produce an SSH-key. This attacker then has the opportunity to accurately reproduce the exact, deterministic state that produced said random number, in essence removing the safety that randomness brings. While it may sound unrealistic, the exponential increase in processing power and the burgeoning field of quantum computing does introduce the possibility that one day, what we perceive as random is nothing more than a simple algorithm to crack.

True random numbers, then, cannot be produced solely through code. These systems require some input that is neither replicable nor reproducible. One method that can be realistically used is the inherently random movement of lava lamps \cite{lavarand}, which is used as a backup source of randomness for Cloudflare\footnote{\href{https://blog.cloudflare.com/randomness-101-lavarand-in-production/}{Cloudflare.com, accessed 2025-03-10}}. Another proposed solution for this is quantum random number generation (\emph{QRNG}) \cite{QRNG}. By reading quantum fluctuations from any given source, for instance an optical signal, the inherent natural unpredictability of said source can be harnessed in order to produce a random number from a state that is nigh impossible to reproduce accurately.

In this project, we will be writing firmware for one such solution, which reads quantum variations from an optical signal. Further details about how this signal is produced will be introduced in section 2 and builds on the work of Clason \cite{Clason2023}. This optical signal will be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}). In turn, these random bits will be processed via Toeplitz extraction \cite{toeplitz} in order to process these bits into random numbers. Some processing has to be done on the microcontrollers themselves in order to ensure that the data is workable, and Toeplitz extraction is a tried and tested method to accomplish this. These random numbers will then be output from the microcontroller to the host computer via USB. This thesis will aim to answer one key research question: How can sampled vacuum fluctuations be processed efficiently in order to output QRNG?

In producing this firmware, several key considerations have to be made in order for this system to be usable in a production environment. The vision for the end product is a simple USB-stick that can be connected to a host, and produce true random numbers from ambient quantum fluctuations. While this system could very well be implemented on physically larger hardware, thus avoiding the constraints that limited hardware introduces, Clason proposes a simpler and cheaper way to achieve QRNG \cite{Clason2023}. In keeping with this, utilizing microcontrollers for processing the raw bits extracted further helps to keep the costs low and the solution reasonably complex. Further details regarding the microcontrollers used in our work will be outlined in section 3. However, due to this portability constraint, our implementation needs to work quickly and efficiently on resource constrained hardware. As such, our main question is broken down into two concrete research areas:

\textbf{Research area 1 (\emph{RA1})}: How can Toeplitz extraction be implemented as effectively as possible on resource constrained hardware in order to process raw bits into a workable random number?

Toeplitz extraction been optimized quite well, and previous research can be utilized for this. However, there are still considerations when implementing the firmware for the microcontroller in order to optimize the code. Our goal is to attempt several implementations in order to find the most optimal implementation with the least amount of CPU-cycles.

\textbf{Research area 2 (\emph{RA2})}: How can we ensure that the output of random numbers is not limited by our firmware, but rather only limited by the USB transfer speed of the microcontroller or the ADC?

There will unequivocally be a bottleneck for the processing speed. For instance, the speed at which the ADC can process the optical signal into raw bits as well as the speed that the USB output can transfer processed random number to the host computer will be limiting factors. Further details on the limitations of the ADC will be outlined in section 3. The slowest of these bottlenecks will inevitably be the limiting factor for any implementation. Our research aims to ensure that our implementation of Toeplitz extraction does not become the limiting factor, but rather processing data fast enough to match or exceed the speed of the hardware.

Section 2 of this article will introduce the theory that allows for QRNG, and how this will be utilized in our works. Section 3 delves further into the hardware and algorithms our work will use, with related works in optimizing Toeplitz extraction listed under section 4. Section 5 will present our methodology as well as some limitations imposed on our work. \textbf{More sections to follow as we finish the article}.

\subsection{2 THEORY}\label{theory}

A majority of the research around this topic stems from physics, with implementations of the technology frequently being published and studied by physicists. As such, a brief introduction to the concepts used in previous research as well as an introduction to previous implementations of this technology will be presented in this section.

The idea of an optical QRNG (\emph{OQRNG}) is not a novel one. The basis of the theory is that intrinsically random properties of a quantum process. Stefanov et. al. \cite{StefanovOptical} proposes using the random choice of a photon between two output signals to generate a random stream of bits, however the theory behind it can be applied to other quantum processes as well. This particular theorem has been implemented by Wayne et. al. \cite{Wayne} to create a quantum number generator. While this article proves the efficacy of OQRNG, it utilizes a slightly different method.

\subsubsection{2.1 Shot noise quantum fluctuations}\label{shot-noise-quantum-fluctuations}

Our work revolves around the measurement of shot noise of vacuum states rather than measuring arrival times of photons. Essentially, this is another quantum process with the same inherently random properties as described by Stefanov et. al. \cite{StefanovOptical}, but instead using shot noise. As described by Niemczuk \cite{shotnoise}, shot noise is minor fluctuations in an electrical current, which is inherently random. Reading this property, then, gives us an intrinsically random source from which to generate a random output, which in turn can be processed into a random number.

Implementations of this theory exist, however with significant drawbacks. Shen et. al. \cite{contender1} presents an implementation using a fairly complex setup, in which a continuous-wave fiber laser is the optical source. They conclude that sampling the shot noise is, indeed, suitable for OQRNG. However, the implementation requires expensive and complex hardware, and the sheer size of the system prohibits it from being portable and easily reproducible in small-scale tests.

A more recent implementation of OQRNG in a smaller scale has been presented by Singh et. al. \cite{singh}. This particular implementation uses a bespoke circuit board where all components are present on a single board -- e.g., this experimental setup contains an integrated ADC, post-processor, entropy controller and entropy generator. While this article cements the viability of OQRNG using shot noise (despite the article not being confirmed as peer reviewed), the bespoke nature of the circuit board makes this experiment difficult to reproduce. As our thesis will use commercially available ADCs and microcontrollers, the only bespoke component is the shot noise generator itself. Furthermore, the Toeplitz extraction is not run on the microcontroller itself in these experiments -- instead, the hashing of these raw bits is done on the receiving computer as this bespoke circuit board featured a relatively weak processor.

In summary, previous research has proven that OQRNG can generate true randomness, and more specifically, Shen et. al. \cite{contender1} and Singh et. al. \cite{singh} both implement OQRNG through readings of shot noise. However, there are limitations in both of these works. Either the system that generates the shot noise is large and complex \cite{contender1} or the system is built on bespoke hardware with limitations in processing power which prevents a fully integrated system \cite{singh}. Furthermore, to the best of our knowledge, most of the work in this field is from the perspective of physicists, and there appears to be little research on this subject in the domain of computer science. Our work aims to bridge this gap by using commercially available hardware (other than the bespoke shot noise generator \cite{Clason2023}) and focuses on implementing Toeplitz extraction directly on the microcontroller. Rather than focusing on the intricacies of quantum fluctuations, we will instead approach this problem from a computer science perspective.

\subsection{3 BACKGROUND}\label{background}

Our work is a practical continuation of the work of Clason \cite{Clason2023}. In this work, quantum shot noise originating from photodiodes was studies, and in so doing a prototype device was constructed. This device read from an optical source, outputting analog voltage from the data ``seen'' by the diodes. A prototype was constructed, in which an LED is read by a photodiode soldered millimeters apart. In this section, we introduce the hardware used for our implementations as well as the considerations taken in order to shift the focus from physics to computer science.

\subsubsection{3.1 Optical RNG module}\label{optical-rng-module}

The one bespoke piece of hardware used in this study is the prototype designed by Clason \cite{Clason2023} as a part of his masters thesis. This device produces the optical shot noise which will be the source of randomness in our work. Moving forward in this article, we will refer to this as the OQRNG-device.

As described in Clason's work \cite{Clason2023}, the OQRNG-device is an electro-optical system which measures optical shot noise, generating quantum randomness. The device has an LED and a photodiode positioned a few millimeters apart, ensuring efficient light coupling. The photodiode detects light from the LED, and converts the light into a current signal, which is sent to a transimpedance amplifier to convert it into measurable voltage. In order to minimize disruptions by other external lights, the system is enclosed in a shielded measurment box.

Whereas the exact quantum mechanisms that ensure that this system ensures randomness and further details regarding the OQRNG-device is better derived directly from Clasons work \cite{Clason2023}, the end result as it correlates to our study is an inherently random, analog voltage current.

\subsubsection{3.2 ADC converter}\label{adc-converter}

This analog current isn't suitable to operate on without further processing. As mentioned in section 1, the signal needs to pass through an ADC and converted into raw bits in order for it to be usable. In his thesis, Clason \cite{Clason2023} suggests a discrete ADC chip capable of analyzing frequencies higher than 25 MHz, as this is the highest frequency studied in his work. The Nyquist-Shannon theorem poses that, in order to accurately reconstruct the signal, the sampling rate must be at least twice the highest frequency component present in the signal, \(f_s \geq 2 f_{\text{max}}\). \textbf{CITATION NEEDED}

\textbf{This section will be filled with more information as soon as the exact ADC we will be using has been selected. The exact requirements needs to be discussed with the project owner before a choice can be made!}

\subsubsection{3.3 Microcontroller}\label{microcontroller}

Microcontrollers (MCUs) are compact and low-power computing devices designed primary for embedded systems and real-time applications. In contrast to general purpose CPUs, MCUs integrates a processor, memory, and peripherals into a single chip, thus enabling efficient and autonomous operation in constrained environments. To program MCUs, various development frameworks and tools such as: Arduino Framework, Espressif IDF, raw C++, etc. Different frameworks can contribute to overhead.

In our project, we will be utilizing several microcontrollers. Primarily, we will be using Teensy 4.1\footnote{\href{https://www.pjrc.com/store/teensy41.html}{Teensy developer documentation, accessed 2025-02-27}} as our baseline, as this controller is relatively cheap yet very powerful.

\subsubsection{\texorpdfstring{3.4 Toeplitz extraction }{3.4 Toeplitz extraction }}\label{toeplitz-extraction}

\textbf{WIP}

\subsubsection{3.5 Summary}\label{summary}

With the assumption that the OQRNG-device produces a truly random analog signal, we can now clearly define the scope in which this thesis operates. Considering the maximum conversion speed from the ADC and the USB-output from the MCU, we have a clear bound between \textbf{PLACEHOLDER} and \textbf{PLACEHOLDER} in which Toeplitz extraction needs to be executed. Any implementation of Toeplitz extraction must then execute fast enough on any given microcontroller feasible for the proposed quantum RNG-thumbstick as to not be the decisive limiting factor.

\subsection{\texorpdfstring{4 RELATED WORKS }{4 RELATED WORKS }}\label{related-works}

\textbf{WIP - Might be good to add related works in the domain of computer science here!}

\subsection{5 METHODOLOGY}\label{methodology}

With the consideration that our work revolves around optimizing Toeplitz extraction in order to quickly process random bits into a random number, we will take an iterative approach. For our tests, we will use a pre-defined stream of raw bits which is loaded into memory on the microcontroller, and run several different implementations of Toeplitz extraction to produce numbers. As we always use a pre-defined bitstream, the result will at this stage be deterministic, giving us a clear indication whether the algorithm works as intended.

However, in order to ensure the results work with varying data, we cannot limit ourselves to simply one stream of bits. The main point of the algorithm is to remove patterns in the bitstream that may lead to less randomized results. As such, we will manually produce several bitstreams to use for our tests -- each with varying degrees of repeated patterns that should be eliminated by the algorithm. Furthermore, in order to see how well our implementation will work with realistic data from the OQRNG generator, we will sample streams of bits directly from the source.

\subsubsection{5.1 Iterative approach}\label{iterative-approach}

Implementation of Toeplitz extraction will begin by a naive implementation not optimized for speed, but rather for accuracy. This implementation will be executed on a separate computer in order to produce the correct random number for every provided bitstring. These will be used as our baseline for accuracy for future iterations.

This naive implementation will then be flashed to our microcontrollers, beginning with Teensy 4.1 as this is the more capable of the microcontrollers used for this experiment. Code to measure the execution speed in milliseconds will be implemented and tested before we load the naive implementation on said microcontroller. We expect that several implementations may be too resource intensive or have a memory complexity far greater than our cheaper, less capable microcontroller are able to handle, and as such these may not be able to be tested until a few iterations of optimization has occurred.

Each iteration will consist of incremental improvements to the algorithm. Our initial investigation has shown several avenues for improving the throughput of the algorithm, and each planned iteration will be discussed briefly. Depending on the empirical results of these iterations, there may be a need for further optimization iterations other than those listed.

\textbf{Iteration 1 - Naive implementation}: The naive implementation consists of a Toeplitz matrix \(T\) and key \(k\), acquired from fixed slices of the bitstream. Then, matrix-vector multiplication will be performed using \(T\) and \(k\), thus performing Toeplitz extraction -- which finally will eliminate jitter and produce a random number. The implementation of this in code will be quite simple, using nested loops to iterate over all values and multiply them. We expect that the throughput of this naive implementation will be far away from optimized iterations, and result in \(O(n^2)\) complexity.

\textbf{Iteration 2 - Efficient data structures}: The naive implementation will not utilize efficient data structures for maximum speed, rather it will likely use fixed-size arrays which need to be iterated over for a time complexity of \(O(n^2)\). Our hypothesis is that more efficient data structures may allow data to be processed in a far more efficient manner -- however, this scenario poses a risk for greater memory complexity which may not be suitable for use on microcontrollers. Which data structures might be most feasible will be evaluated during this iteration and as such is not preordained.

\textbf{Iteration 3 - Bitshifting}: We expect that Toeplitz extraction can be significantly improved with bitwise XOR operations, thus reducing overhead and memory usage. Instead of explicitly constructing the Toeplitz matrix, a right-shift operation can be used to dynamically reconstruct matrix rows. Additionally, matrix-vector multiplication can be improved using XOR operations to extract bits more efficient, thus minimizing unnecessary computations.

\textbf{Iteration 4 - Batching}: Finally, we consider the concept of batching larger amounts of bits for processing. Consider that an input buffer of bits is read from the ADC and stored, waiting for processing. Rather than taking 64 bits, and shifting them one by one, we can take two batches of 64 bits and multiply them directly using XOR bitshifting -- eliminating the need to process these bit-by-bit. As we will have a constant stream of bits from the ADC, we theorize that this method will allow for better performance in real-time processing over reading individual bits.

\textbf{Iteration 5 - ARM Hardware instructions}: Rather than performing bitshifting operations in the code itself, certain microcontrollers come equipped with a separate processor specifically for bitshift-operations. Offloading the shifting to these processors rather than running them on the main CPU may allow faster processing of the data than performing the shifting in the code itself. However, this operation isn't natively supported by Arduino, and will potentially lead to extreme rewrites of the code which may prove too time consuming to do for two microcontrollers (as the code will not be reusable between controllers). As such, this optimization may only be done for one controller or left for future work.

Each scenario will first be executed on a single thread, and multiple threads may theoretically be executed concurrently to further optimize the data. The feasibility of this highly depends on the performance of every individual implementation. Depending on the results during our experimentation, this may yield yet another iteration.

\subsubsection{\texorpdfstring{5.2 Evaluation }{5.2 Evaluation }}\label{evaluation}

Due to the bounds introduced by the hardware limitations, evaluating the throughput of each iteration essentially consists of measuring the execution time. In order to verify accuracy, the output of random numbers should be identical for each sample bitstring tested. Certain iterations might, however, use larger parts of the bitstring, and thus produce a slightly variable result.

As discussed in section 3, the hardware used will impose a clear bound on how quickly our implementation needs to process the bits in order to match the speed of the ADC, as well as the output speed of the USB-port, both in \(MB/s\). Hyncica et. al. \cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements. An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (as the process for this may vary greatly between controllers). As we will use fixed-size bitstrings for evaluation, we can then derive the throughput of the algorithm in \(MB/s\) as follows:

\[
Throughput_{MB/s} = \frac{DataSize_{bits}}{Execution
Time_{ms}} \times \frac{1}{8} \times \frac{1000}{10^6} \text{\phantom{12}(1)}
\]

This measurement allows us to place the throughput of our algorithm soundly in the bounds imposed on us by the hardware. During implementation, measurements of time complexity may be discussed with regards to the code -- but in the end, the performance of the code as it runs is what will be evaluated most thoroughly.

\subsubsection{5.2 Limitations}\label{limitations}

Our proposed iterations all assume that the limited hardware will support it. Whereas we are confident that Teensy 4.1 will be able to handle each iteration step (even the naive implementation), the remaining microcontrollers with lower specifications might not be suitable for the first iterations. Testing the implementations on different microcontrollers could turn out to be unfeasible -- however, this remains to be seen during the experimentation.

\newpage

\phantom{HAHA I CAN CURSE HERE! FUCK YOU!} \newpage

\subsection{CHANGELOG}\label{changelog}

2025-02-14: Added background section, smaller reviews to introduction.

2025-02-28: Template adjusted, added methodology. Started review of background and theory to add stronger correlation to computer science. Not yet finished due to review of articles as well as some additional information required from the project owner. The update to theory and background should be considered a heavy work in progress at this stage.

2025-03-XX: Moved evaluation down in the methodology in order to provide a better flow. Elaborated further on Toeplitz extraction and ADC converters, as well as motivating the selection of these.

### 2025-03-10 docs/style/article-final.tex
some more intro and small details, one more reference added

--- 2025-03-10 docs/style/article-body.tex
@@
+\documentclass{sigchi}
+
+% Arabic page numbers for submission.  Remove this line to eliminate
+% page numbers for the camera ready copy
+% \pagenumbering{arabic}
+
+% Load basic packages
+\usepackage{balance}       % to better equalize the last page
+\usepackage{graphics}      % for EPS, load graphicx instead 
+\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
+\usepackage{txfonts}
+\usepackage{mathptmx}
+\usepackage[pdflang={en-US},pdftex]{hyperref}
+\usepackage{color}
+\usepackage{booktabs}
+\usepackage{textcomp}
+
+% Some optional stuff you might like/need.
+\usepackage{microtype}        % Improved Tracking and Kerning
+% \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
+\usepackage{ccicons}          % Cite your images correctly!
+% \usepackage[utf8]{inputenc} % for a UTF8 editor only
+
+% If you want to use todo notes, marginpars etc. during creation of
+% your draft document, you have to enable the "chi_draft" option for
+% the document class. To do this, change the very first line to:
+% "\documentclass[chi_draft]{sigchi}". You can then place todo notes
+% by using the "\todo{...}"  command. Make sure to disable the draft
+% option again before submitting your final document.
+\usepackage{todonotes}
+
+% Paper metadata (use plain text, for PDF inclusion and later
+% re-using, if desired).  Use \emtpyauthor when submitting for review
+% so you remain anonymous.
+\def\plaintitle{Efficient Exctraction of True Random Numbers from Quantum System on Resource Constrained Hardware}
+\def\plainauthor{First Author, Second Author, Third Author,
+  Fourth Author, Fifth Author, Sixth Author}
+\def\emptyauthor{}
+\def\plainkeywords{Authors' choice; of terms; separated; by
+  semicolons; include commas, within terms only; this section is required.}
+\def\plaingeneralterms{Documentation, Standardization}
+
+% llt: Define a global style for URLs, rather that the default one
+\makeatletter
+\def\url@leostyle{%
+  \@ifundefined{selectfont}{
+    \def\UrlFont{\sf}
+  }{
+    \def\UrlFont{\small\bf\ttfamily}
+  }}
+\makeatother
+\urlstyle{leo}
+
+% To make various LaTeX processors do the right thing with page size.
+\def\pprw{8.5in}
+\def\pprh{11in}
+\special{papersize=\pprw,\pprh}
+\setlength{\paperwidth}{\pprw}
+\setlength{\paperheight}{\pprh}
+\setlength{\pdfpagewidth}{\pprw}
+\setlength{\pdfpageheight}{\pprh}
+
+% Make sure hyperref comes last of your loaded packages, to give it a
+% fighting chance of not being over-written, since its job is to
+% redefine many LaTeX commands.
+\definecolor{linkColor}{RGB}{6,125,233}
+\hypersetup{%
+  pdftitle={\plaintitle},
+% Use \plainauthor for final version.
+%  pdfauthor={\plainauthor},
+  pdfauthor={\emptyauthor},
+  pdfkeywords={\plainkeywords},
+  pdfdisplaydoctitle=true, % For Accessibility
+  bookmarksnumbered,
+  pdfstartview={FitH},
+  colorlinks,
+  citecolor=black,
+  filecolor=black,
+  linkcolor=black,
+  urlcolor=linkColor,
+  breaklinks=true,
+  hypertexnames=false
+}
+
+% create a shortcut to typeset table headings
+% \newcommand\tabhead[1]{\small\textbf{#1}}
+
+% End of preamble. Here it comes the document.
+\begin{document}
+
+\title{\plaintitle}
+
+\numberofauthors{2}
+\author{%
+	\alignauthor{Love Arreborn\\
+		\email{lovar063@student.liu.se}}\\
+	\alignauthor{Nadim Lakrouz\\
+		\email{nadla777@student.liu.se}}\\
+}
+
+\maketitle
+
 \section{1 INTRODUCTION}\label{introduction}
@@
 2025-03-XX: Moved evaluation down in the methodology in order to provide a better flow. Elaborated further on Toeplitz extraction and ADC converters, as well as motivating the selection of these.
+% BALANCE COLUMNS
+\balance{}
+
+% REFERENCES FORMAT
+% References must be the same font size as other body text.
+\bibliographystyle{SIGCHI-Reference-Format}
+\bibliography{sample}
+
+\end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2025-03-13 docs/style/article-body.tex
added one related work

@@
 
-In this project, we will be writing firmware for one such solution, which reads quantum variations from an optical signal.
-Further details about how this signal is produced will be introduced in section 2 and builds on the work of Clason \cite{Clason2023}.
+In this project, we will be writing firmware for a quantum number generator, which generates true randomness by reading shot noise from an optical signal.
+Further details about how this signal is produced will be introduced in section 2 and 3, and builds on the work of Clason \cite{Clason2023}.
 This optical signal will be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}).
@@
 In his thesis, Clason \cite{Clason2023} suggests a discrete ADC chip capable of analyzing frequencies higher than 25 MHz, as this is the highest frequency studied in his work.
-The Nyquist-Shannon theorem poses that, in order to accurately reconstruct the signal, the sampling rate must be at least twice the highest frequency component present in the signal, \(f_s \geq 2 f_{\text{max}}\).
-\textbf{CITATION NEEDED}
+However, in the interest of keeping the implementation light and cheap, we will be using ADCs that provide less samples per second and lower frequencies.
+This is done mainly for ease of development and access to this hardware.
 
-\textbf{This section will be filled with more information as soon as the exact ADC we will be using has been selected.
-The exact requirements needs to be discussed with the project owner before a choice can be made!}
+Initially, we will utilize MAX11102AUB\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} with an effective sample rate of 2 million samples per second (MSPS).
+This ADC provides a 12 bit sample size, providing roughly 2.86 MB/s of sampled data per second, derived by the following calculation.
+
+\[
+Data Rate = \frac{2,000,000_{MSPS} \times 12}{8 \times 1,048,576} \approx 2.86 MB/s \text{\phantom{123} (1)}
+\]
+
+Many microcontrollers furthermore come equipped with internal ADCs that can be utilized, and while these provide a lower sample size (often around 1 MSPS), the ease of development may be prudent to utilize for this proof-of-concept.
+While our initial ADC has a fairly low throughput, this can always be upgraded if it ends up becoming too limiting.
+The final output from the ADC, whether built into the microcontroller or an external one, will be a stream of raw bits, as the analog signal from the OQRNG-device is processed.
 
@@
 Microcontrollers (MCUs) are compact and low-power computing devices designed primary for embedded systems and real-time applications.
-In contrast to general purpose CPUs, MCUs integrates a processor, memory, and peripherals into a single chip, thus enabling efficient and autonomous operation in constrained environments.
-To program MCUs, various development frameworks and tools such as: Arduino Framework, Espressif IDF, raw C++, etc.
-Different frameworks can contribute to overhead.
+Unlike general purpose CPUs, MCUs integrates a processor, memory, and peripherals into a single chip.
+Thus, enabling efficient and autonomous operation in constrained environments.
+Various development frameworks and tools are available for programming MCUs, including the Arduino Framework, manufacturer specific SDKs and raw C++, among others.
+Each framework presents different trade-offs in terms of performance overhead, efficiency and ease of development.
 
-In our project, we will be utilizing several microcontrollers.
-Primarily, we will be using Teensy 4.1\footnote{\href{https://www.pjrc.com/store/teensy41.html}{Teensy developer documentation, accessed 2025-02-27}} as our baseline, as this controller is relatively cheap yet very powerful.
+Modern microcontrollers feature advanced microarchitectural elements to enhance processing capabilities.
+Some of high performance MCUs utilize dual-issue superscalar architectures, enabling them to run several instructions in one cycle.
+This type of design greatly improves single-threaded capabilities, which makes these MCUs ideal for tasks that need real-time data processing.
+The use of dual-issue architectures in embedded devices has demonstrated high execution speeds, along with the added benefit of improved energy efficiency.
+Thus, making them a great option for applications that require low power but high performance.
+Research conducted by Lozano et al.
+\cite{dual-issue} specifically explores energy efficient dual-issue processors designs.
+This study shows that optimizing instruction throughput and power consumption can significantly improve processing efficiency in embedded systems.
+While energy efficiency is not the primary goal for this study, ensuring that the resulting USB-device does not consume too much power is tangentially relevant for real-world usage.
+
+Since MCUs often function under strict timing requirements, it is critical to have effective ways to access memory and transfer data for processing in real time.
+High-performance MCUs enhance memory usage in various ways.
+Some of them use Tightly Coupled Memory (TCM), which gives fast SRAM with specific access routes for important data, avoiding cache misses and guaranteeing consistent performance.
+Moreover, instruction and data caching techniques, including instruction prefetching and branch prediction, help minimize execution delays in computationally intensive real-time applications.
+Another important aspect is Direct Memory Access (DMA), which enables data transfer between peripherals such as the ADC and RAM, without CPU intervention.
+This offloading reduces processing overhead, allowing the MCU to manage fast data transfers effectively.
+These improvements are especially significant for Toeplitz extraction, where large amount of random data needs to be processed and sent quickly with low delay.
+Efficient memory management guarantees that randomness extraction can occur rapidly without major slowdowns in computing.
+Both approaches will be tested during development.
+
+The Teensy 4.1\footnote{\href{https://www.pjrc.com/store/teensy41.html}{Teensy developer documentation, accessed 2025-02-27}}, based on the ARM Cortex-M7 is especially good for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and DSP capabilities.
+The floating-point unit (FPU) and SIMD-style DSP instructions imporove how quickly it can perform bitwise and arithmetic tasks, which are crucial for quick Toeplitz extraction.
+Although, the Cortex-M7 does not have special features for cryptography, its ability to execute multiple instructions quickly, combined with fast memory access and an efficient pipeline design, allows it to effectively manage and process randomness extraction.
+Moreover, its USB High-Speed interface allows rapid transmission of extracted random data to an external system, ensuring minimal bottlenecks in high-rate randomness generation.
 
@@
 
-\textbf{WIP}
+The raw bits from the ADC can potentially have some deterministic patterns, and as such have to be processed somehow in order to remove these patterns.
+Several methods exist for this purpose, and for our work, we will perform this preprocessing via Toeplitz extraction.
+The main focus of this study is to implement this extraction algorithm as effectively as possible on resource constrained hardware.
+
+A detailed account of the inner workings of Toeplitz extraction can be found in the work of Chouhan et al.
+\cite{toeplitz-desc}.
+This work focuses on implementing Toeplitz extraction on field-programmable gate-arrays (FPGA), but some specific details can be derived from their work.
+As these authors describe, Toeplitz extraction is a strong contender for our work due to a lower computational complexity than other alternatives, as well as a relatively easy algorithm to use.
+This extraction utilizes either matrix multiplication or hashing between a pseudo-random seed and the raw data provided from a high-entropy source of randomness -- in our case, the OQRNG-device.
+
+To summarize the theoretical working of Toeplitz extraction (as explained by Chouhan et al.
+\cite{toeplitz-desc}), the sampled raw bit matrix (\(T\)) are multiplied with a pre-determined seed matrix (\(K\)).
+The size of the seed is directly dependant on the size of the sampled data, and can be fixed or continually re-sampled as needed.
+To ensure high levels of entropy, our intuition is that re-sampling the seed from the OQRNG-device continually is prudent.
+The sample and seed with then be processed with matrix multiplication to remove deterministic patterns, and produce a bitstring that results in our randomly generated number.
 
@@
 With the assumption that the OQRNG-device produces a truly random analog signal, we can now clearly define the scope in which this thesis operates.
-Considering the maximum conversion speed from the ADC and the USB-output from the MCU, we have a clear bound between \textbf{PLACEHOLDER} and \textbf{PLACEHOLDER} in which Toeplitz extraction needs to be executed.
+Considering the maximum conversion speed from the ADC and the USB-output from the MCU, we have a clear bound between 2.86 MS/s (for the ADC) and 480 MB/s (for the USB output) in which Toeplitz extraction needs to be executed.
+Any speeds over 2.86 MB/s allows us to upgrade the ADC iteratively to continue increasing the output speed.
 Any implementation of Toeplitz extraction must then execute fast enough on any given microcontroller feasible for the proposed quantum RNG-thumbstick as to not be the decisive limiting factor.
@@
 
-\subsubsection{\texorpdfstring{5.2 Evaluation }{5.2 Evaluation }}\label{evaluation}
+\subsubsection{5.2 Evaluation}\label{evaluation}
 
-Due to the bounds introduced by the hardware limitations, evaluating the throughput of each iteration essentially consists of measuring the execution time.
+As discussed in section 3, the hardware used will impose a clear bound on how quickly our implementation needs to process the bits in order to match the speed of the ADC, as well as the output speed of the USB-port, both in \(MB/s\).
 In order to verify accuracy, the output of random numbers should be identical for each sample bitstring tested.
@@
 
-As discussed in section 3, the hardware used will impose a clear bound on how quickly our implementation needs to process the bits in order to match the speed of the ADC, as well as the output speed of the USB-port, both in \(MB/s\).
 Hyncica et.
 al.
-\cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements.
+\cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements of an algorithms execution speed.
 An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (as the process for this may vary greatly between controllers).
@@
 \[
-Throughput_{MB/s} = \frac{DataSize_{bits}}{Execution
-Time_{ms}} \times \frac{1}{8} \times \frac{1000}{10^6} \text{\phantom{12}(1)}
+Throughput*{MB/s} = \frac{DataSize*{bits}}{Execution Time_{ms}} \times
+\frac{1}{8} \times \frac{1000}{10^6} \text{\phantom{12}(2)}
 \]
@@
 
-\newpage
+\subsection{6 EXPERIMENTATION}\label{experimentation}
 
-\phantom{HAHA I CAN CURSE HERE! FUCK YOU!} \newpage
+Each iteration results in an implementation written in C, all of which will be supplied as an appendix to this paper.
+Furthermore, to simplify the implementation process, we will use the Arduino framework to communicate with the hardware.
+Whereas this introduces some overhead, it drastically reduces the time spent per iteration, and is deemed to be a suitable starting point.
 
@@
 
-2025-03-XX: Moved evaluation down in the methodology in order to provide a better flow. Elaborated further on Toeplitz extraction and ADC converters, as well as motivating the selection of these.
+2025-03-10: Moved evaluation down in the methodology in order to provide a better flow.
+Elaborated further on Toeplitz extraction and ADC converters, as well as motivating the selection of these.
+Some additional information added in introduction as motivation for the work.
+
+2025-03-12: Elaborated on background, as well as adding more details regarding hardware.
+Note that the hardware selected is subject to change over time.
+Further elaborated on related works in optimizing Toeplitz extraction.

### 2025-03-13 docs/style/article-final.tex
added one related work

@@
 
-2025-03-XX: Moved evaluation down in the methodology in order to provide a better flow. Elaborated further on Toeplitz extraction and ADC converters, as well as motivating the selection of these.
+2025-03-10: Moved evaluation down in the methodology in order to provide a better flow.
+Elaborated further on Toeplitz extraction and ADC converters, as well as motivating the selection of these.
+Some additional information added in introduction as motivation for the work.
+
+2025-03-12: Elaborated on background, as well as adding more details regarding hardware.
+Note that the hardware selected is subject to change over time.
+Further elaborated on related works in optimizing Toeplitz extraction.
 % BALANCE COLUMNS

### 2025-03-13 docs/style/article-body.tex
added pico 2 and esp32 information

@@
 High-performance MCUs enhance memory usage in various ways.
-Some of them use Tightly Coupled Memory (TCM), which gives fast SRAM with specific access routes for important data, avoiding cache misses and guaranteeing consistent performance.
+Some of them use Tightly Coupled Memory\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} (TCM), which gives fast SRAM with specific access routes for important data, avoiding cache misses and guaranteeing consistent performance.
 Moreover, instruction and data caching techniques, including instruction prefetching and branch prediction, help minimize execution delays in computationally intensive real-time applications.
-Another important aspect is Direct Memory Access (DMA), which enables data transfer between peripherals such as the ADC and RAM, without CPU intervention.
+Another important aspect is Direct Memory Access\footnote{\href{https://www.sciencedirect.com/topics/computer-science/direct-memory-access}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (DMA), which enables data transfer between peripherals such as the ADC and RAM, without CPU intervention.
 This offloading reduces processing overhead, allowing the MCU to manage fast data transfers effectively.
@@
 The Teensy 4.1\footnote{\href{https://www.pjrc.com/store/teensy41.html}{Teensy developer documentation, accessed 2025-02-27}}, based on the ARM Cortex-M7 is especially good for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and DSP capabilities.
-The floating-point unit (FPU) and SIMD-style DSP instructions imporove how quickly it can perform bitwise and arithmetic tasks, which are crucial for quick Toeplitz extraction.
+The floating-point unit\footnote{\href{https://www.sciencedirect.com/topics/computer-science/floating-point-unit}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (FPU) and SIMD-style DSP instructions improves how quickly it can perform bitwise and arithmetic tasks, which are crucial for quick Toeplitz extraction.
+SIMD-controlled DSP architectures, as described by Han et al.
+\cite{simd-dsp}, leverage parallel vectorized computation to accelerate matrix operations.
+Thus, making them highly effective for Toeplitz matrix-vector multiplications.
 Although, the Cortex-M7 does not have special features for cryptography, its ability to execute multiple instructions quickly, combined with fast memory access and an efficient pipeline design, allows it to effectively manage and process randomness extraction.
-Moreover, its USB High-Speed interface allows rapid transmission of extracted random data to an external system, ensuring minimal bottlenecks in high-rate randomness generation.
+Moreover, its USB High-Speed (480 Mbit/sec) interface, supporting both USB device and USB host modes, allows rapid transmission of extracted random data to an external system, ensuring minimal bottlenecks in high-rate randomness generation.
+
+In order to evaluate just how much constraint our implementation can allow, our aim is to try our implementation on other MCUs with varying levels of power and hardware support.
+Whereas Teensy 4.1 is our primary development platform, we aim to run our implementations on Raspberry Pi Pico 2\footnote{\href{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}{Raspberry Pi Pico 2 documentation, accessed 2025-03-13}} as well as ESP32-S3\footnote{\href{https://www.espressif.com/sites/default/files/documentation/esp32-s3_datasheet_en.pdf}{ESP32-S3 documentation, accessed 2025-03-13}}.
+Due to the lower computational power of these MCUs, there may be significant issues in utilizing these weaker models, yet they are significantly cheaper and easier to access.
 
@@
 
-\textbf{WIP - Might be good to add related works in the domain of computer science here!}
+Most related works found tend to revolve around optimizing Toeplitz using more advanced hardware, and few works seem to delve into evaluating the implementation in code.
+
+As mentioned in section 3.4, Chouhan et al.
+\cite{toeplitz-desc} utilized FPGA to implement Toeplitz extraction specifically for OQRNG, a work also studied by Zhang et al.
+\cite{zhang}.
+Furthermore, Zhang et al.~utilized a standardized min-entropy evaluation to ensure true, unbiased randomness in their result.
+Both of these implementations utilize powerful hardware where the matrix multiplication is offloaded to FPGA.
+These implementations provided extraction speeds of between 3.36 Gpbs \cite{zhang} to 26.57 Gbps \cite{toeplitz-desc}.
+Neither of these implementations operate on constrained hardware, instead creating bespoke circuit boards for their works.
 

### 2025-03-13 docs/style/article-body.tex
stashing...

@@
 In order to evaluate just how much constraint our implementation can allow, our aim is to try our implementation on other MCUs with varying levels of power and hardware support.
-Whereas Teensy 4.1 is our primary development platform, we aim to run our implementations on Raspberry Pi Pico 2\footnote{\href{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}{Raspberry Pi Pico 2 documentation, accessed 2025-03-13}} as well as ESP32-S3\footnote{\href{https://www.espressif.com/sites/default/files/documentation/esp32-s3_datasheet_en.pdf}{ESP32-S3 documentation, accessed 2025-03-13}}.
+Whereas Teensy 4.1 is our primary development platform which we will evaluate closely, we aim to run our implementations on Raspberry Pi Pico 2\footnote{\href{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}{Raspberry Pi Pico 2 documentation, accessed 2025-03-13}} as well as ESP32-S3\footnote{\href{https://www.espressif.com/sites/default/files/documentation/esp32-s3_datasheet_en.pdf}{ESP32-S3 documentation, accessed 2025-03-13}}.
 Due to the lower computational power of these MCUs, there may be significant issues in utilizing these weaker models, yet they are significantly cheaper and easier to access.
+Testing of these will consist solely of running the implementation on these controllers and measuring execution speed.
 
@@
 \cite{toeplitz-desc} utilized FPGA to implement Toeplitz extraction specifically for OQRNG, a work also studied by Zhang et al.
-\cite{zhang}.
+\cite{toeplitz}.
 Furthermore, Zhang et al.~utilized a standardized min-entropy evaluation to ensure true, unbiased randomness in their result.
 Both of these implementations utilize powerful hardware where the matrix multiplication is offloaded to FPGA.
-These implementations provided extraction speeds of between 3.36 Gpbs \cite{zhang} to 26.57 Gbps \cite{toeplitz-desc}.
+These implementations provided extraction speeds of between 3.36 Gpbs \cite{toeplitz} to 26.57 Gbps \cite{toeplitz-desc}.
 Neither of these implementations operate on constrained hardware, instead creating bespoke circuit boards for their works.

### 2025-03-13 docs/style/article-body.tex
added more details of toeplitz

@@
 The sample and seed with then be processed with matrix multiplication to remove deterministic patterns, and produce a bitstring that results in our randomly generated number.
+An example of how this extraction works can be seen below.
+
+\[
+T =
+\begin{bmatrix}
+1 & 0 & 1 & 1 & 0 & 1 & 0 & 1 \\
+0 & 1 & 1 & 0 & 1 & 0 & 1 & 1 \\
+1 & 1 & 0 & 1 & 0 & 1 & 1 & 0 \\
+1 & 0 & 1 & 1 & 1 & 0 & 1 & 0
+\end{bmatrix}
+\quad
+k =
+\begin{bmatrix}
+1 \\ 0 \\ 1 \\ 1 \\ 0 \\ 1 \\ 0 \\ 1
+\end{bmatrix}
+\]
+
+\[
+h(k) = T \cdot x \mod 2
+\]
+
+\phantom{TEMP LINEBREAK}
+
+We then perform XOR multiplication on the matrices:
+
+\[
+h(k) =
+\begin{bmatrix}
+(1 \oplus 0 \oplus 1 \oplus 1 \oplus 0 \oplus 1 \oplus 0 \oplus 1) \\
+(0 \oplus 1 \oplus 1 \oplus 0 \oplus 1 \oplus 0 \oplus 1 \oplus 1) \\
+(1 \oplus 1 \oplus 0 \oplus 1 \oplus 0 \oplus 1 \oplus 1 \oplus 0) \\
+(1 \oplus 0 \oplus 1 \oplus 1 \oplus 1 \oplus 0 \oplus 1 \oplus 0)
+\end{bmatrix}
+\]
+
+Which finally simplifies to:
+
+\[
+h(k) =
+\begin{bmatrix}
+1 \\
+0 \\
+1 \\
+1
+\end{bmatrix}
+\]
+
+This allows us to intepret the result as a random 4-bit integer:
+
+\[
+h(k) = (1011)_2 = 11_{10}
+\]
+
+Whereas this example is quite small in order to demonstrate how the algorithm operates, this can of course be scaled up significantly with larger values for both \(T\) and \(k\) respectively, granting larger integers.
+The main focus of this work is implementing this algorithm as efficiently as possible on our MCUs, and as such, several optimization efforts need to be taken into account during our experimentation.
 
@@
 Neither of these implementations operate on constrained hardware, instead creating bespoke circuit boards for their works.
+
+Another important point that is often neglected in randomness extraction using Toeplitz matrices is how the seed key is handled when forming the Toeplitzmatrix.
+Numerous systems, such as those developed by Chouhan et al.
+\cite{toeplitz-desc} and Zhang et al.
+\cite{toeplitz}, depends on fixed seeds.
+Fixed seeds can create security risks over prolonged operation time.
+To tackle this problem, Lin et al \cite{lin} proposed a method for seed-renewable Toeplitz post-processing in QRNG.
+Their strategy incorporates a dynamic seed pool within the FPGA, where each instance of post-processing picks a new, randomly selected seed.
+Thus, minimizing temporal correlations between extractions.
+Furthermore, an external seed updateing mechanism via PCIe ensures that seeds are refreshed whenever a certain secutiry limit is reached.
+Compared to fixed-seed methods, this renewable approach enhances cryptographic robustness and ensures sustained high-security randomness extraction in real-world applications.
+
+Efficient Toeplitz matrix-vector multiplication (TMVM) is critical for optimizing randomness extraction that relies on Toeplitz, particulary in constrained hardware environments.
+Liao et al.
+\cite{liao} showed that this process could be greatly accelerated using Fast Fourier Transform (FFT) and its inverser (IFFT).
+Thus, reducing computational complexity from \(O(n^2)\) to \(O(n log n)\).
+Their implementation on FPGA utilized this approach for deep neural networks, resulting in a 28.7 times decrease in model size while still achieving fast inference speeds.
+By using FFT and IFFT acceleration, Toeplitz post-processing for randomness extraction could achieve higher throughput.
+Thus, potentially could improve performance and reduce latency.
 

### 2025-03-14 docs/style/article-body.tex
linebreak
final corrections before hand-in

@@
 
-In computer science, there are many applications for randomly generated numbers. From generating keys for cryptography
-
+In computer science, there are many applications for randomly generated numbers.
+From generating keys for cryptography, salting password hashes, load balancing in distributed systems, memory addressing and so much more.
 However, the process of producing these random numbers tends to be pseudo-random, e.g.~utilizing the current states of various modules \cite{randomness}.
@@
 
-As described in Clason's work \cite{Clason2023}, the OQRNG-device is an electro-optical system which measures optical shot noise, generating quantum randomness.
+As described in Clasons work \cite{Clason2023}, the OQRNG-device is an electro-optical system which measures optical shot noise, generating quantum randomness.
 The device has an LED and a photodiode positioned a few millimeters apart, ensuring efficient light coupling.
@@
 Various development frameworks and tools are available for programming MCUs, including the Arduino Framework, manufacturer specific SDKs and raw C++, among others.
-Each framework presents different trade-offs in terms of performance overhead, efficiency and ease of development.
+Each framework presents different trade-offs in terms of performance overhead, efficiency, and ease of development.
 
@@
 Thus, making them highly effective for Toeplitz matrix-vector multiplications.
-Although, the Cortex-M7 does not have special features for cryptography, its ability to execute multiple instructions quickly, combined with fast memory access and an efficient pipeline design, allows it to effectively manage and process randomness extraction.
+The Cortex-M7 does not have special features for cryptography, its ability to execute multiple instructions quickly, combined with fast memory access and an efficient pipeline design, allows it to effectively manage and process randomness extraction.
 Moreover, its USB High-Speed (480 Mbit/sec) interface, supporting both USB device and USB host modes, allows rapid transmission of extracted random data to an external system, ensuring minimal bottlenecks in high-rate randomness generation.
@@
 \cite{toeplitz-desc}), the sampled raw bit matrix (\(T\)) are multiplied with a pre-determined seed matrix (\(K\)).
-The size of the seed is directly dependant on the size of the sampled data, and can be fixed or continually re-sampled as needed.
+The size of the seed is directly dependent on the size of the sampled data, and can be fixed or continually re-sampled as needed.
 To ensure high levels of entropy, our intuition is that re-sampling the seed from the OQRNG-device continually is prudent.
@@
 
-This allows us to intepret the result as a random 4-bit integer:
+This allows us to interpret the result as a random 4-bit integer:
 
@@
 
-Another important point that is often neglected in randomness extraction using Toeplitz matrices is how the seed key is handled when forming the Toeplitzmatrix.
+Another important point that is often neglected in randomness extraction using Toeplitz matrices is how the seed key is handled when forming the Toeplitz matrix.
 Numerous systems, such as those developed by Chouhan et al.
@@
 Thus, minimizing temporal correlations between extractions.
-Furthermore, an external seed updateing mechanism via PCIe ensures that seeds are refreshed whenever a certain secutiry limit is reached.
+Furthermore, an external seed updating mechanism via PCIe ensures that seeds are refreshed whenever a certain security limit is reached.
 Compared to fixed-seed methods, this renewable approach enhances cryptographic robustness and ensures sustained high-security randomness extraction in real-world applications.
@@
 Liao et al.
-\cite{liao} showed that this process could be greatly accelerated using Fast Fourier Transform (FFT) and its inverser (IFFT).
+\cite{liao} showed that this process could be greatly accelerated using Fast Fourier Transform (FFT) and its inverse (IFFT).
 Thus, reducing computational complexity from \(O(n^2)\) to \(O(n log n)\).
@@
 Whereas this introduces some overhead, it drastically reduces the time spent per iteration, and is deemed to be a suitable starting point.
+\newpage
 
@@
 
-2025-03-12: Elaborated on background, as well as adding more details regarding hardware.
+2025-03-13: Elaborated on background, as well as adding more details regarding hardware.
 Note that the hardware selected is subject to change over time.

### 2025-04-22 project/tests/pipe.py
added some article details, adding pipe script and processed data. still things to work on for the test script

@@ -1,0 +1,74 @@
+#!/usr/bin/env python3
+
+import sys
+import serial
+import time
+import re
+
+if len(sys.argv) != 3:
+  print('Usage: python3 pipe.py <input.bin> <serial_port>')
+  sys.exit(1)
+
+input_path = f'data/{sys.argv[1]}'
+serial_path = sys.argv[2]
+output_path = f'results/{sys.argv[1]}'
+
+try:
+  with serial.Serial(serial_path, 115200, timeout=1) as ser:
+    time.sleep(2)
+    ser.reset_input_buffer()
+
+    print(f'Sending bits from {input_path} to {serial_path}, please wait...')
+    with open(input_path, 'rb') as f:
+      for byte in f.read():
+        bits = format(byte, '08b')
+        _ = ser.write(bits.encode())
+        time.sleep(0.0005)
+
+    print('Data sent to MCU, reading output...')
+
+    buffer = ''
+    extracted_outputs: list[str] = []
+
+    # read until timeout
+    while True:
+      chunk = ser.read(128).decode(errors='ignore')
+      if not chunk:
+        break
+      buffer += chunk
+    
+    buffer = buffer.split('out:')
+    for i, bits in enumerate(buffer):
+      bits = bits.strip()
+      bits = ''.join(b for b in bits if b in '01')
+
+      if not bits:
+        print(f'{i}:\tWARN: Block contained invalid characters: {bits}')
+        continue
+
+      print(f'{i}:\t{bits[:8]}... ({len(bits)} bits)')
+      extracted_outputs.append(bits)
+
+    all_bits = ''.join(extracted_outputs)
+
+    print(f'Total blocks received: {len(extracted_outputs)}')
+    print(f'Total bits collected: {len(all_bits)}')
+    print(f'Writing binary output to {output_path}...')
+
+    byte_chunks = [all_bits[i:i+8] for i in range(0, len(all_bits), 8)]
+
+    with open(output_path, 'wb') as f:
+      for i, chunk in enumerate(byte_chunks):
+        if len(chunk) < 8:
+          chunk = chunk.ljust(8, '0')  # pad final byte
+        _ = f.write(int(chunk, 2).to_bytes(1, byteorder='big'))
+
+    print(f'Wrote {len(byte_chunks)} bytes to "{output_path}"')
+
+except serial.SerialException as e:
+  print(f'Serial error: {e}')
+  sys.exit(1)
+
+except FileNotFoundError as e:
+  print(f'File error: {e}')
+  sys.exit(1)

### 2025-04-22 project/utils/pipe.py

@@ -1,13 +1,53 @@
-import sys
+# import sys
+#
+# if len(sys.argv) != 3:
+#   print("Usage: python3 pipe.py input.bin /dev/tty.usbmodemXXXX")
+#   sys.exit(1)
+#
+# input_file = sys.argv[1]
+# serial_port = sys.argv[2]
+#
+# with open(input_file, "rb") as f, open(serial_port, "wb", buffering=0) as out:
+#   for byte in f.read():
+#     bits = format(byte, "08b")
+#     out.write(bits.encode())
 
-if len(sys.argv) != 3:
-  print("Usage: python3 pipe.py input.bin /dev/tty.usbmodemXXXX")
+import sys
+import serial
+import time
+
+if len(sys.argv) != 4:
+  print("Usage: python3 pipe.py input.bin /dev/tty.usbmodemXXXX output.bin")
   sys.exit(1)
 
-input_file = sys.argv[1]
-serial_port = sys.argv[2]
+input_path = sys.argv[1]
+serial_path = sys.argv[2]
+output_path = sys.argv[3]
 
-with open(input_file, "rb") as f, open(serial_port, "wb", buffering=0) as out:
-  for byte in f.read():
-    bits = format(byte, "08b")
-    out.write(bits.encode())
+with serial.Serial(serial_path, 115200, timeout=1) as ser:
+  time.sleep(2)
+  ser.reset_input_buffer()
+
+  with open(input_path, "rb") as f:
+    for byte in f.read():
+      bits = format(byte, "08b")
+      ser.write(bits.encode())
+      time.sleep(0.0005)
+
+  extracted_bits = ""
+  while True:
+    if ser.in_waiting:
+      line = ser.readline().decode(errors='ignore').strip()
+      if line.startswith("out:"):
+        extracted_bits += line[4:]
+    else:
+      break
+
+byte_chunks = [extracted_bits[i:i+8] for i in range(0, len(extracted_bits), 8)]
+with open(output_path, "wb") as f:
+  for chunk in byte_chunks:
+    if len(chunk) < 8:
+      chunk = chunk.ljust(8, '0')
+    f.write(int(chunk, 2).to_bytes(1, byteorder='big'))
+
+print(f"Wrote {len(byte_chunks)} bytes of extracted output to {output_path}")

### 2025-04-23 docs/style/article-body.tex
added some details from initial experimentation

@@
 However, there are still considerations when implementing the firmware for the microcontroller in order to optimize the code.
-Our goal is to attempt several implementations in order to find the most optimal implementation with the least amount of CPU-cycles.
+Our goal is to attempt several implementations in order to find the most optimal implementation with the least amount of effective processing time spent on the algorithm.
 
@@
 
-Initially, we will utilize MAX11102AUB\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} with an effective sample rate of 2 million samples per second (MSPS).
+Many microcontrollers furthermore come equipped with internal ADCs that can be utilized, and while these provide a lower sample size (often around 1 MSPS), the ease of development may be prudent to utilize for this proof-of-concept.
+While our initial ADC has a fairly low throughput, this can always be upgraded if it ends up becoming too limiting.
+
+Should these internal ADCs prove too limiting, we propose utilizing MAX11102AUB\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} with an effective sample rate of 2 million samples per second (MSPS).
 This ADC provides a 12 bit sample size, providing roughly 2.86 MB/s of sampled data per second, derived by the following calculation.
@@
 
-Many microcontrollers furthermore come equipped with internal ADCs that can be utilized, and while these provide a lower sample size (often around 1 MSPS), the ease of development may be prudent to utilize for this proof-of-concept.
-While our initial ADC has a fairly low throughput, this can always be upgraded if it ends up becoming too limiting.
 The final output from the ADC, whether built into the microcontroller or an external one, will be a stream of raw bits, as the analog signal from the OQRNG-device is processed.
@@
 To summarize the theoretical working of Toeplitz extraction (as explained by Chouhan et al.
-\cite{toeplitz-desc}), the sampled raw bit matrix (\(T\)) are multiplied with a pre-determined seed matrix (\(K\)).
+\cite{toeplitz-desc}), the sampled raw bit matrix (\(T\)) is multiplied with a pre-determined seed matrix (\(K\)).
 The size of the seed is directly dependent on the size of the sampled data, and can be fixed or continually re-sampled as needed.
@@
 With the consideration that our work revolves around optimizing Toeplitz extraction in order to quickly process random bits into a random number, we will take an iterative approach.
-For our tests, we will use a pre-defined stream of raw bits which is loaded into memory on the microcontroller, and run several different implementations of Toeplitz extraction to produce numbers.
+For our tests, we will use a pre-defined stream of raw bits which is sent to the microcontroller via USB, and run several different implementations of Toeplitz extraction to produce numbers.
 As we always use a pre-defined bitstream, the result will at this stage be deterministic, giving us a clear indication whether the algorithm works as intended.
@@
 This naive implementation will then be flashed to our microcontrollers, beginning with Teensy 4.1 as this is the more capable of the microcontrollers used for this experiment.
-Code to measure the execution speed in milliseconds will be implemented and tested before we load the naive implementation on said microcontroller.
+Code to measure the execution speed in microseconds will be implemented and tested before we load the naive implementation on said microcontroller.
 We expect that several implementations may be too resource intensive or have a memory complexity far greater than our cheaper, less capable microcontroller are able to handle, and as such these may not be able to be tested until a few iterations of optimization has occurred.
@@
 
+In order to evaluate the efficacy of each iteration, we will test the code by processing our pre-determined bitstreams, saving the processed bits in a binary file and evaluating this file with the command line utility \texttt{ent}.
+This utility runs a series of tests on the data to determine its level of entropy the file has -- rather, how information dense the file is.
+A perfect result (\emph{e.g.
+entirely random}) has an entropy value of \(8.0\), whereas acceptable entropy values for our tests are greater than \(7.99\), denoting a high level of randomness.
+Furthermore, \texttt{ent} can verify the arithmetic mean of the file, summing all the bits in the file and dividing it by the length of the file.
+A good distribution of randomness should have an arithmetic mean of \(~127.5\).
+To ensure each iteration works as expected, we will compare the \texttt{ent} result from each iteration against the original baseline, ensuring that the resulting processed bitstring is identical between iterations.
+There are a few more measurements shown by \texttt{ent}, and more details regarding these can be found in the utilities manual page\footnote{\href{https://manpages.ubuntu.com/manpages/trusty/man1/ent.1.html}{Manual page for \texttt{ent}, accessed 2025-04-23.}}.
+
 As discussed in section 3, the hardware used will impose a clear bound on how quickly our implementation needs to process the bits in order to match the speed of the ADC, as well as the output speed of the USB-port, both in \(MB/s\).
@@
 
-Each iteration results in an implementation written in C, all of which will be supplied as an appendix to this paper.
+Each iteration results in an implementation written in C++, all of which will be supplied as an appendix to this paper.
 Furthermore, to simplify the implementation process, we will use the Arduino framework to communicate with the hardware.
 Whereas this introduces some overhead, it drastically reduces the time spent per iteration, and is deemed to be a suitable starting point.
+
+Initially, we created a script to simulate a stream of bits from a file, and pre-computed the expected results using a trusted version of Toeplitz extraction.
+With the expected results saved to a file on the host computer, each iteration could then be deployed to the MCU, and tested with the same data as the baseline.
+As the algorithm is fully deterministic with this uniform data, the exact same results in the same order should be produced by all iterations.
+The simulation script then simply asserts that the MCU produces the same random string of bits from the resulting iterations as with the baseline.
+
+As our system does not rely on interrupts in any capacity, measuring the execution speed of the implemented Toeplitz extraction function can be done using the built in \texttt{micros()}-function provided by Arduino.
+This provides us with an accurate measurement of the execution time per function call.
+The average execution time per test set over each iteration will be presented in \textbf{\texttt{TABLE\ 1}}, along with the fastest and slowest executions per iteration respectively.
+Furthermore, the entropy value on the processed bitstream will be noted in the table to demonstrate that the important measurement, e.g.~the randomness, still holds true between iterations.
+
 \newpage
@@
 Further elaborated on related works in optimizing Toeplitz extraction.
+
+2025-04-22: Begun including details regarding initial experimentation, updating details regarding experiments that had to change (e.g.~no baseline on separate hardware).

### 2025-04-23 docs/style/article-final.tex
added some details from initial experimentation

@@
 Further elaborated on related works in optimizing Toeplitz extraction.
+
+2025-04-22: Begun including details regarding initial experimentation, updating details regarding experiments that had to change (e.g.~no baseline on separate hardware).
 % BALANCE COLUMNS

### 2025-04-24 project/tests/pipe.py
added filtering output from mcu to pipe.py & led indicator

@@ -4,71 +4,95 @@ import sys
 import serial
 import time
 import re
+import datetime
 
 if len(sys.argv) != 3:
-  print('Usage: python3 pipe.py <input.bin> <serial_port>')
-  sys.exit(1)
+    print('Usage: python3 pipe.py <input.bin> <serial_port>')
+    sys.exit(1)
 
 input_path = f'data/{sys.argv[1]}'
 serial_path = sys.argv[2]
 output_path = f'results/{sys.argv[1]}'
 
+
+def filter_debug(buffer: str):
+    timestamp = datetime.datetime.now().strftime("%Y%m%d_%H%M%S")
+    log_path = f'logs/debug_{timestamp}.log'
+
+    debug_buff = []
+    data_buff = []
+    for line in buffer.splitlines():
+        line = line.strip()
+        if line.startswith('['):
+            debug_buff.append(line)
+        else:
+            data_buff.append(line)
+
+    with open(log_path, 'w') as log_file:
+        log_file.write('\n'.join(debug_buff))
+
+    print(f'Saved debug output to {log_path}')
+    return '\n'.join(data_buff)
+
+
 try:
-  with serial.Serial(serial_path, 115200, timeout=1) as ser:
-    time.sleep(2)
-    ser.reset_input_buffer()
+    with serial.Serial(serial_path, 115200, timeout=1) as ser:
+        # time.sleep(0.00005)
+        ser.reset_input_buffer()
 
-    print(f'Sending bits from {input_path} to {serial_path}, please wait...')
-    with open(input_path, 'rb') as f:
-      for byte in f.read():
-        bits = format(byte, '08b')
-        _ = ser.write(bits.encode())
-        time.sleep(0.0005)
+        print(
+            f'Sending bits from {input_path} to {serial_path}, please wait...')
+        with open(input_path, 'rb') as f:
+            for byte in f.read():
+                bits = format(byte, '08b')
+                _ = ser.write(bits.encode())
+                time.sleep(0.0005)
 
-    print('Data sent to MCU, reading output...')
+        print('Data sent to MCU, reading output...')
 
-    buffer = ''
-    extracted_outputs: list[str] = []
+        buffer = ''
+        extracted_outputs: list[str] = []
 
-    # read until timeout
-    while True:
-      chunk = ser.read(128).decode(errors='ignore')
-      if not chunk:
-        break
-      buffer += chunk
-    
-    buffer = buffer.split('out:')
[…]

### 2025-04-24 project/tests/tests.py

@@ -9,98 +9,105 @@ import glob
 
 SERIAL_PORTNO = 115200
 
+
 def serial_select() -> str:
-  os = platform.system()
-  port = None
+    os = platform.system()
+    port = None
 
-  match os:
-    case 'Linux':
-      port = glob.glob('/dev/ttyACM*') + glob.glob('/dev/ttyUSB*')
-    case 'Darwin':
-      port = glob.glob('/dev/tty.usbmodem*') + glob.glob('/dev/tty.usbserial*') + glob.glob('/dev/cu.usbmodem*')
-    case 'Windows':
-      print('Error: Ew. Exiting due to cringe.')
-      sys.exit(1)
-    case _:
-      print(f'Error: Unsupported OS "{os}", exiting.')
-      sys.exit(1)
+    match os:
+        case 'Linux':
+            port = glob.glob('/dev/ttyACM*') + glob.glob('/dev/ttyUSB*')
+        case 'Darwin':
+            port = glob.glob('/dev/tty.usbmodem*') + \
+                glob.glob('/dev/tty.usbserial*') + \
+                glob.glob('/dev/cu.usbmodem*')
+        case 'Windows':
+            port = ['COM3']
+        case _:
+            print(f'Error: Unsupported OS "{os}", exiting.')
+            sys.exit(1)
 
-  if len(port) == 0:
-    print('Error: No serial device found. Please connect a microcontroller and try again.')
-    sys.exit(1)
+    if len(port) == 0:
+        print(
+            'Error: No serial device found. Please connect a microcontroller and try again.')
+        sys.exit(1)
 
-  print('Available ports:')
-  for i, p in enumerate(port):
-    print(f'{i}: {p}')
+    print('Available ports:')
+    for i, p in enumerate(port):
+        print(f'{i}: {p}')
 
-  while True:
-    try:
-      sel = int(input('Select port: '))
-    except:
-      print('Not a number.')
-      continue
+    while True:
+        try:
+            sel = int(input('Select port: '))
+        except:
+            print('Not a number.')
+            continue
 
-    if sel >= 0 and sel < len(port):
-      return port[sel]
+        if sel >= 0 and sel < len(port):
+            return port[sel]
 
-    print(f'Invalid port {sel}.')
+        print(f'Invalid port {sel}.')
+
 
 def run_tests(port: str) -> None:
-  no_tests = len(glob.glob('data/*.txt'))
-  failed = 0
[…]

### 2025-04-25 project/tests/pipe.py
tests update

@@ -51,10 +51,6 @@ try:
 
     all_bits = ''.join(extracted_outputs)
 
-    print(f'Total blocks received: {len(extracted_outputs)}')
-    print(f'Total bits collected: {len(all_bits)}')
-    print(f'Writing binary output to {output_path}...')
-
     byte_chunks = [all_bits[i:i+8] for i in range(0, len(all_bits), 8)]
 
     with open(output_path, 'wb') as f:

### 2025-04-25 project/tests/tests.py

@@ -1,8 +1,12 @@
 #!/usr/bin/env python3
 
+from math import nan
+from statistics import mean
 import sys
+from typing import Any
 import serial
 import time
+import pprint
 
 import platform
 import glob
@@ -26,7 +30,7 @@ def serial_select() -> str:
       sys.exit(1)
 
   if len(port) == 0:
-    print('Error: No serial device found. Please connect a microcontroller and try again.')
+    print('❌ Error: No serial device found. Please connect a microcontroller and try again.')
     sys.exit(1)
 
   print('Available ports:')
@@ -43,56 +47,64 @@ def serial_select() -> str:
     if sel >= 0 and sel < len(port):
       return port[sel]
 
-    print(f'Invalid port {sel}.')
+    print(f'⚠️ Invalid port {sel}.')
 
 def run_tests(port: str) -> None:
-  no_tests = len(glob.glob('data/*.txt'))
+  tests = sorted(glob.glob('data/*.bin'))
+  results = sorted(glob.glob('results/*.bin'))
   failed = 0
 
+  test_results: dict[str, dict[str, Any]] = {}
+
   with serial.Serial(port, SERIAL_PORTNO, timeout=1) as ser:
-    for i in range(no_tests):
-      data = f'data/{i}.txt'
-      result = f'results/{i}.txt'
+    ser.reset_input_buffer()
+    ser.reset_output_buffer()
+    for test, res in zip(tests, results):
+      print(f'🕑 Sending bits from {test}, please wait...')
+      with open(test, 'rb') as f:
+        for byte in f.read():
+          bits = format(byte, '08b')
+          _ = ser.write(bits.encode())
+          time.sleep(0.0005)
 
-      with open(data, 'r') as f:
-        data_bits = f.read()
+      print('🕑 Data sent to MCU, reading output...')
 
-      with open(result, 'r') as f:
-        expected_outputs = [line.strip() for line in f if line.strip()]
+      extracted_outputs: list[str] = []
+      times: list[int] = []
 
-      print(f'Running test {i}:')
-      received_outputs: list[str] = []
-      ser.reset_input_buffer()
-
-      for bit in data_bits:
-        _ = ser.write(bit.encode())
-        time.sleep(0.0005)
-
-        if ser.in_waiting:
-          line = ser.readline().decode(errors='ignore').strip()
-          if line.startswith('out:'):
-            received_outputs.append(line[4:])
-
-      if len(received_outputs) != len(expected_outputs):
[…]

### 2025-04-25 project/tests/tests.py
tests with generation

@@ -1,18 +1,18 @@
 #!/usr/bin/env python3
 
-from math import nan
 from statistics import mean
 import sys
 from typing import Any
 import serial
-import time
 import pprint
 
 import platform
 import glob
+import os
 
 SERIAL_PORTNO = 115200
-
+BIT_CHUNK_SIZE = 64
+OUTPUT_LEN = 32
 
 def serial_select() -> str:
     os = platform.system()
@@ -28,22 +28,22 @@ def serial_select() -> str:
         case 'Windows':
             port = ['COM3']
         case _:
-            print(f'Error: Unsupported OS "{os}", exiting.')
+            print(f'❌ Error: Unsupported OS "{os}", exiting.')
             sys.exit(1)
 
     if len(port) == 0:
         print('❌ Error: No serial device found. Please connect a microcontroller and try again.')
         sys.exit(1)
 
-    print('Available ports:')
+    print('❓ Available ports:')
     for i, p in enumerate(port):
         print(f'{i}: {p}')
 
     while True:
         try:
-            sel = int(input('Select port: '))
+            sel = int(input('⌨️ Select port: '))
         except:
-            print('Not a number.')
+            print('❌ Not a number.')
             continue
 
         if sel >= 0 and sel < len(port):
@@ -52,70 +52,175 @@ def serial_select() -> str:
         print(f'⚠️ Invalid port {sel}.')
 
 def run_tests(port: str) -> None:
-  tests = sorted(glob.glob('data/*.bin'))
+  tests   = sorted(glob.glob('data/*.bin'))
   results = sorted(glob.glob('results/*.bin'))
-  failed = 0
+  failed  = 0
 
   test_results: dict[str, dict[str, Any]] = {}
 
   with serial.Serial(port, SERIAL_PORTNO, timeout=1) as ser:
     ser.reset_input_buffer()
     ser.reset_output_buffer()
-    for test, res in zip(tests, results):
-      print(f'🕑 Sending bits from {test}, please wait...')
-      with open(test, 'rb') as f:
-        for byte in f.read():
-          bits = format(byte, '08b')
-          _ = ser.write(bits.encode())
-          time.sleep(0.0005)
 
-      print('🕑 Data sent to MCU, reading output...')
+    for test_path, result_path in zip(tests, results):
+      with open(test_path, 'rb') as f:
+        data_bytes = f.read()
+      all_bits = ''.join(format(b, '08b') for b in data_bytes)
 
       extracted_outputs: list[str] = []
[…]

### 2025-04-25 project/tests/tests.py
tests with generation v2

@@ -187,20 +187,19 @@ def generate_results(port: str) -> None:
 
       for i in range(0, len(all_bits), BIT_CHUNK_SIZE):
         chunk = all_bits[i:i + BIT_CHUNK_SIZE]
-        _ = ser.write(chunk.encode())
+        ser.write(chunk.encode())
         ser.flush()
 
-        while True:
-          line = ser.readline().decode(errors='ignore').strip()
-          if not line:
-            continue
-          if not line.startswith('out:'):
-            print(f'WARN: unexpected MCU line: {line}')
-            continue
+        # read exactly one response, break on timeout or error
+        line = ser.readline()  # timeout=1s
+        if not line:
+          raise RuntimeError("Timeout waiting for MCU response")
+        line = line.decode(errors='ignore').strip()
+        if not line.startswith('out:'):
+          raise RuntimeError(f"Bad MCU response: {line}")
 
-          payload, _, _ = line[4:].partition('took:')
-          chunks_out.append(payload)
-          break
+        payload, _, _ = line[4:].partition('took:')
+        chunks_out.append(payload)
 
       full_out = ''.join(chunks_out)
       byte_vals = [

### 2025-04-25 project/tests/tests.py
update

@@ -200,6 +200,8 @@ def generate_results(port: str) -> None:
 
         payload, _, _ = line[4:].partition('took:')
         chunks_out.append(payload)
+        print(payload)
+        break
 
       full_out = ''.join(chunks_out)
       byte_vals = [

### 2025-04-25 project/tests/tests.py
test script works

@@ -1,18 +1,18 @@
 #!/usr/bin/env python3
 
-from statistics import mean
 import sys
-from typing import Any
 import serial
-import pprint
-
 import platform
 import glob
 import os
+import re
 
-SERIAL_PORTNO = 115200
+SERIAL_BAUD = 6000000
 BIT_CHUNK_SIZE = 64
 OUTPUT_LEN = 32
+
+BIT_CHUNK_BYTES = 8
+OUT_BYTES = 4
 
 def serial_select() -> str:
     os = platform.system()
@@ -52,66 +52,51 @@ def serial_select() -> str:
         print(f'⚠️ Invalid port {sel}.')
 
 def run_tests(port: str) -> None:
-  tests   = sorted(glob.glob('data/*.bin'))
-  results = sorted(glob.glob('results/*.bin'))
-  failed  = 0
+    tests   = sorted(glob.glob('data/*.bin'))
+    results = sorted(glob.glob('results/*.bin'))
+    failed  = 0
 
-  test_results: dict[str, dict[str, Any]] = {}
+    with serial.Serial(port, SERIAL_BAUD, timeout=1) as ser:
+        ser.reset_input_buffer()
+        ser.reset_output_buffer()
+        for test_path, result_path in zip(tests, results):
+            data = open(test_path, 'rb').read()
+            chunks_out = []
+            for offset in range(0, len(data), BIT_CHUNK_BYTES):
+                chunk = data[offset : offset + BIT_CHUNK_BYTES]
+                if len(chunk) < BIT_CHUNK_BYTES:
+                    chunk = chunk.ljust(BIT_CHUNK_BYTES, b'\x00')
 
-  with serial.Serial(port, SERIAL_PORTNO, timeout=1) as ser:
-    ser.reset_input_buffer()
-    ser.reset_output_buffer()
+                written = ser.write(chunk)
+                if written != len(chunk):
+                    raise RuntimeError(
+                        f'❌ Wrote {written}/{len(chunk)} bytes; expected {len(chunk)}'
+                    )
 
-    for test_path, result_path in zip(tests, results):
-      with open(test_path, 'rb') as f:
-        data_bytes = f.read()
-      all_bits = ''.join(format(b, '08b') for b in data_bytes)
+                resp = ser.read(OUT_BYTES)
+                if len(resp) != OUT_BYTES:
+                    raise RuntimeError(
[…]

### 2025-04-25 project/tests/tests.py
finished test script

@@ -5,7 +5,8 @@ import serial
 import platform
 import glob
 import os
-import re
+from statistics import mean
+import argparse
 
 SERIAL_BAUD = 6000000
 BIT_CHUNK_SIZE = 64
@@ -51,6 +52,32 @@ def serial_select() -> str:
 
         print(f'⚠️ Invalid port {sel}.')
 
+def process_chunks(data, ser, ms=False):
+    chunks_out = []
+    for offset in range(0, len(data), BIT_CHUNK_BYTES):
+        chunk = data[offset : offset + BIT_CHUNK_BYTES]
+        if len(chunk) < BIT_CHUNK_BYTES:
+            chunk = chunk.ljust(BIT_CHUNK_BYTES, b'\x00')
+
+        written = ser.write(chunk)
+        if written != len(chunk):
+            raise RuntimeError(
+                f'❌ Wrote {written}/{len(chunk)} bytes; expected {len(chunk)}'
+            )
+
+        resp = ser.read(OUT_BYTES)
+        if len(resp) != OUT_BYTES:
+            raise RuntimeError(
+                f'❌ Incomplete response from MCU: got {len(resp)} bytes, expected {OUT_BYTES}'
+            )
+        if ms:
+             chunks_out.append(int.from_bytes(resp, 'big'))
+        else:
+            bits = ''.join(f'{b:08b}' for b in resp)
+            chunks_out.append(bits)
+    return chunks_out
+
+
 def run_tests(port: str) -> None:
     tests   = sorted(glob.glob('data/*.bin'))
     results = sorted(glob.glob('results/*.bin'))
@@ -61,26 +88,7 @@ def run_tests(port: str) -> None:
         ser.reset_output_buffer()
         for test_path, result_path in zip(tests, results):
             data = open(test_path, 'rb').read()
-            chunks_out = []
-            for offset in range(0, len(data), BIT_CHUNK_BYTES):
-                chunk = data[offset : offset + BIT_CHUNK_BYTES]
-                if len(chunk) < BIT_CHUNK_BYTES:
-                    chunk = chunk.ljust(BIT_CHUNK_BYTES, b'\x00')
-
-                written = ser.write(chunk)
-                if written != len(chunk):
-                    raise RuntimeError(
-                        f'❌ Wrote {written}/{len(chunk)} bytes; expected {len(chunk)}'
-                    )
-
-                resp = ser.read(OUT_BYTES)
-                if len(resp) != OUT_BYTES:
-                    raise RuntimeError(
[…]

### 2025-04-25 project/tests/tests.py
qol in test script

@@ -40,17 +40,21 @@ def serial_select() -> str:
     for i, p in enumerate(port):
         print(f'{i}: {p}')
 
-    while True:
-        try:
-            sel = int(input('⌨️ Select port: '))
-        except:
-            print('❌ Not a number.')
-            continue
+    if len(port) > 1:
+        while True:
+            try:
+                sel = int(input('⌨️ Select port: '))
+            except:
+                print('❌ Not a number.')
+                continue
 
-        if sel >= 0 and sel < len(port):
-            return port[sel]
+            if sel >= 0 and sel < len(port):
+                return port[sel]
 
-        print(f'⚠️ Invalid port {sel}.')
+            print(f'⚠️ Invalid port {sel}.')
+    else:
+        print(f'✅ Auto-selecting only port: {port[0]}')
+
 
 def process_chunks(data, ser, ms=False):
     chunks_out = []

### 2025-04-25 project/tests/tests.py
bugfix

@@ -54,6 +54,7 @@ def serial_select() -> str:
             print(f'⚠️ Invalid port {sel}.')
     else:
         print(f'✅ Auto-selecting only port: {port[0]}')
+        return port[0]
 
 
 def process_chunks(data, ser, ms=False):

### 2025-04-25 project/baseline/cpp/toeplitz.cpp
more qol and some details in article

@@ -1,76 +1,0 @@
-#include <chrono>
-#include <iostream>
-#include <stdexcept>
-#include <vector>
-
-using namespace std;
-
-enum LOG_LEVEL {
-  LOG_HIGH,
-  LOG_MED,
-  LOG_LOW
-};
-
-constexpr int NO_ITERATIONS{1000};
-constexpr LOG_LEVEL DEBUG{LOG_MED};
-
-// Needs at least m + n - 1 bits in seed where:
-//    m is the output
-//    n is the size of raw bits
-vector<bool> toeplitz_extraction(const vector<bool> &raw_bits,
-                                 const vector<bool> &seed_bits,
-                                 size_t output_len) {
-  size_t n{raw_bits.size()};
-  size_t m{output_len};
-
-  if (seed_bits.size() < m + n - 1) {
-    throw runtime_error{"Seed is smaller than m + n - 1"};
-  }
-
-  vector<bool> output(m, false);
-
-  for (size_t i{0}; i < m; i++) {
-    bool sum = false;
-    for (size_t j{0}; j < n; j++) {
-      if (raw_bits[j] && seed_bits[i + j]) {
-        sum = !sum;
-      }
-    }
-    output[i] = sum;
-  }
-
-  return output;
-}
-
-int main() {
-  vector<bool> raw_bits{true, false, true, true, false, true, false, false};
-
-  // Simulating seed bits rather than fetching from real source
-  // This should be fetched from sampled data later
-  // Constant or variable seed?
-  vector<bool> seed_bits{true, false, true,  false, true, false,
-                         true, true,  false, true,  false};
-
-  chrono::nanoseconds total_duration{0};
-
-  for (int i{0}; i < NO_ITERATIONS; i++) {
-    auto start{chrono::high_resolution_clock::now()};
-
-    vector<bool> extracted_bits{toeplitz_extraction(raw_bits, seed_bits, 4)};
-
-    auto end{chrono::high_resolution_clock::now()};
-    auto duration{chrono::duration_cast<chrono::nanoseconds>(end - start)};
-    if (DEBUG == LOG_HIGH) {
-      cout << "Iteration " << i << ": " << duration.count() << " ns." << endl;
-    } else if (DEBUG == LOG_MED) {
-      if (duration.count() > 300) {
-        cout << "Iteration " << i << ": " << duration.count() << " ns." << endl;
-      }
-    }
-    total_duration += duration;
-  }
-
[…]

### 2025-04-25 project/baseline/cpp/toeplitz_v2.cpp

@@ -1,48 +1,0 @@
-#include <iostream>
-#include <stdexcept>
-#include <vector>
-
-using namespace std;
-
-//raw bits - input from entropy source 
-//seed - a part of entropy, is it good?
-vector<int> toeplitz_extraction(const vector<int> &raw_bits,
-                                const vector<int> &seed_bits,
-                                size_t output_len) {
-  size_t n = raw_bits.size();
-  size_t m = output_len;
-
-  if (seed_bits.size() < m + n - 1) {
-    throw runtime_error("Seed is smaller than m + n - 1");
-  }
-
-  vector<int> output(m, 0);
-
-  for (size_t i = 0; i < m; i++) {
-    int sum = 0;
-    for (size_t j = 0; j < n; j++) {
-      sum += raw_bits[j] * seed_bits[i + j];
-    }
-    output[i] = sum % 2;
-  }
-
-  return output;
-}
-
-//now, how to make it real time? 
-int main() {
-  vector<int> raw_bits{1, 1, 0, 0};
-
-  vector<int> seed_bits{1, 1, 0, 1, 0, 1};
-
-  vector<int> extracted_bits = toeplitz_extraction(raw_bits, seed_bits, 3);
-
-  cout << "Extracted bits: ";
-  for (int bit : extracted_bits) {
-    cout << bit;
-  }
-  cout << endl;
-
-  return 0;
-}
-

### 2025-04-25 project/iter3/tests/pipe.py

@@ -1,0 +1,102 @@
+#!/usr/bin/env python3
+
+import sys
+import serial
+import time
+import re
+import datetime
+
+if len(sys.argv) != 3:
+    print('Usage: python3 pipe.py <input.bin> <serial_port>')
+    sys.exit(1)
+
+input_path = f'data/{sys.argv[1]}'
+serial_path = sys.argv[2]
+output_path = f'results/{sys.argv[1]}'
+
+
+def filter_debug(buffer: str):
+    timestamp = datetime.datetime.now().strftime("%Y%m%d_%H%M%S")
+    log_path = f'logs/debug_{timestamp}.log'
+
+    debug_buff = []
+    data_buff = []
+    for line in buffer.splitlines():
+        line = line.strip()
+        if line.startswith('['):
+            debug_buff.append(line)
+        else:
+            data_buff.append(line)
+
+    with open(log_path, 'w') as log_file:
+        log_file.write('\n'.join(debug_buff))
+
+    print(f'Saved debug output to {log_path}')
+    return '\n'.join(data_buff)
+
+
+try:
+    with serial.Serial(serial_path, 115200, timeout=1) as ser:
+        # time.sleep(0.00005)
+        ser.reset_input_buffer()
+
+        print(
+            f'Sending bits from {input_path} to {serial_path}, please wait...')
+        with open(input_path, 'rb') as f:
+            for byte in f.read():
+                bits = format(byte, '08b')
+                _ = ser.write(bits.encode())
+                time.sleep(0.0005)
+
+        print('Data sent to MCU, reading output...')
+
+        buffer = ''
+        extracted_outputs: list[str] = []
+
+        # read until timeout
+        while True:
+            chunk = ser.read(128).decode(errors='ignore')
+            if not chunk:
+                break
+            buffer += chunk
+
+        buffer = filter_debug(buffer)
+        buffer = buffer.split('out:')
+        for i, bits in enumerate(buffer):
+            bits = bits.strip()
+            bits = ''.join(b for b in bits if b in '01')
+
+            if not bits:
+                print(f'{i}:\tWARN: Block contained invalid characters: {bits}')
+                continue
+
+            print(f'{i}:\t{bits[:8]}... ({len(bits)} bits)')
+            extracted_outputs.append(bits)
+
+<<<<<<< HEAD
[…]

### 2025-04-25 project/iter4/tests/tests.py

@@ -7,17 +7,13 @@ import glob
 import os
 from statistics import mean
 import argparse
-import subprocess
-from pathlib import Path
 
 SERIAL_BAUD     = 6000000
-BIT_CHUNK_SIZE  = 2048
-OUTPUT_LEN      = BIT_CHUNK_SIZE // 2
+BIT_CHUNK_SIZE  = 1024
+OUTPUT_LEN      = 512
 
 BIT_CHUNK_BYTES = BIT_CHUNK_SIZE // 8
 OUT_BYTES = OUTPUT_LEN // 8
-
-current_iter = 0
 
 def serial_select() -> str:
     os = platform.system()
@@ -61,7 +57,7 @@ def serial_select() -> str:
         return port[0]
 
 
-def process_chunks(data, ser, ms=False, raw=False):
+def process_chunks(data, ser, ms=False):
     if ms:
         output_chunk_bytes = 4
     else:
@@ -86,23 +82,16 @@ def process_chunks(data, ser, ms=False, raw=False):
             )
         if ms:
              chunks_out.append(int.from_bytes(resp, 'big'))
-        elif raw:
-            chunks_out.append(resp)
         else:
             bits = ''.join(f'{b:08b}' for b in resp)
             chunks_out.append(bits)
-
     return chunks_out
 
 
 def run_tests(port: str) -> None:
     tests   = sorted(glob.glob('data/*.bin'))
-    results = sorted(glob.glob(f'results/{OUTPUT_LEN}/*.bin'))
+    results = sorted(glob.glob('results/*.bin'))
     failed  = 0
-
-    if len(tests) != len(results):
-        print('❌ Mismatch in test and results. Generate new ones first!')
-        sys.exit(1)
 
     with serial.Serial(port, SERIAL_BAUD, timeout=1) as ser:
         ser.reset_input_buffer()
@@ -130,9 +119,7 @@ def run_tests(port: str) -> None:
 
 def generate_results(port: str) -> None:
     tests = sorted(glob.glob('data/*.bin'))
-    os.makedirs(f'results/{OUTPUT_LEN}', exist_ok=True)
-
-    ent_string = ''
+    os.makedirs('results', exist_ok=True)
 
     with serial.Serial(port, SERIAL_BAUD, timeout=1) as ser:
         ser.reset_input_buffer()
@@ -140,36 +127,18 @@ def generate_results(port: str) -> None:
 
         for test_path in tests:
             print(f'🕑 Running {test_path}…')
[…]

### 2025-04-25 project/iter4/tests/pipe.py
some

@@ -1,102 +1,0 @@
-#!/usr/bin/env python3
-
-import sys
-import serial
-import time
-import re
-import datetime
-
-if len(sys.argv) != 3:
-    print('Usage: python3 pipe.py <input.bin> <serial_port>')
-    sys.exit(1)
-
-input_path = f'data/{sys.argv[1]}'
-serial_path = sys.argv[2]
-output_path = f'results/{sys.argv[1]}'
-
-
-def filter_debug(buffer: str):
-    timestamp = datetime.datetime.now().strftime("%Y%m%d_%H%M%S")
-    log_path = f'logs/debug_{timestamp}.log'
-
-    debug_buff = []
-    data_buff = []
-    for line in buffer.splitlines():
-        line = line.strip()
-        if line.startswith('['):
-            debug_buff.append(line)
-        else:
-            data_buff.append(line)
-
-    with open(log_path, 'w') as log_file:
-        log_file.write('\n'.join(debug_buff))
-
-    print(f'Saved debug output to {log_path}')
-    return '\n'.join(data_buff)
-
-
-try:
-    with serial.Serial(serial_path, 115200, timeout=1) as ser:
-        # time.sleep(0.00005)
-        ser.reset_input_buffer()
-
-        print(
-            f'Sending bits from {input_path} to {serial_path}, please wait...')
-        with open(input_path, 'rb') as f:
-            for byte in f.read():
-                bits = format(byte, '08b')
-                _ = ser.write(bits.encode())
-                time.sleep(0.0005)
-
-        print('Data sent to MCU, reading output...')
-
-        buffer = ''
-        extracted_outputs: list[str] = []
-
-        # read until timeout
-        while True:
-            chunk = ser.read(128).decode(errors='ignore')
-            if not chunk:
-                break
-            buffer += chunk
-
-        buffer = filter_debug(buffer)
-        buffer = buffer.split('out:')
-        for i, bits in enumerate(buffer):
-            bits = bits.strip()
-            bits = ''.join(b for b in bits if b in '01')
-
-            if not bits:
-                print(f'{i}:\tWARN: Block contained invalid characters: {bits}')
-                continue
-
-            print(f'{i}:\t{bits[:8]}... ({len(bits)} bits)')
-            extracted_outputs.append(bits)
-
-<<<<<<< HEAD
[…]

### 2025-04-25 project/iter4/tests/tests.py

@@ -1,215 +1,0 @@
-#!/usr/bin/env python3
-
-import sys
-import serial
-import platform
-import glob
-import os
-from statistics import mean
-import argparse
-
-SERIAL_BAUD     = 6000000
-BIT_CHUNK_SIZE  = 1024
-OUTPUT_LEN      = 512
-
-BIT_CHUNK_BYTES = BIT_CHUNK_SIZE // 8
-OUT_BYTES = OUTPUT_LEN // 8
-
-def serial_select() -> str:
-    os = platform.system()
-    port = None
-
-    match os:
-        case 'Linux':
-            port = glob.glob('/dev/ttyACM*') + glob.glob('/dev/ttyUSB*')
-        case 'Darwin':
-            port = glob.glob('/dev/tty.usbmodem*') + \
-                glob.glob('/dev/tty.usbserial*') + \
-                glob.glob('/dev/cu.usbmodem*')
-        case 'Windows':
-            port = ['COM3']
-        case _:
-            print(f'❌ Error: Unsupported OS "{os}", exiting.')
-            sys.exit(1)
-
-    if len(port) == 0:
-        print('❌ Error: No serial device found. Please connect a microcontroller and try again.')
-        sys.exit(1)
-
-    print('❓ Available ports:')
-    for i, p in enumerate(port):
-        print(f'{i}: {p}')
-
-    if len(port) > 1:
-        while True:
-            try:
-                sel = int(input('⌨️ Select port: '))
-            except:
-                print('❌ Not a number.')
-                continue
-
-            if sel >= 0 and sel < len(port):
-                return port[sel]
-
-            print(f'⚠️ Invalid port {sel}.')
-    else:
-        print(f'✅ Auto-selecting only port: {port[0]}')
-        return port[0]
-
-
-def process_chunks(data, ser, ms=False):
-    if ms:
-        output_chunk_bytes = 4
-    else:
-        output_chunk_bytes = OUT_BYTES
-
-    chunks_out = []
-    for offset in range(0, len(data), BIT_CHUNK_BYTES):
-        chunk = data[offset : offset + BIT_CHUNK_BYTES]
-        if len(chunk) < BIT_CHUNK_BYTES:
-            chunk = chunk.ljust(BIT_CHUNK_BYTES, b'\x00')
-
-        written = ser.write(chunk)
-        if written != len(chunk):
-            raise RuntimeError(
[…]

### 2025-04-27 docs/style/article-body.tex
minor correction
first tabel, updated teensy makefile

@@
-\section{1 INTRODUCTION}\label{introduction}
+\hypertarget{introduction}{%
+\section{1 INTRODUCTION}\label{introduction}}
 
@@
 
-\subsection{2 THEORY}\label{theory}
+\hypertarget{theory}{%
+\subsection{2 THEORY}\label{theory}}
 
@@
 
-\subsubsection{2.1 Shot noise quantum fluctuations}\label{shot-noise-quantum-fluctuations}
+\hypertarget{shot-noise-quantum-fluctuations}{%
+\subsubsection{2.1 Shot noise quantum fluctuations}\label{shot-noise-quantum-fluctuations}}
 
@@
 
-\subsection{3 BACKGROUND}\label{background}
+\hypertarget{background}{%
+\subsection{3 BACKGROUND}\label{background}}
 
@@
 
-\subsubsection{3.1 Optical RNG module}\label{optical-rng-module}
+\hypertarget{optical-rng-module}{%
+\subsubsection{3.1 Optical RNG module}\label{optical-rng-module}}
 
@@
 
-\subsubsection{3.2 ADC converter}\label{adc-converter}
+\hypertarget{adc-converter}{%
+\subsubsection{3.2 ADC converter}\label{adc-converter}}
 
@@
 
-\subsubsection{3.3 Microcontroller}\label{microcontroller}
+\hypertarget{microcontroller}{%
+\subsubsection{3.3 Microcontroller}\label{microcontroller}}
 
@@
 
-\subsubsection{\texorpdfstring{3.4 Toeplitz extraction }{3.4 Toeplitz extraction }}\label{toeplitz-extraction}
+\hypertarget{toeplitz-extraction}{%
+\subsubsection{\texorpdfstring{3.4 Toeplitz extraction }{3.4 Toeplitz extraction }}\label{toeplitz-extraction}}
 
@@
 To summarize the theoretical working of Toeplitz extraction (as explained by Chouhan et al.
-\cite{toeplitz-desc}), the sampled raw bit matrix (\(T\)) is multiplied with a pre-determined seed matrix (\(K\)).
+\cite{toeplitz-desc}), a pre-determined seed matrix (\(T\)) is multiplied with the sampled raw bit matrix (\(K\)).
 The size of the seed is directly dependent on the size of the sampled data, and can be fixed or continually re-sampled as needed.
@@
 
-\subsubsection{3.5 Summary}\label{summary}
+\hypertarget{summary}{%
+\subsubsection{3.5 Summary}\label{summary}}
 
@@
 
-\subsection{\texorpdfstring{4 RELATED WORKS }{4 RELATED WORKS }}\label{related-works}
+\hypertarget{related-works}{%
+\subsection{\texorpdfstring{4 RELATED WORKS }{4 RELATED WORKS }}\label{related-works}}
 
@@
 
-\subsection{5 METHODOLOGY}\label{methodology}
+\hypertarget{methodology}{%
+\subsection{5 METHODOLOGY}\label{methodology}}
 
@@
 
-\subsubsection{5.1 Iterative approach}\label{iterative-approach}
+\hypertarget{iterative-approach}{%
+\subsubsection{5.1 Iterative approach}\label{iterative-approach}}
 
@@
 
-\subsubsection{5.2 Evaluation}\label{evaluation}
+\hypertarget{evaluation}{%
+\subsubsection{5.2 Evaluation}\label{evaluation}}
 
@@
 
-\subsubsection{5.2 Limitations}\label{limitations}
+\hypertarget{limitations}{%
+\subsubsection{5.2 Limitations}\label{limitations}}
 
@@
 
-\subsection{6 EXPERIMENTATION}\label{experimentation}
+\hypertarget{experimentation}{%
+\subsection{6 EXPERIMENTATION}\label{experimentation}}
 
@@
 
+For our tests, we used six files with varying degrees of patterns in the bitstring, ranging from quite disastrous to acceptable.
+We constructed a simple test script to facilitate easier testing of each iteration.
+This script allowed us to easily test each implementation by automating several tasks -- generating a baseline binary file with data processed by a good, proven implementation of Toeplitz extraction, giving us a way to ensure each iteration still produces the correct result.
+Furthermore, this script allows us to define different sample sizes for the MCU to utilize, and can extract the amount of time the algorithm takes to execute (\emph{see table 1}) per each sample size.
+
+Below we'll detail the results for each iteration.
+
 \newpage
 
-\subsection{CHANGELOG}\label{changelog}
+\hypertarget{results}{%
+\subsection{\texorpdfstring{RESULTS }{RESULTS }}\label{results}}
+
+All iterations was tested on 6 different input datasets, with results variation less than 0.1 microseconds.
+Thus, they wont be included in tabells unless result is unexpected.
+All the tables showen below has been tested on ``biased\_bits\_bad''.
+
+Due to compatebility issues, pico pi 2 was tested on different machine, creating some minor variation in result -- about 1 microsecond.
+
+\hypertarget{iteration-1}{%
+\subsubsection{Iteration 1}\label{iteration-1}}
+
+\[
+\centering
+\begin{tabular}{|c|c|c|c|c|c|c|}
+\hline
+\textbf{Bit size} & \multicolumn{3}{c|}{\textbf{Teensy (µs)}} & \multicolumn{3}{c|}{\textbf{Pico (µs)}} \\
+\hline
+& Min & Max & Avg & Min & Max & Avg \\
+\hline
+64   & 13 & 14 & 13.1564 & 103 & 117 & 106.3914 \\
+512  & 788 & 791 & 788.3139 & 5296 & 5311 & 5302.4979 \\
+1024 & 3124 & 3130 & 3124.0580 & 21108 & 21119 & 21111.2163 \\
+\hline
+\end{tabular}
+\caption{Minimum, maximum and average processing time for Teensy and Pico.}
+\label{tab:teensy_pico_min_max_avg}
+\end{table}
+\]
+
+\hypertarget{changelog}{%
+\subsection{CHANGELOG}\label{changelog}}
 
@@
 2025-04-22: Begun including details regarding initial experimentation, updating details regarding experiments that had to change (e.g.~no baseline on separate hardware).
+
+2025-04-27: Added first table with test data.

### 2025-04-27 docs/style/article-final.tex
minor correction
first tabel, updated teensy makefile

@@
 
-\section{1 INTRODUCTION}\label{introduction}
+\hypertarget{introduction}{%
+\section{1 INTRODUCTION}\label{introduction}}
 
@@
 2025-04-22: Begun including details regarding initial experimentation, updating details regarding experiments that had to change (e.g.~no baseline on separate hardware).
+
+2025-04-27: Added first table with test data.
 % BALANCE COLUMNS

### 2025-04-28 docs/style/article-body.tex
updates to article

--- 2025-04-23 docs/style/article-body.tex
@@
 
+For our tests, we used six files with varying degrees of patterns in the bitstring, ranging from quite disastrous to acceptable.
+We constructed a simple test script to facilitate easier testing of each iteration.
+This script allowed us to easily test each implementation by automating several tasks -- generating a baseline binary file with data processed by a good, proven implementation of Toeplitz extraction, giving us a way to ensure each iteration still produces the correct result.
+Furthermore, this script allows us to define different sample sizes for the MCU to utilize, and can extract the amount of time the algorithm takes to execute (\emph{see table 1}) per each sample size.
+
+Below we'll detail the results for each iteration.
+
 \newpage
+
+\subsection{\texorpdfstring{RESULTS }{RESULTS }}\label{results}
+
+All iterations was tested on 6 different input datasets, with some results varying less than 0.1 microseconds.
+The results with such slight variations have been omitted from the table.
+All the tables shown below has been tested on the dataset `biased\_bits\_bad' which can be found as an appendix.
+
+Due to compatibility issues, Pi Pico 2 was tested on different machine, creating further variation in the results -- about 1 microsecond.
+
+\subsubsection{Iteration 1}\label{iteration-1}
+
+The first iteration brought few surprises, as it ran quite well on Teensy. To our surprise, it executed well enough on Pi Pico 2 as well, albeit with \(~8\times\) slower execution speed.
+
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1000}}
+  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1583}}
+  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1583}}
+  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1583}}
+  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1417}}
+  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1417}}
+  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1417}}@{}}
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Bit size}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Teensy Min (µs)}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Teensy Max (µs)}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Teensy Avg (µs)}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Pico Min (µs)}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Pico Max (µs)}
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+\textbf{Pico Avg (µs)}
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+64 & 13 & 14 & 13.1564 & 103 & 117 & 106.3914 \\
+512 & 788 & 791 & 788.3139 & 5296 & 5311 & 5302.4979 \\
+1024 & 3124 & 3130 & 3124.0580 & 21108 & 21119 & 21111.2163 \\
+\end{longtable}
+
+\subsubsection{\texorpdfstring{Iteration 2 }{Iteration 2 }}\label{iteration-2}
+
+Although the updated version included enhanced data structures, theoretical and practical analysis indicated that these modifications do not provide advantages for this project.
+Due to the small size of the input data (64, 512, 1024 bits) - using plain, contiguous memory layouts such as \texttt{std::vector\textless{}int\textgreater{}} proves to be very efficient in terms of access speed and memory usage.
+Current C++ compilers can effectively optimize these simple structures, often generating faster and more compact machine code than manually optimised alternatives.
+Introducing additional data structures or indirection would only add overhead without improving performance.
+
+From an algorithmic perspective, Toeplitz extraction inherently requires a double summation loop, which leads to a time complexity of \(O(n^2)\) no matter the optimization applied internally.
+Thus, making attempts to change it would not result in lower asymptotic cost.
+
+In summary, given the size and type of this system, a straightforward and minimal approach is not just adequate, but also recommended.
+
+\subsubsection{Iteration 3}\label{iteration-3}
+
+Coming soon!
 

### 2025-04-30 docs/style/article-body.tex
pushing latest version

@@
 All iterations was tested on 6 different input datasets, with some results varying less than 0.1 microseconds.
-The results with such slight variations have been omitted from the table.
-All the tables shown below has been tested on the dataset `biased\_bits\_bad' which can be found as an appendix.
+These results with such slight variations have been omitted from the table.
+All the tables shown below has been tested on the dataset \texttt{biased\_bits\_bad.bin} which can be found as an appendix -- and whereas more files have been used, this file has been arbitrarily selected to present the data.
 
@@
 
-The first iteration brought few surprises, as it ran quite well on Teensy. To our surprise, it executed well enough on Pi Pico 2 as well, albeit with \(~8\times\) slower execution speed.
+The first iteration brought few surprises, as it ran quite well on Teensy.
+To our surprise, it executed well enough on Pi Pico 2 as well, albeit with \textasciitilde{}\(8\times\) slower execution speed.
+This first implementation is the same one used to generate the baseline results, and used to ensure following iterations all generate correct output.
 
-\begin{longtable}[]{@{}
-  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1000}}
-  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1583}}
-  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1583}}
-  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1583}}
-  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1417}}
-  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1417}}
-  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1417}}@{}}
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Bit size}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Teensy Min (µs)}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Teensy Max (µs)}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Teensy Avg (µs)}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Pico Min (µs)}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Pico Max (µs)}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
-\textbf{Pico Avg (µs)}
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-64 & 13 & 14 & 13.1564 & 103 & 117 & 106.3914 \\
-512 & 788 & 791 & 788.3139 & 5296 & 5311 & 5302.4979 \\
-1024 & 3124 & 3130 & 3124.0580 & 21108 & 21119 & 21111.2163 \\
-\end{longtable}
+Worth noting is that this iteration (\_and all following iterations that generate the same result) have achieved entropy scores from \texttt{ent} that are greater than 7.99 for all tested data while ensuring it cannot be compressed further.
+This ensures that the algorithm has successfully removed as much of the patterns as can be reasonably expected from the raw binary data while maintaining the random nature of the data.
+
+\centering
+
+\begin{tabular}{|c|c|c|c|c|c|c|} \hline \textbf{Bit size} &
+\multicolumn{3}{c|}{\textbf{Teensy (µs)}} & \multicolumn{3}{c|}{\textbf{Pico
+(µs)}} \\ \hline & Min & Max & Avg & Min & Max & Avg \\ \hline 64 & 13 & 14 &
+13.1564 & 103 & 117 & 106.3914 \\ 512 & 788 & 791 & 788.3139 & 5296 & 5311 &
+5302.4979 \\ 1024 & 3124 & 3130 & 3124.0580 & 21108 & 21119 & 21111.2163 \\
+\hline \end{tabular} \caption{Minimum, maximum and average processing time for
+Teensy and Pico.} \label{tab:teensy_pico_min_max_avg}
+
+\textbackslash end\{table\}
 

### 2025-05-07 docs/style/article-body.tex
some restructure after mid-seminar

@@
 Another proposed solution for this is quantum random number generation (\emph{QRNG}) \cite{QRNG}.
-By reading quantum fluctuations from any given source, for instance an optical signal, the inherent natural unpredictability of said source can be harnessed in order to produce a random number from a state that is nigh impossible to reproduce accurately.
+By reading quantum fluctuation signals from any given source, for instance an optical signal, the inherent natural unpredictability of said signal can be harnessed in order to produce a random number from a state that is nigh impossible to reproduce accurately.
 
-In this project, we will be writing firmware for a quantum number generator, which generates true randomness by reading shot noise from an optical signal.
-Further details about how this signal is produced will be introduced in section 2 and 3, and builds on the work of Clason \cite{Clason2023}.
-This optical signal will be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}).
-In turn, these random bits will be processed via Toeplitz extraction \cite{toeplitz} in order to process these bits into random numbers.
-Some processing has to be done on the microcontrollers themselves in order to ensure that the data is workable, and Toeplitz extraction is a tried and tested method to accomplish this.
+Further details about how this signal is produced and sampled will be introduced in section 2 and 3, and builds on the work of Clason \cite{Clason2023}.
+This optical signal needs to be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}).
+Some processing of the data has to be done in order to ensure that the data is workable.
+As such, these random bits will be processed via Toeplitz extraction \cite{toeplitz} in order to process these bits into random numbers.
 These random numbers will then be output from the microcontroller to the host computer via USB.
@@
 
-\textbf{Research area 2 (\emph{RA2})}: How can we ensure that the output of random numbers is not limited by our firmware, but rather only limited by the USB transfer speed of the microcontroller or the ADC?
+\textbf{Research area 2 (\emph{RA2})}: Can we ensure that the output of random numbers is not limited primarily by firmware, but rather limited by the processing power or the USB transfer speed of the microcontroller, alternatively by the ADC?
 
@@
 Should these internal ADCs prove too limiting, we propose utilizing MAX11102AUB\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} with an effective sample rate of 2 million samples per second (MSPS).
-This ADC provides a 12 bit sample size, providing roughly 2.86 MB/s of sampled data per second, derived by the following calculation.
+This ADC provides a 12 bit sample size, providing roughly 24 Mbit/s of sampled data per second, derived by the following calculation.
 
 \[
-Data Rate = \frac{2,000,000_{MSPS} \times 12}{8 \times 1,048,576} \approx 2.86 MB/s \text{\phantom{123} (1)}
+\text{ADC Throughput}
+= \frac{2,000,000_{\mathrm{MSPS}}\times12}{1,000,000}
+\approx24\ \mathrm{Mbit/s}\phantom{123}(1)
 \]
@@
 
-\[
-T =
-\begin{bmatrix}
-1 & 0 & 1 & 1 & 0 & 1 & 0 & 1 \\
-0 & 1 & 1 & 0 & 1 & 0 & 1 & 1 \\
-1 & 1 & 0 & 1 & 0 & 1 & 1 & 0 \\
-1 & 0 & 1 & 1 & 1 & 0 & 1 & 0
-\end{bmatrix}
-\quad
-k =
-\begin{bmatrix}
-1 \\ 0 \\ 1 \\ 1 \\ 0 \\ 1 \\ 0 \\ 1
-\end{bmatrix}
-\]
+\begin{verbatim}
+Where n = number of input bits,
+      m = number of output bits:
 
-\[
-h(k) = T \cdot x \mod 2
-\]
+Let X be the input array (length n)
+Let T be the seed array (length n + m – 1)
+Let Y be a new array of length m
 
-\phantom{TEMP LINEBREAK}
+For i from 0 to m - 1:
+  sum = 0
+  For j from 0 to n - 1:
+    sum = sum + X[j] * T[i + j]
+  Y[i] = sum mod 2
 
-We then perform XOR multiplication on the matrices:
-
-\[
-h(k) =
-\begin{bmatrix}
-(1 \oplus 0 \oplus 1 \oplus 1 \oplus 0 \oplus 1 \oplus 0 \oplus 1) \\
-(0 \oplus 1 \oplus 1 \oplus 0 \oplus 1 \oplus 0 \oplus 1 \oplus 1) \\
-(1 \oplus 1 \oplus 0 \oplus 1 \oplus 0 \oplus 1 \oplus 1 \oplus 0) \\
-(1 \oplus 0 \oplus 1 \oplus 1 \oplus 1 \oplus 0 \oplus 1 \oplus 0)
-\end{bmatrix}
-\]
-
-Which finally simplifies to:
-
-\[
-h(k) =
-\begin{bmatrix}
-1 \\
-0 \\
-1 \\
-1
-\end{bmatrix}
-\]
-
-This allows us to interpret the result as a random 4-bit integer:
-
-\[
-h(k) = (1011)_2 = 11_{10}
-\]
+Return Y
+\end{verbatim}
 
@@
 With the assumption that the OQRNG-device produces a truly random analog signal, we can now clearly define the scope in which this thesis operates.
-Considering the maximum conversion speed from the ADC and the USB-output from the MCU, we have a clear bound between 2.86 MS/s (for the ADC) and 480 MB/s (for the USB output) in which Toeplitz extraction needs to be executed.
+Considering the maximum conversion speed from the ADC and the USB-output from the MCU, we have a clear bound over 24 Mbit/s (imposed by the ADC) in which Toeplitz extraction needs to be executed.
 Any speeds over 2.86 MB/s allows us to upgrade the ADC iteratively to continue increasing the output speed.
@@
 Code to measure the execution speed in microseconds will be implemented and tested before we load the naive implementation on said microcontroller.
-We expect that several implementations may be too resource intensive or have a memory complexity far greater than our cheaper, less capable microcontroller are able to handle, and as such these may not be able to be tested until a few iterations of optimization has occurred.
-
 Each iteration will consist of incremental improvements to the algorithm.
@@
 
-In order to evaluate the efficacy of each iteration, we will test the code by processing our pre-determined bitstreams, saving the processed bits in a binary file and evaluating this file with the command line utility \texttt{ent}.
-This utility runs a series of tests on the data to determine its level of entropy the file has -- rather, how information dense the file is.
-A perfect result (\emph{e.g.
-entirely random}) has an entropy value of \(8.0\), whereas acceptable entropy values for our tests are greater than \(7.99\), denoting a high level of randomness.
-Furthermore, \texttt{ent} can verify the arithmetic mean of the file, summing all the bits in the file and dividing it by the length of the file.
-A good distribution of randomness should have an arithmetic mean of \(~127.5\).
-To ensure each iteration works as expected, we will compare the \texttt{ent} result from each iteration against the original baseline, ensuring that the resulting processed bitstring is identical between iterations.
-There are a few more measurements shown by \texttt{ent}, and more details regarding these can be found in the utilities manual page\footnote{\href{https://manpages.ubuntu.com/manpages/trusty/man1/ent.1.html}{Manual page for \texttt{ent}, accessed 2025-04-23.}}.
+In order to evaluate the efficacy of each iteration, we will test the code by processing our pre-determined bitstreams, saving the processed bits in a binary file and evaluating this file with the command line utility \texttt{ent}\footnote{\href{https://manpages.ubuntu.com/manpages/trusty/man1/ent.1.html}{Manual page for \texttt{ent}, accessed 2025-04-23.}}, which is a simple way of evaluating the randomness of binary data.
 
-As discussed in section 3, the hardware used will impose a clear bound on how quickly our implementation needs to process the bits in order to match the speed of the ADC, as well as the output speed of the USB-port, both in \(MB/s\).
+As discussed in section 3, the hardware used will impose a clear bound on how quickly our implementation needs to process the bits in order to match the speed of the ADC and the output speed of the USB-port in \(Mbit/s\).
 In order to verify accuracy, the output of random numbers should be identical for each sample bitstring tested.
@@
 An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (as the process for this may vary greatly between controllers).
-As we will use fixed-size bitstrings for evaluation, we can then derive the throughput of the algorithm in \(MB/s\) as follows:
+As we will use fixed-size bitstrings for evaluation, we can then derive the throughput of the algorithm in \(Mbit/s\) as follows:
 
 \[
-Throughput*{MB/s} = \frac{DataSize*{bits}}{Execution Time_{ms}} \times
-\frac{1}{8} \times \frac{1000}{10^6} \text{\phantom{12}(2)}
+\text{Throughput}{\mathrm{Mbit/s}}
+= \frac{DataSize{\mathrm{bits}}}{ExecutionTime_{\mathrm{ms}}}
+\times 10^{-3}
+\phantom{12}(2)
 \]
@@
 This measurement allows us to place the throughput of our algorithm soundly in the bounds imposed on us by the hardware.
-During implementation, measurements of time complexity may be discussed with regards to the code -- but in the end, the performance of the code as it runs is what will be evaluated most thoroughly.
+Plugging in the \(24 Mbit/s\) bound imposed by the ADC with an arbitrarily chosen 64-bit sample size, we can derive the average execution speed in microseconds:
 
-\subsubsection{5.2 Limitations}\label{limitations}
+\[
+T_{\mathrm{ms}}
+= \frac{64}{24}\times10^{-3}\ \mathrm{ms}
+\approx 2.667\times10^{-3}\ \mathrm{ms}
+=2.667\ \mu\mathrm{s}.
+\]
+
+\subsubsection{5.3 Limitations}\label{limitations}
 
@@
 
-Worth noting is that this iteration (\_and all following iterations that generate the same result) have achieved entropy scores from \texttt{ent} that are greater than 7.99 for all tested data while ensuring it cannot be compressed further.
+Worth noting is that this iteration (\emph{and all following iterations that generate the same result}) have achieved entropy scores from \texttt{ent} that are greater than 7.99 for all tested data while ensuring it cannot be compressed further.
 This ensures that the algorithm has successfully removed as much of the patterns as can be reasonably expected from the raw binary data while maintaining the random nature of the data.
 
-\centering
-
 \begin{tabular}{|c|c|c|c|c|c|c|} \hline \textbf{Bit size} &
-\multicolumn{3}{c|}{\textbf{Teensy (µs)}} & \multicolumn{3}{c|}{\textbf{Pico
-(µs)}} \\ \hline & Min & Max & Avg & Min & Max & Avg \\ \hline 64 & 13 & 14 &
-13.1564 & 103 & 117 & 106.3914 \\ 512 & 788 & 791 & 788.3139 & 5296 & 5311 &
-5302.4979 \\ 1024 & 3124 & 3130 & 3124.0580 & 21108 & 21119 & 21111.2163 \\
-\hline \end{tabular} \caption{Minimum, maximum and average processing time for
-Teensy and Pico.} \label{tab:teensy_pico_min_max_avg}
-
-\textbackslash end\{table\}
+\multicolumn{3}{c|}{\textbf{Teensy ($\mu s$)}} &
+\multicolumn{3}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline & Min & Max & Avg & Min &
+Max & Avg \\ \hline 64 & 13 & 14 & 13.1564 & 103 & 117 & 106.3914 \\ 512 & 788 &
+791 & 788.3139 & 5296 & 5311 & 5302.4979 \\ 1024 & 3124 & 3130 & 3124.0580 &
+21108 & 21119 & 21111.2163 \\ \hline \end{tabular} \caption{Minimum, maximum and
+average processing time for Teensy and Pico.}
+\label{tab:teensy_pico_min_max_avg}
 

### 2025-05-07 project/tests/tests.py
allot of tests, new data & iteration

@@ -11,7 +11,7 @@ import subprocess
 from pathlib import Path
 
 SERIAL_BAUD = 6000000
-BIT_CHUNK_SIZE = 1024
+BIT_CHUNK_SIZE = 128
 OUTPUT_LEN = BIT_CHUNK_SIZE // 2
 
 BIT_CHUNK_BYTES = BIT_CHUNK_SIZE // 8
@@ -164,8 +164,8 @@ def generate_results(port: str) -> None:
 
 def log_times(port: str) -> None:
     while True:
-        current_iter = input('⌨️ Select iteration (1 - 5): ')
-        if current_iter in ['1', '2', '3', '4', '5']:
+        current_iter = input('⌨️ Select iteration (1 - 6): ')
+        if current_iter in ['1', '2', '3', '4', '5', '6']:
             break
         print('❌ Not a valid iteration.')
 

### 2025-05-15 docs/style/article-body.tex
edited method

@@
 This optical signal needs to be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}).
-Some processing of the data has to be done in order to ensure that the data is workable.
-As such, these random bits will be processed via Toeplitz extraction \cite{toeplitz} in order to process these bits into random numbers.
-These random numbers will then be output from the microcontroller to the host computer via USB.
-This thesis will aim to answer one key research question: How can sampled vacuum fluctuations be processed efficiently in order to output QRNG?
+Some processing of the data has to be done in order to ensure that the data is workable and to remove potential deterministic patterns from these random bits.
+One method for this post-processing we will explore in this work is Toeplitz extraction \cite{toeplitz}, typically performed on the host computer utilizing the randomly generated numbers.
+While this system could very well be implemented on physically larger hardware, thus avoiding the constraints that limited hardware introduces, Clason proposes a simpler and cheaper way to achieve QRNG \cite{Clason2023}.
+In keeping with this, utilizing microcontrollers rather than a host computer for processing the raw bits extracted further helps to keep the costs low and the solution reasonably complex.
+Additionally, this allows the system to be self-contained, for instance a simple and portable USB-device which both generates and processes true random numbers.
+Further details regarding the microcontrollers used in our work will be outlined in section 3.
 
-In producing this firmware, several key considerations have to be made in order for this system to be usable in a production environment.
-The vision for the end product is a simple USB-stick that can be connected to a host, and produce true random numbers from ambient quantum fluctuations.
-While this system could very well be implemented on physically larger hardware, thus avoiding the constraints that limited hardware introduces, Clason proposes a simpler and cheaper way to achieve QRNG \cite{Clason2023}.
-In keeping with this, utilizing microcontrollers for processing the raw bits extracted further helps to keep the costs low and the solution reasonably complex.
-Further details regarding the microcontrollers used in our work will be outlined in section 3.
-However, due to this portability constraint, our implementation needs to work quickly and efficiently on resource constrained hardware.
-As such, our main question is broken down into two concrete research areas:
+However, due to the lower processing power of the average microcontroller, any implementation of Toeplitz extraction needs to work quickly and efficiently despite this hardware constraint.
+Experimenting with efficient implementations of this well-defined algorithm is the main focus of this thesis, and the key research areas explored in this work are as follows:
 
@@
 
-\textbf{Research area 2 (\emph{RA2})}: Can we ensure that the output of random numbers is not limited primarily by firmware, but rather limited by the processing power or the USB transfer speed of the microcontroller, alternatively by the ADC?
+\textbf{Research area 2 (\emph{RA2})}: Can we ensure that the output of random numbers is not primarily limited primarily by our implementation, but rather limited by the processing power or the USB transfer speed of the microcontroller, alternatively by the ADC?
 
@@
 Section 3 delves further into the hardware and algorithms our work will use, with related works in optimizing Toeplitz extraction listed under section 4.
-Section 5 will present our methodology as well as some limitations imposed on our work.
-\textbf{More sections to follow as we finish the article}.
+Section 5 will present our methodology and implementation strategy, as well as some limitations imposed on our work.
+Finally, section 7 will present the results of our experimentation.
 
@@
 
-Microcontrollers (MCUs) are compact and low-power computing devices designed primary for embedded systems and real-time applications.
-Unlike general purpose CPUs, MCUs integrates a processor, memory, and peripherals into a single chip.
-Thus, enabling efficient and autonomous operation in constrained environments.
-Various development frameworks and tools are available for programming MCUs, including the Arduino Framework, manufacturer specific SDKs and raw C++, among others.
-Each framework presents different trade-offs in terms of performance overhead, efficiency, and ease of development.
-
-Modern microcontrollers feature advanced microarchitectural elements to enhance processing capabilities.
-Some of high performance MCUs utilize dual-issue superscalar architectures, enabling them to run several instructions in one cycle.
-This type of design greatly improves single-threaded capabilities, which makes these MCUs ideal for tasks that need real-time data processing.
-The use of dual-issue architectures in embedded devices has demonstrated high execution speeds, along with the added benefit of improved energy efficiency.
-Thus, making them a great option for applications that require low power but high performance.
-Research conducted by Lozano et al.
-\cite{dual-issue} specifically explores energy efficient dual-issue processors designs.
-This study shows that optimizing instruction throughput and power consumption can significantly improve processing efficiency in embedded systems.
-While energy efficiency is not the primary goal for this study, ensuring that the resulting USB-device does not consume too much power is tangentially relevant for real-world usage.
+Microcontrollers (MCUs) are compact and low-power computing devices designed for embedded systems and real-time operations, and suitable as a processing unit for the purposes of this work.
+Unlike general purpose CPUs, an MCU integrates a processor, memory and peripherals (\emph{such as an ADC, for instance}) into a single chip.
+Furthermore, modern MCUs often feature advanced microarchitectural elements to enhance processing capabilities on single threads (\emph{such as dual-issue superscalar architectures, allowing the MCU to run several instructions per CPU cycle}), making them suitable candidates for the post-processing required for OQRNG-data.
 
@@
 
-The Teensy 4.1\footnote{\href{https://www.pjrc.com/store/teensy41.html}{Teensy developer documentation, accessed 2025-02-27}}, based on the ARM Cortex-M7 is especially good for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and DSP capabilities.
-The floating-point unit\footnote{\href{https://www.sciencedirect.com/topics/computer-science/floating-point-unit}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (FPU) and SIMD-style DSP instructions improves how quickly it can perform bitwise and arithmetic tasks, which are crucial for quick Toeplitz extraction.
+In our work, we intend to use Teensy 4.1\footnote{\href{https://www.pjrc.com/store/teensy41.html}{Teensy developer documentation, accessed 2025-02-27}}, based on the ARM Cortex-M7.
+This MCU is especially suitable for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and Digital Signal Processing (\emph{DSP}) capabilities.
+The floating-point unit\footnote{\href{https://www.sciencedirect.com/topics/computer-science/floating-point-unit}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (FPU) and Single Instruction, Multiple Data (\emph{SIMD}) style DSP instructions improves how quickly it can perform bitwise and arithmetic tasks, which are crucial for quick Toeplitz extraction.
 SIMD-controlled DSP architectures, as described by Han et al.
-\cite{simd-dsp}, leverage parallel vectorized computation to accelerate matrix operations.
-Thus, making them highly effective for Toeplitz matrix-vector multiplications.
-The Cortex-M7 does not have special features for cryptography, its ability to execute multiple instructions quickly, combined with fast memory access and an efficient pipeline design, allows it to effectively manage and process randomness extraction.
-Moreover, its USB High-Speed (480 Mbit/sec) interface, supporting both USB device and USB host modes, allows rapid transmission of extracted random data to an external system, ensuring minimal bottlenecks in high-rate randomness generation.
+\cite{simd-dsp}, leverage parallel vectorized computation to accelerate matrix operations -- making them highly effective for Toeplitz matrix-vector multiplications.
 
-In order to evaluate just how much constraint our implementation can allow, our aim is to try our implementation on other MCUs with varying levels of power and hardware support.
+In order to evaluate how efficent our implementation can become, our aim is to try our implementation on other MCUs with varying levels of power and hardware support.
 Whereas Teensy 4.1 is our primary development platform which we will evaluate closely, we aim to run our implementations on Raspberry Pi Pico 2\footnote{\href{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}{Raspberry Pi Pico 2 documentation, accessed 2025-03-13}} as well as ESP32-S3\footnote{\href{https://www.espressif.com/sites/default/files/documentation/esp32-s3_datasheet_en.pdf}{ESP32-S3 documentation, accessed 2025-03-13}}.
 Due to the lower computational power of these MCUs, there may be significant issues in utilizing these weaker models, yet they are significantly cheaper and easier to access.
-Testing of these will consist solely of running the implementation on these controllers and measuring execution speed.
+Testing of these will consist solely of running the implementation on these controllers and measuring execution speed and correctness of the output.
 
-\subsubsection{\texorpdfstring{3.4 Toeplitz extraction }{3.4 Toeplitz extraction }}\label{toeplitz-extraction}
+\subsubsection{3.4 Toeplitz extraction}\label{toeplitz-extraction}
 
@@
 Let X be the input array (length n)
-Let T be the seed array (length n + m – 1)
+Let T be the seed matrix (length n + m – 1)
 Let Y be a new array of length m
@@
 
-Whereas this example is quite small in order to demonstrate how the algorithm operates, this can of course be scaled up significantly with larger values for both \(T\) and \(k\) respectively, granting larger integers.
 The main focus of this work is implementing this algorithm as efficiently as possible on our MCUs, and as such, several optimization efforts need to be taken into account during our experimentation.
@@
 
-\subsection{\texorpdfstring{4 RELATED WORKS }{4 RELATED WORKS }}\label{related-works}
+\subsection{4 RELATED WORKS}\label{related-works}
 
@@
 The main point of the algorithm is to remove patterns in the bitstream that may lead to less randomized results.
-As such, we will manually produce several bitstreams to use for our tests -- each with varying degrees of repeated patterns that should be eliminated by the algorithm.
-Furthermore, in order to see how well our implementation will work with realistic data from the OQRNG generator, we will sample streams of bits directly from the source.
+As such, we will sample several bitstreams from the OQRNG-device to use for our tests -- each with varying degrees of repeated patterns that should be eliminated by the algorithm.
+All bitstrings tested are available as an appendix to this paper.
 
-\subsubsection{5.1 Iterative approach}\label{iterative-approach}
+\subsubsection{5.1 Implementation strategy}\label{implementation-strategy}
 
-Implementation of Toeplitz extraction will begin by a naive implementation not optimized for speed, but rather for accuracy.
-This implementation will be executed on a separate computer in order to produce the correct random number for every provided bitstring.
-These will be used as our baseline for accuracy for future iterations.
+Our implementation of the Toeplitz extractor followed a structured, iterative approach divided into two phases.
+Phase one focused on exploring performance improvements through incremental algorithmic changes.
+Phase two then addressed architectural inefficiencies identified during phase one.
+Each phase then consisted of a number of individual iterations, in which performance was evaluated in terms of both correctness and execution time on the target hardware.
 
-This naive implementation will then be flashed to our microcontrollers, beginning with Teensy 4.1 as this is the more capable of the microcontrollers used for this experiment.
-Code to measure the execution speed in microseconds will be implemented and tested before we load the naive implementation on said microcontroller.
-Each iteration will consist of incremental improvements to the algorithm.
-Our initial investigation has shown several avenues for improving the throughput of the algorithm, and each planned iteration will be discussed briefly.
-Depending on the empirical results of these iterations, there may be a need for further optimization iterations other than those listed.
+\paragraph{5.2 Phase one}\label{phase-one}
 
-\textbf{Iteration 1 - Naive implementation}: The naive implementation consists of a Toeplitz matrix \(T\) and key \(k\), acquired from fixed slices of the bitstream.
-Then, matrix-vector multiplication will be performed using \(T\) and \(k\), thus performing Toeplitz extraction -- which finally will eliminate jitter and produce a random number.
-The implementation of this in code will be quite simple, using nested loops to iterate over all values and multiply them.
-We expect that the throughput of this naive implementation will be far away from optimized iterations, and result in \(O(n^2)\) complexity.
+Initially, we require a ``naive'' version designed to prioritize correctness over speed.
+This version was first executed on a separate computer to generate reference output for various input bitstrings, which were later used as accuracy baselines.
+The naive implementation was then flashed onto the Teensy 4.1 microcontroller, where execution time was measured in microseconds.
+Each subsequent iteration introduced controlled modifications aimed at improving throughput.
 
-\textbf{Iteration 2 - Efficient data structures}: The naive implementation will not utilize efficient data structures for maximum speed, rather it will likely use fixed-size arrays which need to be iterated over for a time complexity of \(O(n^2)\).
-Our hypothesis is that more efficient data structures may allow data to be processed in a far more efficient manner -- however, this scenario poses a risk for greater memory complexity which may not be suitable for use on microcontrollers.
-Which data structures might be most feasible will be evaluated during this iteration and as such is not preordained.
+The initial implementation followed the pseudocode described in section 3.4, using matrix multiplication over raw input and seed data.
+It relied on \texttt{std::vector\textless{}int\textgreater{}} for storage and used nested loops to compute each output bit.
+From this point, this implementation was improved over the coming iterations, and the new implementation verified in the same manner as the initial version.
 
-\textbf{Iteration 3 - Bitshifting}: We expect that Toeplitz extraction can be significantly improved with bitwise XOR operations, thus reducing overhead and memory usage.
-Instead of explicitly constructing the Toeplitz matrix, a right-shift operation can be used to dynamically reconstruct matrix rows.
-Additionally, matrix-vector multiplication can be improved using XOR operations to extract bits more efficient, thus minimizing unnecessary computations.
+\textbf{Iteration 1 - Data structures:} This iteration kept the same algorithmic logic as the initial implementation, but experimented with data structures such as raw pointers and hash maps to enhance the use of \texttt{std::vector}.
+The goal was to reduce or eliminate the reliance on nested loops.
 
-\textbf{Iteration 4 - Batching}: Finally, we consider the concept of batching larger amounts of bits for processing.
-Consider that an input buffer of bits is read from the ADC and stored, waiting for processing.
-Rather than taking 64 bits, and shifting them one by one, we can take two batches of 64 bits and multiply them directly using XOR bitshifting -- eliminating the need to process these bit-by-bit.
-As we will have a constant stream of bits from the ADC, we theorize that this method will allow for better performance in real-time processing over reading individual bits.
+\textbf{Iteration 2 - Bitshifting:} Basic bitwise operation was introduced to replace arithmetic whenever possible.
+Multiplication was replaced with logical \texttt{AND\ \&} and modulo operations with bit masking \texttt{\&\ 1}.
+The goal was to reduce the number of instructions and improve per-bit processing speed.
 
-\textbf{Iteration 5 - ARM Hardware instructions}: Rather than performing bitshifting operations in the code itself, certain microcontrollers come equipped with a separate processor specifically for bitshift-operations.
-Offloading the shifting to these processors rather than running them on the main CPU may allow faster processing of the data than performing the shifting in the code itself.
-However, this operation isn't natively supported by Arduino, and will potentially lead to extreme rewrites of the code which may prove too time consuming to do for two microcontrollers (as the code will not be reusable between controllers).
-As such, this optimization may only be done for one controller or left for future work.
+\textbf{Iteration 3 - Batching and Hardware optimization:} This iteration focused on optimizing performance through batching and the use of ARM-native instructions, beginning by testing batching alone.
+This was followed by isolated use of ARM instructions such as \texttt{\_\_builtin\_popcountll()} (\emph{which counts the number of set bits in an unsigned integer}).
+After establishing their individual effect we combined both techniques, multiple batch sizes were tested to determine their impact.
+More details and benchmarks for each configuration can be found in the section 6.
 
-Each scenario will first be executed on a single thread, and multiple threads may theoretically be executed concurrently to further optimize the data.
-The feasibility of this highly depends on the performance of every individual implementation.
-Depending on the results during our experimentation, this may yield yet another iteration.
+\paragraph{5.3 Phase two}\label{phase-two}
 
-\subsubsection{5.2 Evaluation}\label{evaluation}
+Phase two focused on addressing inefficiencies and design issues that were unintentionally introduced during earlier iterations.
+Rather than continuing with new algorithmic ideas, this phase aimed to identify and fix structural problems.
+Several assumptions from phase one were re-evaluated -- such as the benefits of certain data structures or abstractions.
 
-In order to evaluate the efficacy of each iteration, we will test the code by processing our pre-determined bitstreams, saving the processed bits in a binary file and evaluating this file with the command line utility \texttt{ent}\footnote{\href{https://manpages.ubuntu.com/manpages/trusty/man1/ent.1.html}{Manual page for \texttt{ent}, accessed 2025-04-23.}}, which is a simple way of evaluating the randomness of binary data.
+\textbf{Iteration 4 - Loop unrolling:} This iteration focused on reducing the number of loops in the extractor by manually unrolling repeated operations.
+The goal was to decrease overhead created by loops.
+Whereas this operation is commonly done by compiler optimization, manually performing this guarantees that we unroll the loops rather than leaving it to the compiler.
 
-As discussed in section 3, the hardware used will impose a clear bound on how quickly our implementation needs to process the bits in order to match the speed of the ADC and the output speed of the USB-port in \(Mbit/s\).
-In order to verify accuracy, the output of random numbers should be identical for each sample bitstring tested.
-Certain iterations might, however, use larger parts of the bitstring, and thus produce a slightly variable result.
+\textbf{Iteration 5 - Removal of vector usage:} This iteration removed \texttt{std::vector} in favor of fixed-size types like \texttt{uint32\_t} and \texttt{uint64\_t} to reduce the overhead introduced by creating and populating this complex data structure.
+
+\textbf{Iteration 6 - Data type exploration:} Following the removal of vectors, this iteration explored alternative static data types to determine the most efficient structure for storing input and seed data.
+
+\textbf{Iteration 7 - Re-implementation of bitshifting:} In this final iteration, bit shifting techniques were reintroduced into the now simplified implementation.
+With high-level abstractions removed and memory structures streamlined, bit-level operations such as shifting and masking were applied.
+
+\subsubsection{5.4 Evaluation}\label{evaluation}
+
+To evaluate the correctness of each implementation, a baseline was generated as discussed in section 5.2.
+Using the naive, initial implementation to process bits and saving for later evaluation gave us a source of truth against which to compare following iterations.
+To verify that the algorithm successfully removed the patterns it should, we verified the measured entropy score with the command line utility \texttt{ent}\footnote{\href{https://manpages.ubuntu.com/manpages/trusty/man1/ent.1.html}{Manual page for \texttt{ent}, accessed 2025-04-23.}}.
+
+We created a script to facilitate easier testing, which is attached to this paper.
+Using this script, all bitstrings used for testing can be evaluated against the baseline, ensuring that the output from the new iteration matches the baseline exactly.
+Furthermore, the script also provides the average execution time of only the Toeplitz extraction in microseconds.
 
@@
 =2.667\ \mu\mathrm{s}.
+\phantom{12}(3)
 \]
+
+In section 6, this calculation will be used to derive the execution speed of the various iterations.
 
@@
 
-\subsection{6 EXPERIMENTATION}\label{experimentation}
-
-Each iteration results in an implementation written in C++, all of which will be supplied as an appendix to this paper.
-Furthermore, to simplify the implementation process, we will use the Arduino framework to communicate with the hardware.
-Whereas this introduces some overhead, it drastically reduces the time spent per iteration, and is deemed to be a suitable starting point.
-
-Initially, we created a script to simulate a stream of bits from a file, and pre-computed the expected results using a trusted version of Toeplitz extraction.
-With the expected results saved to a file on the host computer, each iteration could then be deployed to the MCU, and tested with the same data as the baseline.
-As the algorithm is fully deterministic with this uniform data, the exact same results in the same order should be produced by all iterations.
-The simulation script then simply asserts that the MCU produces the same random string of bits from the resulting iterations as with the baseline.
-
-As our system does not rely on interrupts in any capacity, measuring the execution speed of the implemented Toeplitz extraction function can be done using the built in \texttt{micros()}-function provided by Arduino.
-This provides us with an accurate measurement of the execution time per function call.
-The average execution time per test set over each iteration will be presented in \textbf{\texttt{TABLE\ 1}}, along with the fastest and slowest executions per iteration respectively.
-Furthermore, the entropy value on the processed bitstream will be noted in the table to demonstrate that the important measurement, e.g.~the randomness, still holds true between iterations.
-
-For our tests, we used six files with varying degrees of patterns in the bitstring, ranging from quite disastrous to acceptable.
-We constructed a simple test script to facilitate easier testing of each iteration.
-This script allowed us to easily test each implementation by automating several tasks -- generating a baseline binary file with data processed by a good, proven implementation of Toeplitz extraction, giving us a way to ensure each iteration still produces the correct result.
-Furthermore, this script allows us to define different sample sizes for the MCU to utilize, and can extract the amount of time the algorithm takes to execute (\emph{see table 1}) per each sample size.
-
-Below we'll detail the results for each iteration.
-
-\newpage
-
-\subsection{\texorpdfstring{RESULTS }{RESULTS }}\label{results}
-
-All iterations was tested on 6 different input datasets, with some results varying less than 0.1 microseconds.
-These results with such slight variations have been omitted from the table.
-All the tables shown below has been tested on the dataset \texttt{biased\_bits\_bad.bin} which can be found as an appendix -- and whereas more files have been used, this file has been arbitrarily selected to present the data.
-
-Due to compatibility issues, Pi Pico 2 was tested on different machine, creating further variation in the results -- about 1 microsecond.
-
-\subsubsection{Iteration 1}\label{iteration-1}
-
-The first iteration brought few surprises, as it ran quite well on Teensy.
-To our surprise, it executed well enough on Pi Pico 2 as well, albeit with \textasciitilde{}\(8\times\) slower execution speed.
-This first implementation is the same one used to generate the baseline results, and used to ensure following iterations all generate correct output.
-
-Worth noting is that this iteration (\emph{and all following iterations that generate the same result}) have achieved entropy scores from \texttt{ent} that are greater than 7.99 for all tested data while ensuring it cannot be compressed further.
-This ensures that the algorithm has successfully removed as much of the patterns as can be reasonably expected from the raw binary data while maintaining the random nature of the data.
+\subsection{6 RESULTS}\label{results}
 
@@
 
-\subsubsection{\texorpdfstring{Iteration 2 }{Iteration 2 }}\label{iteration-2}
-
-Although the updated version included enhanced data structures, theoretical and practical analysis indicated that these modifications do not provide advantages for this project.
-Due to the small size of the input data (64, 512, 1024 bits) - using plain, contiguous memory layouts such as \texttt{std::vector\textless{}int\textgreater{}} proves to be very efficient in terms of access speed and memory usage.
-Current C++ compilers can effectively optimize these simple structures, often generating faster and more compact machine code than manually optimised alternatives.
-Introducing additional data structures or indirection would only add overhead without improving performance.
-
-From an algorithmic perspective, Toeplitz extraction inherently requires a double summation loop, which leads to a time complexity of \(O(n^2)\) no matter the optimization applied internally.
-Thus, making attempts to change it would not result in lower asymptotic cost.
-
-In summary, given the size and type of this system, a straightforward and minimal approach is not just adequate, but also recommended.
-
-\subsubsection{Iteration 3}\label{iteration-3}
-
-Coming soon!
+Table 1 presents the execution speeds of each iteration as listed in sections 5.3 and 5.4, alongside the exectution speed in Mbit/s as calculated by utilizing the average value in (3).
+My thought then is that we here talk about the results and what they demonstrate, before moving on to draw conclusions and attempt to answer our research questions in section 7.
 

### 2025-05-20 docs/style/article-body.tex
fixed most small comments from meeting

@@
 However, the process of producing these random numbers tends to be pseudo-random, e.g.~utilizing the current states of various modules \cite{randomness}.
-These numbers do not generate true randomness, and in order to heighten security other methods of generating these are required.
-Current random number generators (\emph{RNG}) are usually implemented in code, using certain states of the host machine as a starting point before running a predetermined algorithm \cite{randomness}.
+Pseudo-random numbers do not generate true randomness, and in order to heighten security, other methods of generating random numbers are required.
+Current random number generators (\emph{RNG}) are usually implemented in computer programs, using certain states of the host machine as a starting point before running a predetermined algorithm \cite{randomness}.
 This pseudo-random number generation (\emph{PRNG}) comes with the drawback that the result is always deterministic, provided that the initial state is known.
@@
 By reading quantum fluctuation signals from any given source, for instance an optical signal, the inherent natural unpredictability of said signal can be harnessed in order to produce a random number from a state that is nigh impossible to reproduce accurately.
+Clason \cite{Clason2023} presents a device that generates a fluctuating analogue signal utilizing this method, which we will use as a basis for our work.
 
-Further details about how this signal is produced and sampled will be introduced in section 2 and 3, and builds on the work of Clason \cite{Clason2023}.
-This optical signal needs to be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}).
-Some processing of the data has to be done in order to ensure that the data is workable and to remove potential deterministic patterns from these random bits.
+Further details about how this signal is produced and sampled will be introduced in section 2 and 3.
+The optical signal output by the device described in \cite{Clason2023} needs to be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}).
+Some post-processing of the said raw bits has to be performed in order to ensure that the bits are workable, and to remove potential deterministic patterns from the data.
 One method for this post-processing we will explore in this work is Toeplitz extraction \cite{toeplitz}, typically performed on the host computer utilizing the randomly generated numbers.
-While this system could very well be implemented on physically larger hardware, thus avoiding the constraints that limited hardware introduces, Clason proposes a simpler and cheaper way to achieve QRNG \cite{Clason2023}.
+
+\[
+    \text{OQRNG Analogue output} \rightarrow \text{ADC-converter} \rightarrow \text{post-processing}
+\]
+
+Clason proposes a simpler and cheaper way to achieve QRNG \cite{Clason2023}.
 In keeping with this, utilizing microcontrollers rather than a host computer for processing the raw bits extracted further helps to keep the costs low and the solution reasonably complex.
@@
 
-We created a script to facilitate easier testing, which is attached to this paper.
-Using this script, all bitstrings used for testing can be evaluated against the baseline, ensuring that the output from the new iteration matches the baseline exactly.
-Furthermore, the script also provides the average execution time of only the Toeplitz extraction in microseconds.
-
 Hyncica et.
@@
 
+We created a script to facilitate easier testing, which is attached to this paper.
+Using this script, all bitstrings used for testing can be evaluated against the baseline, ensuring that the output from the new iteration matches the baseline exactly.
+Furthermore, the script also provides the average execution time of only the Toeplitz extraction in microseconds.
+
 \begin{tabular}{|c|c|c|c|c|c|c|} \hline \textbf{Bit size} &
@@
 My thought then is that we here talk about the results and what they demonstrate, before moving on to draw conclusions and attempt to answer our research questions in section 7.
+
+\subsection{7 CONCLUSION}\label{conclusion}
 
@@
 2025-04-27: Added first table with test data.
+
+2025-05-20: Unfortunately, we've missed adding concrete details about changes made up until this point. Fix later.

### 2025-05-20 docs/style/article-final.tex
fixed most small comments from meeting

@@
 2025-04-27: Added first table with test data.
+
+2025-05-20: Unfortunately, we've missed adding concrete details about changes made up until this point. Fix later.
 % BALANCE COLUMNS

### 2025-05-21 docs/style/article-body.tex
pushing algorithm update

@@
 By reading quantum fluctuation signals from any given source, for instance an optical signal, the inherent natural unpredictability of said signal can be harnessed in order to produce a random number from a state that is nigh impossible to reproduce accurately.
-Clason \cite{Clason2023} presents a device that generates a fluctuating analogue signal utilizing this method, which we will use as a basis for our work.
+Clason \cite{Clason2023} presents a device that generates a fluctuating analogue signal utilizing this method.
 
-Further details about how this signal is produced and sampled will be introduced in section 2 and 3.
 The optical signal output by the device described in \cite{Clason2023} needs to be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}).
@@
 One method for this post-processing we will explore in this work is Toeplitz extraction \cite{toeplitz}, typically performed on the host computer utilizing the randomly generated numbers.
+The post-processing finally yields a random number.
+This relationship can be seen in Figure 1.
 
-\[
-    \text{OQRNG Analogue output} \rightarrow \text{ADC-converter} \rightarrow \text{post-processing}
-\]
+\begin{figure}[ht] \centering \begin{tikzpicture}[>=latex] % top row, 2 cm apart
+\node (A) {RNG-module}; \node (B) [right=2cm of A] {ADC-conversion};
 
-Clason proposes a simpler and cheaper way to achieve QRNG \cite{Clason2023}.
-In keeping with this, utilizing microcontrollers rather than a host computer for processing the raw bits extracted further helps to keep the costs low and the solution reasonably complex.
-Additionally, this allows the system to be self-contained, for instance a simple and portable USB-device which both generates and processes true random numbers.
-Further details regarding the microcontrollers used in our work will be outlined in section 3.
+    \node (C) [below=1cm of B] {Post-processing};
+    \node (D) [left=2cm of C]  {Random number};
 
-However, due to the lower processing power of the average microcontroller, any implementation of Toeplitz extraction needs to work quickly and efficiently despite this hardware constraint.
+    \draw[->] (A) -- (B);
+    \draw[->] (B) -- (C);
+    \draw[->] (C) -- (D);
+
+\end{tikzpicture} \caption{An abstract of the steps required for OQRNG.}
+\label{fig:linear-system} \end{figure}
+
+Clason proposes a simpler and cheaper way to achieve QRNG \cite{Clason2023} than what has been done in previous research.
+These systems have generally been large, bulky and expensive, whereas the method proposed is portable enough to theoretically be installed in a portable device -- for instance a USB-thumbstick.
+In keeping with this, utilizing microcontrollers rather than a host computer for processing the raw bits extracted allows this system to be self-contained and portable, and further helps to keep the costs low and the solution reasonably complex.
+
+However, due to the limited processing power of the average microcontroller, any implementation of Toeplitz extraction needs to work quickly and efficiently despite this hardware constraint.
 Experimenting with efficient implementations of this well-defined algorithm is the main focus of this thesis, and the key research areas explored in this work are as follows:
@@
 
-Toeplitz extraction been optimized quite well, and previous research can be utilized for this.
+Toeplitz extraction has been optimized quite well, and previous research can be utilized to address this research question.
 However, there are still considerations when implementing the firmware for the microcontroller in order to optimize the code.
@@
 For instance, the speed at which the ADC can process the optical signal into raw bits as well as the speed that the USB output can transfer processed random number to the host computer will be limiting factors.
-Further details on the limitations of the ADC will be outlined in section 3.
+Further details on the limitations of the ADC will be outlined in Section 3.
 The slowest of these bottlenecks will inevitably be the limiting factor for any implementation.
@@
 The idea of an optical QRNG (\emph{OQRNG}) is not a novel one.
-The basis of the theory is that intrinsically random properties of a quantum process.
+The basis of the theory is the intrinsically random properties of a quantum process.
 Stefanov et.
@@
 This particular implementation uses a bespoke circuit board where all components are present on a single board -- e.g., this experimental setup contains an integrated ADC, post-processor, entropy controller and entropy generator.
-While this article cements the viability of OQRNG using shot noise (despite the article not being confirmed as peer reviewed), the bespoke nature of the circuit board makes this experiment difficult to reproduce.
+While this article cements the viability of OQRNG using shot noise (\emph{despite the article not being confirmed as peer reviewed}), the bespoke nature of the circuit board makes this experiment difficult to reproduce.
 As our thesis will use commercially available ADCs and microcontrollers, the only bespoke component is the shot noise generator itself.
@@
 Furthermore, to the best of our knowledge, most of the work in this field is from the perspective of physicists, and there appears to be little research on this subject in the domain of computer science.
-Our work aims to bridge this gap by using commercially available hardware (other than the bespoke shot noise generator \cite{Clason2023}) and focuses on implementing Toeplitz extraction directly on the microcontroller.
+Our work aims to bridge this gap by using commercially available hardware (\emph{other than the bespoke shot noise generator \cite{Clason2023}}) and focuses on implementing Toeplitz extraction directly on the microcontroller.
 Rather than focusing on the intricacies of quantum fluctuations, we will instead approach this problem from a computer science perspective.
@@
 Our work is a practical continuation of the work of Clason \cite{Clason2023}.
-In this work, quantum shot noise originating from photodiodes was studies, and in so doing a prototype device was constructed.
-This device read from an optical source, outputting analog voltage from the data ``seen'' by the diodes.
-A prototype was constructed, in which an LED is read by a photodiode soldered millimeters apart.
-In this section, we introduce the hardware used for our implementations as well as the considerations taken in order to shift the focus from physics to computer science.
+In this work, quantum shot noise originating from photodiodes was studied, and in so doing a prototype device was constructed.
+This prototype utilized an LED soldered millimeters apart from a photodiode, which in turn outputs a variable analog signal from the device.
+In this section, we introduce the remaining hardware used for our implementations as well as the considerations taken in order to shift the focus from physics to computer science.
 
@@
 
-Whereas the exact quantum mechanisms that ensure that this system ensures randomness and further details regarding the OQRNG-device is better derived directly from Clasons work \cite{Clason2023}, the end result as it correlates to our study is an inherently random, analog voltage current.
+Whereas the exact quantum mechanisms that ensure that this system ensures randomness and further details regarding the OQRNG-device is better derived directly from Clasons work \cite{Clason2023}, the primary concern for our study is the inherently random, analog voltage current produced by the system.
 
@@
 
-This analog current isn't suitable to operate on without further processing.
-As mentioned in section 1, the signal needs to pass through an ADC and converted into raw bits in order for it to be usable.
+This analog current is not suitable to operate on without further processing.
+As mentioned in Section 1, the signal needs to pass through an ADC to be converted into raw bits.
 In his thesis, Clason \cite{Clason2023} suggests a discrete ADC chip capable of analyzing frequencies higher than 25 MHz, as this is the highest frequency studied in his work.
@@
 
-Many microcontrollers furthermore come equipped with internal ADCs that can be utilized, and while these provide a lower sample size (often around 1 MSPS), the ease of development may be prudent to utilize for this proof-of-concept.
+Many microcontrollers furthermore come equipped with internal ADCs that can be utilized, and while these provide a lower sample size (\emph{often around 1 MSPS}), the ease of development may be prudent to utilize for this proof-of-concept.
 While our initial ADC has a fairly low throughput, this can always be upgraded if it ends up becoming too limiting.
 
-Should these internal ADCs prove too limiting, we propose utilizing MAX11102AUB\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} with an effective sample rate of 2 million samples per second (MSPS).
+Should these internal ADCs prove too limiting, we propose utilizing MAX11102AUB\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} with an effective sample rate of 2 million samples per second (\emph{MSPS}).
 This ADC provides a 12 bit sample size, providing roughly 24 Mbit/s of sampled data per second, derived by the following calculation.
@@
 
-Microcontrollers (MCUs) are compact and low-power computing devices designed for embedded systems and real-time operations, and suitable as a processing unit for the purposes of this work.
-Unlike general purpose CPUs, an MCU integrates a processor, memory and peripherals (\emph{such as an ADC, for instance}) into a single chip.
+Microcontrollers (\emph{MCUs}) are compact and low-power computing devices designed for embedded systems and real-time operations, and suitable as a processing unit for the purposes of this work.
+Unlike general CPUs, an MCU integrates a processor, memory and peripherals (\emph{such as an ADC}) into a single chip.
 Furthermore, modern MCUs often feature advanced microarchitectural elements to enhance processing capabilities on single threads (\emph{such as dual-issue superscalar architectures, allowing the MCU to run several instructions per CPU cycle}), making them suitable candidates for the post-processing required for OQRNG-data.
@@
 High-performance MCUs enhance memory usage in various ways.
-Some of them use Tightly Coupled Memory\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} (TCM), which gives fast SRAM with specific access routes for important data, avoiding cache misses and guaranteeing consistent performance.
-Moreover, instruction and data caching techniques, including instruction prefetching and branch prediction, help minimize execution delays in computationally intensive real-time applications.
+Some of them use Tightly Coupled Memory\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} (\emph{TCM}), which gives fast SRAM with specific access routes for important data, avoiding cache misses and guaranteeing consistent performance.
+Moreover, instruction and data caching techniques, including instruction pre-fetching and branch prediction, help minimize execution delays in computationally intensive real-time applications.
 Another important aspect is Direct Memory Access\footnote{\href{https://www.sciencedirect.com/topics/computer-science/direct-memory-access}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (DMA), which enables data transfer between peripherals such as the ADC and RAM, without CPU intervention.
 This offloading reduces processing overhead, allowing the MCU to manage fast data transfers effectively.
-These improvements are especially significant for Toeplitz extraction, where large amount of random data needs to be processed and sent quickly with low delay.
+These improvements are especially significant for Toeplitz extraction, where large amount of random data needs to be processed and sent quickly with low delays.
 Efficient memory management guarantees that randomness extraction can occur rapidly without major slowdowns in computing.
@@
 This MCU is especially suitable for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and Digital Signal Processing (\emph{DSP}) capabilities.
-The floating-point unit\footnote{\href{https://www.sciencedirect.com/topics/computer-science/floating-point-unit}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (FPU) and Single Instruction, Multiple Data (\emph{SIMD}) style DSP instructions improves how quickly it can perform bitwise and arithmetic tasks, which are crucial for quick Toeplitz extraction.
+The floating-point unit\footnote{\href{https://www.sciencedirect.com/topics/computer-science/floating-point-unit}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (\emph{FPU}) and Single Instruction, Multiple Data (\emph{SIMD}) style DSP instructions improve how quickly it can perform bitwise and arithmetic tasks, which is crucial for quick Toeplitz extraction.
 SIMD-controlled DSP architectures, as described by Han et al.
@@
 The raw bits from the ADC can potentially have some deterministic patterns, and as such have to be processed somehow in order to remove these patterns.
-Several methods exist for this purpose, and for our work, we will perform this preprocessing via Toeplitz extraction.
+Several methods exist for this purpose, and for our work, we will perform this pre-processing via Toeplitz extraction.
 The main focus of this study is to implement this extraction algorithm as effectively as possible on resource constrained hardware.
@@
 
-To summarize the theoretical working of Toeplitz extraction (as explained by Chouhan et al.
-\cite{toeplitz-desc}), a pre-determined seed matrix (\(T\)) is multiplied with the sampled raw bit matrix (\(K\)).
+To summarize the theoretical working of Toeplitz extraction (\emph{as explained by Chouhan et al.
+\cite{toeplitz-desc}}), a pre-determined seed matrix (\(T\)) is multiplied with the sampled raw bit matrix (\(K\)).
 The size of the seed is directly dependent on the size of the sampled data, and can be fixed or continually re-sampled as needed.
 To ensure high levels of entropy, our intuition is that re-sampling the seed from the OQRNG-device continually is prudent.
-The sample and seed with then be processed with matrix multiplication to remove deterministic patterns, and produce a bitstring that results in our randomly generated number.
+The sample and seed will then be processed with matrix multiplication to remove deterministic patterns, and produce a bitstring that results in our randomly generated number.
 An example of how this extraction works can be seen below.
 
-\begin{verbatim}
-Where n = number of input bits,
-      m = number of output bits:
-
-Let X be the input array (length n)
-Let T be the seed matrix (length n + m – 1)
-Let Y be a new array of length m
-
-For i from 0 to m - 1:
-  sum = 0
-  For j from 0 to n - 1:
-    sum = sum + X[j] * T[i + j]
-  Y[i] = sum mod 2
-
-Return Y
-\end{verbatim}
+\begin{algorithm}[ht] \caption{Toeplitz extraction}\label{alg:bit-conv}
+\begin{algorithmic}[1] \REQUIRE \(x[0\,..\,n-1]\) \Comment{input bit array of
+length \(n\)} \REQUIRE \(t[0\,..\,n+m-2]\) \Comment{seed matrix of length
+\(n+m-1\)} \ENSURE \(y[0\,..\,m-1]\) \Comment{output bit array of length \(m\)}
+\FOR{\(i = 0\) to \(m - 1\)} \STATE sum = 0 \FOR{\(j = 0\) to \(n - 1\)} \STATE
+sum = sum + \(x[j] \* t[i + j]\) \ENDFOR \STATE \(y[i] =
+\mathrm{sum}\;\bmod\;2\) \ENDFOR \RETURN \(y\) \end{algorithmic} \end{algorithm}
 
@@
 With the assumption that the OQRNG-device produces a truly random analog signal, we can now clearly define the scope in which this thesis operates.
-Considering the maximum conversion speed from the ADC and the USB-output from the MCU, we have a clear bound over 24 Mbit/s (imposed by the ADC) in which Toeplitz extraction needs to be executed.
+Considering the maximum conversion speed from the ADC and the USB-output from the MCU, we have a clear bound over 24 Mbit/s (\emph{imposed by the ADC}) in which Toeplitz extraction needs to be executed.
 Any speeds over 2.86 MB/s allows us to upgrade the ADC iteratively to continue increasing the output speed.
@@
 Thus, minimizing temporal correlations between extractions.
-Furthermore, an external seed updating mechanism via PCIe ensures that seeds are refreshed whenever a certain security limit is reached.
+Furthermore, an external seed updating mechanism via Peripheral Component Interconnect Express (\emph{PCIe}) ensures that seeds are refreshed whenever a certain security limit is reached.
 Compared to fixed-seed methods, this renewable approach enhances cryptographic robustness and ensures sustained high-security randomness extraction in real-world applications.
 
-Efficient Toeplitz matrix-vector multiplication (TMVM) is critical for optimizing randomness extraction that relies on Toeplitz, particulary in constrained hardware environments.
+Efficient Toeplitz matrix-vector multiplication (\emph{TMVM}) is critical for optimizing randomness extraction that relies on Toeplitz, particulary in constrained hardware environments.
 Liao et al.
-\cite{liao} showed that this process could be greatly accelerated using Fast Fourier Transform (FFT) and its inverse (IFFT).
-Thus, reducing computational complexity from \(O(n^2)\) to \(O(n log n)\).
+\cite{liao} showed that this process could be greatly accelerated using Fast Fourier Transform (\emph{FFT}) and its inverse (\emph{IFFT}) -- reducing computational complexity from \(O(n^2)\) to \(O(n \log n)\).
 Their implementation on FPGA utilized this approach for deep neural networks, resulting in a 28.7 times decrease in model size while still achieving fast inference speeds.
@@
 al.
-\cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements of an algorithms execution speed.
+\cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements of the execution speed of an algorithm.
 An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (as the process for this may vary greatly between controllers).
@@
 \[
-\text{Throughput}{\mathrm{Mbit/s}}
-= \frac{DataSize{\mathrm{bits}}}{ExecutionTime_{\mathrm{ms}}}
+Throughput_\mathrm{Mbit/s}
+= \frac{DataSize_\mathrm{bits}}{ExecutionTime_\mathrm{ms}}
 \times 10^{-3}
@@
 \[
-T_{\mathrm{ms}}
-= \frac{64}{24}\times10^{-3}\ \mathrm{ms}
+\frac{64}{24}\times10^{-3}\ \mathrm{ms}
 \approx 2.667\times10^{-3}\ \mathrm{ms}
@@
 
-\subsubsection{5.3 Limitations}\label{limitations}
+\subsubsection{5.5 Limitations}\label{limitations}
 
-Our proposed iterations all assume that the limited hardware will support it.
-Whereas we are confident that Teensy 4.1 will be able to handle each iteration step (even the naive implementation), the remaining microcontrollers with lower specifications might not be suitable for the first iterations.
+Our proposed iterations all assume that the limited hardware will support it. Whereas we are confident that Teensy 4.1 will be able to handle each iteration
+
+step (\emph{even the naive implementation}), the remaining microcontrollers with lower specifications might not be suitable for the first iterations.
 Testing the implementations on different microcontrollers could turn out to be unfeasible -- however, this remains to be seen during the experimentation.
@@
 
-Table 1 presents the execution speeds of each iteration as listed in sections 5.3 and 5.4, alongside the exectution speed in Mbit/s as calculated by utilizing the average value in (3).
+Table 1 presents the execution speeds of each iteration as listed in sections 5.3 and 5.4, alongside the execution speed in Mbit/s as calculated by utilizing the average value in (3).
 My thought then is that we here talk about the results and what they demonstrate, before moving on to draw conclusions and attempt to answer our research questions in section 7.

### 2025-05-21 pres/common.tex
apparently missed some files?

\documentclass[noamsthm, english, xcolor=x11names, xcolor=svgnames, xcolor=dvipsnames]{beamer}
\usepackage[swedish]{babel}
\usepackage{etoolbox} % set
\usepackage{parskip}
\usepackage{graphicx,txfonts}
\usepackage[T1]{fontenc}
\usepackage{enumitem}
\usepackage{mathabx}
\usepackage{alltt}
\usetheme[
  dark,%
  %% Standard colors {blue, green, turquoise} for title, outline and end pages
  titlecolor=blue,%
  outlinecolor=green,%
  endcolor=turquoise,%
  %% Complementary colors {orange, purple, yellow, gray} for some details
  complementary=orange,%
  %% All colors {blue, green, turquoise, orange, purple, yellow, gray} for blocks
  blockcolor=blue,%
  %% Font schemes {latex, standard, sansserif, calibri, liu} to use
  font=calibri,%
  %license=ccby,%
  %% Select color themes {highcontrast, superhighcontrast} with high contrast or add handout option in documentclass
  %highcontrast,%
  %superhighcontrast,%
  %% Remove outline before each section and end page with {nooutline, noendpage}
  outline,%
  %nooutline, %
  %noendpage,%
  %% Add navigation symbols with {navigation}
  %navigation,%
  %% Add total number of frames to frame count with {totalframes}
  totalframes,%
  % Remove parts of header with {noheadertitle, noheaderauthor, noheaderdate, noheadernumber, minimalheader, noheader}
  noheadertitle,%
  noheaderauthor,% 
  noheaderdate,%
  %noheadernumber,%
  %% Show institute in title frame
  showinstitute,%
  %% Show outline in two columns
  %outlinecolumns,
  %footerimage,
]{LiU}

% make LiU-logo black
\makeatletter
\def\LiU@logo@color@footline{black}
\makeatother

\usepackage{listings}
\usepackage{tikz}
\usepackage{../tikz-uml}
\usepackage[x11names]{xcolor}
\usepackage[forceshell,pgf,debug,outputdir={graphs/}]{dot2texi}

\setitemize{label=\usebeamerfont*{itemize item}%
  \usebeamercolor[fg]{itemize item}
  \usebeamertemplate{itemize item}}
\setenumerate[1]{%
  label=\protect\usebeamerfont{enumerate item}%
        \protect\usebeamercolor[fg]{enumerate item}%
        \insertenumlabel.}
\setenumerate[2]{%
  label=\protect\usebeamerfont{enumerate subitem}%
        \protect\usebeamercolor[fg]{enumerate subitem}%
        \insertsubenumlabel.}
%% \setenumerate[2]{%
%%   label=\protect\usebeamerfont{enumerate subitem}%
%%         \protect\usebeamercolor[fg]{enumerate subitem}%
%%         (\alph{enumii})}

%% Add a text on end page, leave blank to add author
\finaltext{}

\usetikzlibrary{shapes,arrows}
\usetikzlibrary{calc, tikzmark, fit}
\usetikzlibrary{positioning, decorations.markings}
\usetikzlibrary{patterns,snakes}

\setmonofont{Liberation Mono}
\lstset{basicstyle=\scriptsize\fontspec{LiberationMono}}
\date{}

\definecolor{LiUBlue}{RGB}{0,185,231}
\definecolor{LiUTurquoise}{RGB}{23,199,210}
\definecolor{LiUGreen}{RGB}{0,207,181}
\definecolor{LiUOrange}{RGB}{255,100,66}
\definecolor{LiUPurple}{RGB}{137,129,211}
\definecolor{LiUYellow}{RGB}{253,239,93}
\definecolor{LiUGray}{RGB}{106,126,145}

\tikzumlset{font=\tiny}

\lstdefinelanguage[11]{C++}[ISO]{C++}{
  morekeywords={alignas,alignof,char16_t,char32_t,constexpr,override,%
    decltype,noexcept,nullptr,static_assert,thread_local,final, nullopt},%
}

%\documentclass[english]{liu}
\lstset{ %
  language=ruby, %
  commentstyle=\color{ForestGreen}, %
  keywordstyle=\color{blue}, %
  stringstyle=\color{BrickRed}, %
  captionpos=t, %
  frame=single, %
  tabsize=2, %
  rulecolor=\color{black}, %
  showstringspaces=false, %
  backgroundcolor=\color{WhiteSmoke},%
  extendedchars=true, %
  escapeinside={@(}{)@},          % if you want to add LaTeX within your code
  basicstyle=\ttfamily\footnotesize 
}

\lstdefinestyle{run}{%
  frame=single,%
  backgroundcolor=\color{Ivory1},%
  language=bash,%
  showstringspaces=false, %
  escapeinside={@(}{)@},          % if you want to add LaTeX within your code
}

\lstdefinelanguage{none}{
  identifierstyle=
}

\lstdefinestyle{terminal}{%
  frame=single,%
  backgroundcolor=\color{Ivory1},%
  language=none,%
  showstringspaces=false, %
  escapeinside={@(}{)@},          % if you want to add LaTeX within your code
  basicstyle=\ttfamily\tiny
}

\lstdefinestyle{noterminal}{%
  style=terminal,%
  frame=none,
  backgroundcolor=\color{white},%
}

\lstdefinestyle{small}{%
  basicstyle=\ttfamily\tiny
}

\lstdefinestyle{medium}{%
  basicstyle=\fontsize{7}{9}\selectfont\ttfamily
}

\lstdefinestyle{smallrun}{%
  style=run,
  basicstyle=\ttfamily\tiny
}

\lstdefinestyle{basic}{
  style=small,
  backgroundcolor=\color{white},%
  frame=none,
}

\hypersetup{
	colorlinks=true,
        urlcolor=cyan,
        linkcolor=.
}

\setbeamercolor*{section in toc}{fg=black}
\setbeamercolor*{subsection in toc}{fg=black}
\setbeamercolor*{section in toc shaded}{fg=LiUgreen0!80!black}
\setbeamercolor*{subsection in toc shaded}{fg=LiUgreen0!80!black}
\setbeamertemplate{section in toc shaded}{\leavevmode\leftskip=2.5em{\hspace{-2.5em}\makebox[2em]{\inserttocsectionnumber}\inserttocsection\par}}
\setbeamertemplate{subsection in toc shaded}{\leavevmode\leftskip=2.9em{\hspace{-0.5em}\inserttocsubsection\par}}

% Use lstinline as item in description
\makeatletter
\newcommand*{\lstitem}[1][]{%
  \setbox0\hbox\bgroup
    \patchcmd{\lst@InlineM}{\@empty}{\@empty\egroup\item[\usebox0]\leavevmode\ignorespaces}{}{}%
    \lstinline[#1]%
}
\makeatother

\newcommand{\lst}[1]{\lstinline{#1}}
\newcommand{\codestyle}[1]{\footnotesize\fontspec{LiberationMono}\texttt{#1}}
\newcommand{\smallcodestyle}[1]{\tiny\fontspec{LiberationMono}\texttt{#1}}
\newfontface\extrafont{Liberation Serif}

\tikzset{fun/.style={inner sep=0, anchor=west, outer sep=0}}

\tikzset{var/.style={minimum height=0.4cm, minimum width=0.4cm, font=\scriptsize\ttfamily}}
\tikzset{var val/.style={var, rectangle, draw=black, font=\tiny\ttfamily}}
\tikzset{myptr/.style={>=latex, decoration={markings,mark=at position 1 with %
      {\arrow[scale=2]{>}}},postaction={decorate}}}
\tikzset{array element/.style={rectangle, thick, draw=black, minimum width=0.7cm, minimum height=0.7cm, font=\ttfamily}}

\tikzset{class/.style={rectangle, draw=black, anchor=center}}

\newcommand{\varbox}[3]
{
  % #1: Name of this variable
  % #2: Value of this variable
  % #3: position
  \node[var, anchor=east] (#1 label) at (#3) {\texttt{#1}};
  \node[var val, anchor=west, xshift=-\pgflinewidth] (#1 value) at (#1 label.east) {\texttt{#2}};
  \node[fit=(#1 label)(#1 value), inner sep=0] (#1) {};
}

\newcommand{\fullvarbox}[4]
{
  % #1: Name of this shape
  % #2: Name of the variable
  % #3: Value of this variable
  % #4: position
  \node[var, anchor=east] (#1 label) at (#4) {\texttt{#2}};
  \node[var val, anchor=west, xshift=-\pgflinewidth] (#1 value) at (#1 label.east) {\texttt{#3}};
  \node[fit=(#1 label)(#1 value), inner sep=0] (#1) {};
}

\newcommand{\extendedvarbox}[4]
{
  % #1: Name of this variable
  % #2: Value of this variable
  % #3: position
  % #4: data type
  \node[var, anchor=east] (#1 label) at (#3) {\texttt{#1}};
  \node[var val, anchor=west, xshift=-\pgflinewidth] (#1 value) at (#1 label.east) {\texttt{#2}};
  \node[var, anchor=north] (#1 type) at (#1 value.south) {\lstinline[style=small]{#4}};
  \node[fit=(#1 label)(#1 value)(#1 type), inner sep=0] (#1) {};
}

\newcommand{\containerbox}[5]
{
  % #1: name of containerbox
  % #2: left most varbox
  % #3: right most varbox
  % #4: top most varbox
  % $5: bottom most varbox

  \draw[thick] let 
  \p{left} = ($(#2.west) - (0.05cm, 0)$), 
  \p{right} = ($(#3.east) + (0.025cm, 0)$),
  \p{top} = ($(#4.north) + (0, 0.05cm)$),
  \p{bottom} = ($(#5.south) - (0, 0.05cm)$),
  \p{top left} = (\x{left}, \y{top}),
  \p{bottom right} = (\x{right}, \y{bottom}),
  \p{bottom left} = (\x{left}, \y{bottom}),
  \p{midpoint} = ($0.5*(\p{bottom left}) + 0.5*(\p{bottom right})$),
  \n{width} = {\x{right} - \x{left}},
  \n{height} = {\y{top} - \y{bottom}}
  in
  node[anchor=north west, minimum width=\n{width}, font=\scriptsize, inner sep=0.1cm] (#1 label) at (\p{bottom left}) {#1}
  node[anchor=north west, rectangle, draw=black, thick, minimum width=\n{width}, minimum height={\n{height}}] (#1 container) at (\p{top left}) {}
  node[fit=(#1 container)(#1 label), inner sep = 0] (#1) {};
}

\newcommand{\nullptr}[1]{
  % #1: which box to make null
  \draw ($(#1.north west) + (\pgflinewidth, -\pgflinewidth)$) -- ($(#1.south east) + (-\pgflinewidth, \pgflinewidth)$);
}

\newcommand{\unused}[1]{
  % #1: which box to make null
  \draw ($(#1.north west) + (0.1cm, -0.1cm)$) -- ($(#1.south east) + (-0.1cm, 0.1cm)$);
  \draw ($(#1.north east) + (-0.1cm, -0.1cm)$) -- ($(#1.south west) + (0.1cm, 0.1cm)$);
}

\tikzset{box/.style={minimum height=0.5cm, minimum width=0.5cm, rectangle, draw=black}}
\tikzset{ptrbox/.style={box, minimum width=0.2cm}}

\newcommand{\nodebox}[3]{
  % #1: name
  % #2: position
  % #3: value
  \node[box, anchor=center] (#1 value) at (#2) {\scriptsize\texttt{#3}};
  \node[ptrbox, anchor=west, xshift=-\pgflinewidth] (#1 next) at (#1 value.east) {};
  \node[inner sep=0, fit=(#1 value)(#1 next)] (#1) {};
}

\newcommand{\pointer}[2]{
  % #1: from
  % #2: to
  \draw[->, >=latex] (#1 next.center) -- (#2.west);
}

\newcommand{\binodebox}[3]{
  % #1: name
  % #2: position
  % #3: value
  \node[box, anchor=center] (#1 value) at (#2) {\scriptsize\texttt{#3}};
  \node[ptrbox, anchor=east, xshift=\pgflinewidth] (#1 prev) at (#1 value.west) {};
  \node[ptrbox, anchor=west, xshift=-\pgflinewidth] (#1 next) at (#1 value.east) {};
  \node[inner sep=0, fit=(#1 prev)(#1 value)(#1 next)] (#1) {};
}

\newcommand{\bipointer}[2]{
  % #1: from
  % #2: to
  \draw[->, >=latex, transform canvas={yshift=0.1cm}] (#1 next.center) -- (#2.west);
  \draw[<-, >=latex, transform canvas={yshift=-0.1cm}] (#1.east) -- (#2 prev.center);
}

\newcommand{\fullcontainerbox}[6]
{
  % #1: name of containerbox
  % #2: label of containerbox
  % #3: left most varbox
  % #4: right most varbox
  % #5: top most varbox
  % $6: bottom most varbox

  \draw[thick] let 
  \p{left} = ($(#3.west) - (0.05cm, 0)$), 
  \p{right} = ($(#4.east) + (0.025cm, 0)$),
  \p{top} = ($(#5.north) + (0, 0.05cm)$),
  \p{bottom} = ($(#6.south) - (0, 0.05cm)$),
  \p{top left} = (\x{left}, \y{top}),
  \p{bottom right} = (\x{right}, \y{bottom}),
  \p{bottom left} = (\x{left}, \y{bottom}),
  \p{midpoint} = ($0.5*(\p{bottom left}) + 0.5*(\p{bottom right})$),
  \n{width} = {\x{right} - \x{left}},
  \n{height} = {\y{top} - \y{bottom}}
  in
  node[anchor=north west, minimum width=\n{width}, font=\scriptsize, inner sep=0.1cm] (#1 label) at (\p{bottom left}) {#2}
  node[anchor=north west, rectangle, draw=black, thick, minimum width=\n{width}, minimum height={\n{height}}] (#1 container) at (\p{top left}) {}
  node[fit=(#1 container)(#1 label), inner sep = 0] (#1) {};
}

\makeatletter
\newcommand{\createarray}[4]{%
  % #1: x
  % #2: y
  % #3: name
  % Every other argument is a new box and its content
  \edef\arrstart{#1}
  \edef\arrheight{#2}%
  \edef\arrname{#3}%
  \pgfmathtruncatemacro{\arrcounter}{0}%
  \node[array element] (\arrname \arrcounter) at ({\arrstart + 0.7*\arrcounter}, \arrheight) {#4};%
  \pgfmathtruncatemacro{\arrcounter}{\arrcounter + 1}%
  \arraynextarg%
}
\newcommand{\arraynextarg}{\@ifnextchar\bgroup{\arraygobblearg}{}}
\newcommand{\arraygobblearg}[1]{%
  \node[array element] (\arrname \arrcounter) at ({\arrstart + 0.7*\arrcounter}, \arrheight) {#1};%
  \pgfmathtruncatemacro{\arrcounter}{\arrcounter + 1}%
  \@ifnextchar\bgroup{\arraygobblearg}{}}
\makeatother

\makeatletter
\newcommand{\createverticalarray}[4]{%
  % #1: x
  % #2: y
  % #3: name
  % Every other argument is a new box and its context
  \edef\arrstart{#2}
  \edef\arrwidth{#1}%
  \edef\arrname{#3}%
  \pgfmathtruncatemacro{\arrcounter}{0}%
  \node[array element] (\arrname \arrcounter) at (\arrwidth, {\arrstart - 0.7*\arrcounter}) {#4};%
  \pgfmathtruncatemacro{\arrcounter}{\arrcounter + 1}%
  \verticalarraynextarg%
}
\newcommand{\verticalarraynextarg}{\@ifnextchar\bgroup{\verticalarraygobblearg}{}}
\newcommand{\verticalarraygobblearg}[1]{%
  \node[array element] (\arrname \arrcounter) at (\arrwidth, {\arrstart - 0.7*\arrcounter}) {#1};%
  \pgfmathtruncatemacro{\arrcounter}{\arrcounter + 1}%
  \@ifnextchar\bgroup{\verticalarraygobblearg}{}}
\makeatother

\newcommand{\underscore}[2]{
  \draw[draw=#2] ($(pic cs:#1beg) - (0.05cm, 0.05cm)$) -- ($(pic cs:#1end) + (0.05cm, -0.05cm)$);
  \draw let 
  \p{left} = ($(pic cs:#1beg) - (0.05cm, 0.05cm)$), 
  \p{right} = ($(pic cs:#1end) + (0.05cm, -0.05cm)$),
  \p{mid} = ($0.5*(\p{left}) + 0.5*(\p{right})$)
  in
  (\p{left}) -- (\p{right})
  coordinate (#1) at (\p{mid});
}

\newcommand{\highlight}[2]{\draw[draw=#2, thick] ($(pic cs:#1beg) - (0.05cm, 0.1cm)$) rectangle ($(pic cs:#1end) + (0.05cm, 0.1cm) + (0, 1ex)$);}

\newcommand{\smallhighlight}[2]{\draw[draw=#2, thick] ($(pic cs:#1beg) - (0.025cm, 0.05cm)$) rectangle ($(pic cs:#1end) + (0.025cm, 0.05cm) + (0, 0.7ex)$);}

\newcommand{\highlightbox}[2]{
  \coordinate (#1 bottomleft) at ($(pic cs:#1beg) + (0.1cm, 0)$);
  \coordinate (#1 topright) at ($(pic cs:#1end) + (-0.1cm, 1ex)$);
  \node[fit=(#1 bottomleft)(#1 topright), draw=#2, thick] (#1) {};
}

\newcommand{\pointto}[5]{\draw[->, >=stealth, thick, draw=red, red] (#1) to [#5] node[midway, #4, red]{#3}(#2);}

\newcommand{\crossout}[1]{\draw[draw=black, thick] ($(pic cs:#1beg) + (0, 1ex)$) -- ($(pic cs:#1end)$);\draw[draw=black, thick] ($(pic cs:#1end) + (0, 1ex)$) -- ($(pic cs:#1beg)$);}

\newcommand{\vtable}[4]
{
  \node[align=left, anchor=north west, font=\scriptsize\ttfamily, inner sep=0] (#1 vtable) at (#2) {
    \begin{tabular}{|l|}
      \hline
      \textbf{vtable for #1} \\
      \hline
      #3 \\
      \hline
    \end{tabular}
  };
}

\newcommand{\pointtocode}[3]
{
  \draw[->, thick,draw=#3] ($(pic cs:#1) + (0, 0.3ex)$) -- ($(pic cs:#2) + (0, 0.3ex)$);
}

\title{Efficient Exctraction of True Random Numbers from Quantum System on Resource Constrained Hardware}
\author{Love Arreborn and Nadim Lakrouz}
\institute{Institutionen för datavetenskap}

### 2025-05-21 pres/opp/stack.rb

@@ -1,0 +1,351 @@
+#!/usr/bin/env ruby
+
+# This file is called rdparse.rb because it implements a Recursive
+# Descent Parser. Read more about the theory on e.g.
+# http://en.wikipedia.org/wiki/Recursive_descent_parser
+
+# 2010-02-11 New version of this file for the 2010 instance of TDP007
+#   which handles false return values during parsing, and has an easy way
+#   of turning on and off debug messages.
+# 2014-02-16 New version that handles { false } blocks and :empty tokens.
+
+require 'logger'
+
+class Rule
+
+  # A rule is created through the rule method of the Parser class, like this:
+  #   rule :term do
+  #     match(:term, '*', :dice) {|a, _, b| a * b }
+  #     match(:term, '/', :dice) {|a, _, b| a / b }
+  #     match(:dice)
+  #   end
+  
+  Match = Struct.new :pattern, :block
+  
+  def initialize(name, parser)
+    @logger = parser.logger
+    # The name of the expressions this rule matches
+    @name = name
+    # We need the parser to recursively parse sub-expressions occurring 
+    # within the pattern of the match objects associated with this rule
+    @parser = parser
+    @matches = []
+    # Left-recursive matches
+    @lrmatches = []
+  end
+  
+  # Add a matching expression to this rule, as in this example:
+  #   match(:term, '*', :dice) {|a, _, b| a * b }
+  # The arguments to 'match' describe the constituents of this expression.
+  def match(*pattern, &block)
+    match = Match.new(pattern, block)
+    # If the pattern is left-recursive, then add it to the left-recursive set
+    if pattern[0] == @name
+      pattern.shift
+      @lrmatches << match
+    else
+      @matches << match
+    end
+  end
+  
+  def parse
+    # Try non-left-recursive matches first, to avoid infinite recursion
+    match_result = try_matches(@matches)
+    return nil if match_result.nil?
+    loop do
+      result = try_matches(@lrmatches, match_result)
+      return match_result if result.nil?
+      match_result = result
+    end
+  end
+
+  private
+  
+  # Try out all matching patterns of this rule
+  def try_matches(matches, pre_result = nil)
+    match_result = nil
+    # Begin at the current position in the input string of the parser
+    start = @parser.pos
+    matches.each do |match|
+      # pre_result is a previously available result from evaluating expressions
+      result = pre_result.nil? ? [] : [pre_result]
+
[…]

### 2025-05-21 pres/opp/stacksolution.rb

@@ -1,0 +1,354 @@
+#!/usr/bin/env ruby
+
+# This file is called rdparse.rb because it implements a Recursive
+# Descent Parser. Read more about the theory on e.g.
+# http://en.wikipedia.org/wiki/Recursive_descent_parser
+
+# 2010-02-11 New version of this file for the 2010 instance of TDP007
+#   which handles false return values during parsing, and has an easy way
+#   of turning on and off debug messages.
+# 2014-02-16 New version that handles { false } blocks and :empty tokens.
+
+require 'logger'
+
+class Rule
+
+  # A rule is created through the rule method of the Parser class, like this:
+  #   rule :term do
+  #     match(:term, '*', :dice) {|a, _, b| a * b }
+  #     match(:term, '/', :dice) {|a, _, b| a / b }
+  #     match(:dice)
+  #   end
+  
+  Match = Struct.new :pattern, :block
+  
+  def initialize(name, parser)
+    @logger = parser.logger
+    # The name of the expressions this rule matches
+    @name = name
+    # We need the parser to recursively parse sub-expressions occurring 
+    # within the pattern of the match objects associated with this rule
+    @parser = parser
+    @matches = []
+    # Left-recursive matches
+    @lrmatches = []
+  end
+  
+  # Add a matching expression to this rule, as in this example:
+  #   match(:term, '*', :dice) {|a, _, b| a * b }
+  # The arguments to 'match' describe the constituents of this expression.
+  def match(*pattern, &block)
+    match = Match.new(pattern, block)
+    # If the pattern is left-recursive, then add it to the left-recursive set
+    if pattern[0] == @name
+      pattern.shift
+      @lrmatches << match
+    else
+      @matches << match
+    end
+  end
+  
+  def parse
+    # Try non-left-recursive matches first, to avoid infinite recursion
+    match_result = try_matches(@matches)
+    return nil if match_result.nil?
+    loop do
+      result = try_matches(@lrmatches, match_result)
+      return match_result if result.nil?
+      match_result = result
+    end
+  end
+
+  private
+  
+  # Try out all matching patterns of this rule
+  def try_matches(matches, pre_result = nil)
+    match_result = nil
+    # Begin at the current position in the input string of the parser
+    start = @parser.pos
+    matches.each do |match|
+      # pre_result is a previously available result from evaluating expressions
[…]

### 2025-05-21 docs/style/article-final.tex

@@
 \usepackage{textcomp}
+\usepackage{tikz}
+\usetikzlibrary{positioning}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 

### 2025-05-21 docs/style/article-body.tex
praying for no merge ocnflicts

@@
 Imagine, then, if a malicious attacker somehow manages to ascertain the state a computer was in when it generated a random number, for instance to produce an SSH-key.
-This attacker then has the opportunity to accurately reproduce the exact, deterministic state that produced said random number, in essence removing the safety that randomness brings.
+This hypothetical attacker has the opportunity to accurately reproduce the exact, deterministic state that produced said random number, in essence removing the safety that randomness brings.
 While it may sound unrealistic, the exponential increase in processing power and the burgeoning field of quantum computing does introduce the possibility that one day, what we perceive as random is nothing more than a simple algorithm to crack.
@@
 The post-processing finally yields a random number.
-This relationship can be seen in Figure 1.
+This relationship can be seen in Figure \ref{fig:linear-system}.
 
@@
 
-\subsubsection{5.1 Implementation strategy}\label{implementation-strategy}
-
 Our implementation of the Toeplitz extractor followed a structured, iterative approach divided into two phases.
@@
 
-\paragraph{5.2 Phase one}\label{phase-one}
+\subsubsection{5.1 Phase one}\label{phase-one}
 
@@
 
-The initial implementation followed the pseudocode described in section 3.4, using matrix multiplication over raw input and seed data.
+The initial implementation followed the pseudocode described in Algorithm \ref{alg:bit-conv}, using matrix multiplication over raw input and seed data.
 It relied on \texttt{std::vector\textless{}int\textgreater{}} for storage and used nested loops to compute each output bit.
@@
 
-\paragraph{5.3 Phase two}\label{phase-two}
+\subsubsection{5.2 Phase two}\label{phase-two}
 
@@
 
-\subsubsection{5.4 Evaluation}\label{evaluation}
+\subsubsection{5.3 Evaluation}\label{evaluation}
 
@@
 
-\subsubsection{5.5 Limitations}\label{limitations}
+\subsubsection{5.4 Limitations}\label{limitations}
 
-Our proposed iterations all assume that the limited hardware will support it. Whereas we are confident that Teensy 4.1 will be able to handle each iteration
-
-step (\emph{even the naive implementation}), the remaining microcontrollers with lower specifications might not be suitable for the first iterations.
+Our proposed iterations all assume that the limited hardware will support it.
+Whereas we are confident that Teensy 4.1 will be able to handle each iteration step (\emph{even the naive implementation}), the remaining microcontrollers with lower specifications might not be suitable for the first iterations.
 Testing the implementations on different microcontrollers could turn out to be unfeasible -- however, this remains to be seen during the experimentation.
@@
 Furthermore, the script also provides the average execution time of only the Toeplitz extraction in microseconds.
+A brief overview of the architecture of this test script can be seen in Figure \ref{fig:testscript-architecture}.
+
+\begin{figure*}[ht] \centering
+\includegraphics[width=\textwidth]{img/testscript.png} \caption{A brief overview
+of the testscript architecture.} \label{fig:testscript-architecture}
+\end{figure*}
 

### 2025-05-21 docs/style/article-body.tex
for incoming pull

@@
 
-\textbf{Iteration 6 - Data type exploration:} Following the removal of vectors, this iteration explored alternative static data types to determine the most efficient structure for storing input and seed data.
-
-\textbf{Iteration 7 - Re-implementation of bitshifting:} In this final iteration, bit shifting techniques were reintroduced into the now simplified implementation.
-With high-level abstractions removed and memory structures streamlined, bit-level operations such as shifting and masking were applied.
+\textbf{Iteration 6 - Data type exploration:} Following the removal of vectors, this iteration explored alternative static data types such as: \texttt{array}, \texttt{unordered\_map} and \texttt{bitset} to determine the most efficient structure for storing input and seed data.
 
@@
 
-\begin{tabular}{|c|c|c|c|c|c|c|} \hline \textbf{Bit size} &
-\multicolumn{3}{c|}{\textbf{Teensy ($\mu s$)}} &
-\multicolumn{3}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline & Min & Max & Avg & Min &
-Max & Avg \\ \hline 64 & 13 & 14 & 13.1564 & 103 & 117 & 106.3914 \\ 512 & 788 &
-791 & 788.3139 & 5296 & 5311 & 5302.4979 \\ 1024 & 3124 & 3130 & 3124.0580 &
-21108 & 21119 & 21111.2163 \\ \hline \end{tabular} \caption{Minimum, maximum and
-average processing time for Teensy and Pico.}
-\label{tab:teensy_pico_min_max_avg}
+\subsubsection{6.1 Phase one}\label{phase-one-1}
 
-Table 1 presents the execution speeds of each iteration as listed in sections 5.3 and 5.4, alongside the execution speed in Mbit/s as calculated by utilizing the average value in (3).
-My thought then is that we here talk about the results and what they demonstrate, before moving on to draw conclusions and attempt to answer our research questions in section 7.
+\textbf{Iteration 1 - Data Structures:}
+
+\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
+\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
+\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 13.1564 & 106.3914
+\\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580 & 21111.2163 \\ \hline
+\caption{A brief overview of the testscript architecture.} \label{tab:iter1}
+
+\end{tabular}
+
+Table \ref{tab:iter1} presents the average execution time of iteration 1 on both Teensy 4.1 and Raspberry Pico Pi 2 across thre input sizes.
+The Teensy consistently outperforms the Pico, with the gap widening as the bit size increases.
+
+\textbf{Iteration 2 - Bitshifting:}
+
+\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
+\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline 64 & 16.4689 \\ 512 &
+1006.6255 \\ 1024 & 3996.5972 \\ \hline \end{tabular}
+
+\label{tab:iter2}
+
+Table \ref{tab:iter2} shows the execution times for iteration 2 on the Teensy 4.1, which utilizes bitshifting instead of the original data structure approach.
+A clear increase in executiom time compared to iteration 1 is observed.
+
+Additionally, the isolated effect of applying a single bitmask operation \texttt{\&\ 1} was evaluated.
+This resulted in a slight reduction in average execution time, from \texttt{13.1564\ µs} to \texttt{13.1000\ µs}, although this specific result is not included in the tables above.
+
+\textbf{Iteration 3 - Batching and Hardware optimization:}
+
+\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
+\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline 64/64 & 43.0760 \\ 512/512
+& 2663.8194 \\ 1024/1024 & 10513.1767 \\ \hline \end{tabular}
+
+\label{tab:iter3}
+
+Table \ref{tab:iter3} presents the average execution time of iteration 3 on the Teensy 4.1 for varying batching and bit sizes.
+Additionally, when performing the same 1024/1024-bit operation, the Raspberry Pi Pico 2 showed average execution time of \texttt{80806.006\ µs} indicating a substantially lower throughput compared to the Teensy.
+
+\subsubsection{6.2 Phase two}\label{phase-two-1}
+
+\textbf{Iteration 4 - Loop unrolling:}
+
+\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
+\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
+\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 9.7017 & 70.7402 \\
+512 & 551.6358 & 3978.1928 \\ 1024 & 2195.1979 & 15830.8784 \\ \hline
+\end{tabular}
+
+\label{tab:iter4}
+
+Table \ref{tab:iter4} presents the execution time of iteration 4, which utilizes a 4x unrolled bit-processing loop.
+Compared to erlier iterations, this approach yields a substantial performance improvments on both the Teensy and the Raspberry Pi Pico Mcu's.
+
+In addition to the results presented in Table \ref{tab:iter4}, a further tests was conducted on the Teensy using a single-loop unrolled implementation for the 64-bit size, resulting in an average execution time of \texttt{6.6626\ µs}.
+An additional fully unrolled variant, where loops were entirely eliminated, produced a measured execution time of \texttt{0.0491\ µs}.
+However, this measurement was later determined to be invalid due to packaging error.
+
+\textbf{Iteration 5 - Removal of vector usage:}
+
+\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
+\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
+\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 0.0501 & 0.2175 \\
+\hline \end{tabular}
+
+\label{tab:iter5}
+
+Table \ref{tab:iter5} presents the results of iteration 5, in which vector structure were removed in favor of fixed-size integer types \texttt{uint32\_t} and \texttt{uint64\_t}.
+This iteration led to an extremely efficient implementation that approaches the physical execution limits of the Teensy 4.1.
+
+\textbf{Iteration 6 - Data type exploration:}
+
+\begin{tabular}{|c|c|} \hline \textbf{Data structure} &
+\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline array & 0.4284 \\
+unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline \end{tabular}
+
+\label{tab:iter6}
+
+Table \ref{tab:iter6} presents the results of iteration 6, which was introduced to address the structural limitations encountered in iteration 5.
+Specifically, the fixed-width output constrains imposed by the removal of vector.
+In this iteration, alternative data structures were evaluated.
+Vector-based approach from iteration 1 serving as a baseline for comparison.
 

### 2025-05-21 docs/article-body.tex
microsec fix

--- 2025-05-21 docs/style/article-final.tex
@@
-\documentclass{sigchi}
-
-% Arabic page numbers for submission.  Remove this line to eliminate
-% page numbers for the camera ready copy
-% \pagenumbering{arabic}
-
-% Load basic packages
-\usepackage{balance}       % to better equalize the last page
-\usepackage{graphics}      % for EPS, load graphicx instead 
-\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
-\usepackage{txfonts}
-\usepackage{mathptmx}
-\usepackage[pdflang={en-US},pdftex]{hyperref}
-\usepackage{color}
-\usepackage{booktabs}
-\usepackage{textcomp}
-\usepackage{tikz}
-\usetikzlibrary{positioning}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-
-% Some optional stuff you might like/need.
-\usepackage{microtype}        % Improved Tracking and Kerning
-% \usepackage[all]{hypcap}    % Fixes bug in hyperref caption linking
-\usepackage{ccicons}          % Cite your images correctly!
-% \usepackage[utf8]{inputenc} % for a UTF8 editor only
-
-% If you want to use todo notes, marginpars etc. during creation of
-% your draft document, you have to enable the "chi_draft" option for
-% the document class. To do this, change the very first line to:
-% "\documentclass[chi_draft]{sigchi}". You can then place todo notes
-% by using the "\todo{...}"  command. Make sure to disable the draft
-% option again before submitting your final document.
-\usepackage{todonotes}
-
-% Paper metadata (use plain text, for PDF inclusion and later
-% re-using, if desired).  Use \emtpyauthor when submitting for review
-% so you remain anonymous.
-\def\plaintitle{Efficient Exctraction of True Random Numbers from Quantum System on Resource Constrained Hardware}
-\def\plainauthor{First Author, Second Author, Third Author,
-  Fourth Author, Fifth Author, Sixth Author}
-\def\emptyauthor{}
-\def\plainkeywords{Authors' choice; of terms; separated; by
-  semicolons; include commas, within terms only; this section is required.}
-\def\plaingeneralterms{Documentation, Standardization}
-
-% llt: Define a global style for URLs, rather that the default one
-\makeatletter
-\def\url@leostyle{%
-  \@ifundefined{selectfont}{
-    \def\UrlFont{\sf}
-  }{
-    \def\UrlFont{\small\bf\ttfamily}
-  }}
-\makeatother
-\urlstyle{leo}
-
-% To make various LaTeX processors do the right thing with page size.
-\def\pprw{8.5in}
-\def\pprh{11in}
-\special{papersize=\pprw,\pprh}
-\setlength{\paperwidth}{\pprw}
-\setlength{\paperheight}{\pprh}
-\setlength{\pdfpagewidth}{\pprw}
-\setlength{\pdfpageheight}{\pprh}
-
-% Make sure hyperref comes last of your loaded packages, to give it a
-% fighting chance of not being over-written, since its job is to
-% redefine many LaTeX commands.
-\definecolor{linkColor}{RGB}{6,125,233}
-\hypersetup{%
-  pdftitle={\plaintitle},
-% Use \plainauthor for final version.
-%  pdfauthor={\plainauthor},
-  pdfauthor={\emptyauthor},
-  pdfkeywords={\plainkeywords},
-  pdfdisplaydoctitle=true, % For Accessibility
-  bookmarksnumbered,
-  pdfstartview={FitH},
-  colorlinks,
-  citecolor=black,
-  filecolor=black,
-  linkcolor=black,
-  urlcolor=linkColor,
-  breaklinks=true,
-  hypertexnames=false
-}
-
-% create a shortcut to typeset table headings
-% \newcommand\tabhead[1]{\small\textbf{#1}}
-
-% End of preamble. Here it comes the document.
-\begin{document}
-
-\title{\plaintitle}
-
-\numberofauthors{2}
-\author{%
-	\alignauthor{Love Arreborn\\
-		\email{lovar063@student.liu.se}}\\
-	\alignauthor{Nadim Lakrouz\\
-		\email{nadla777@student.liu.se}}\\
-}
-
-\maketitle
-
-\section{1 INTRODUCTION}\label{introduction}
+\hypertarget{introduction}{%
+\section{1 INTRODUCTION}\label{introduction}}
 
@@
 From generating keys for cryptography, salting password hashes, load balancing in distributed systems, memory addressing and so much more.
-However, the process of producing these random numbers tends to be pseudo-random, e.g.~utilizing the current states of various modules \cite{randomness}.
+However, the process of producing these random numbers tends to be pseudo-random, e.g.~utilizing the current states of various modules {[}@randomness{]}.
 Pseudo-random numbers do not generate true randomness, and in order to heighten security, other methods of generating random numbers are required.
-Current random number generators (\emph{RNG}) are usually implemented in computer programs, using certain states of the host machine as a starting point before running a predetermined algorithm \cite{randomness}.
+Current random number generators (\emph{RNG}) are usually implemented in computer programs, using certain states of the host machine as a starting point before running a predetermined algorithm {[}@randomness{]}.
 This pseudo-random number generation (\emph{PRNG}) comes with the drawback that the result is always deterministic, provided that the initial state is known.
@@
 These systems require some input that is neither replicable nor reproducible.
-One method that can be realistically used is the inherently random movement of lava lamps \cite{lavarand}, which is used as a backup source of randomness for Cloudflare\footnote{\href{https://blog.cloudflare.com/randomness-101-lavarand-in-production/}{Cloudflare.com, accessed 2025-03-10}}.
-Another proposed solution for this is quantum random number generation (\emph{QRNG}) \cite{QRNG}.
+One method that can be realistically used is the inherently random movement of lava lamps {[}@lavarand{]}, which is used as a backup source of randomness for Cloudflare\footnote{\href{https://blog.cloudflare.com/randomness-101-lavarand-in-production/}{Cloudflare.com, accessed 2025-03-10}}.
+Another proposed solution for this is quantum random number generation (\emph{QRNG}) {[}@QRNG{]}.
 By reading quantum fluctuation signals from any given source, for instance an optical signal, the inherent natural unpredictability of said signal can be harnessed in order to produce a random number from a state that is nigh impossible to reproduce accurately.
-Clason \cite{Clason2023} presents a device that generates a fluctuating analogue signal utilizing this method.
+Clason {[}@Clason2023{]} presents a device that generates a fluctuating analogue signal utilizing this method.
 
-The optical signal output by the device described in \cite{Clason2023} needs to be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}).
+The optical signal output by the device described in {[}@Clason2023{]} needs to be converted to a stream of random, raw bits via an Analog to Digital Converter (\emph{ADC}).
 Some post-processing of the said raw bits has to be performed in order to ensure that the bits are workable, and to remove potential deterministic patterns from the data.
-One method for this post-processing we will explore in this work is Toeplitz extraction \cite{toeplitz}, typically performed on the host computer utilizing the randomly generated numbers.
+One method for this post-processing we will explore in this work is Toeplitz extraction {[}@toeplitz{]}, typically performed on the host computer utilizing the randomly generated numbers.
 The post-processing finally yields a random number.
@@
 
-Clason proposes a simpler and cheaper way to achieve QRNG \cite{Clason2023} than what has been done in previous research.
+Clason proposes a simpler and cheaper way to achieve QRNG {[}@Clason2023{]} than what has been done in previous research.
 These systems have generally been large, bulky and expensive, whereas the method proposed is portable enough to theoretically be installed in a portable device -- for instance a USB-thumbstick.
@@
 al.
-\cite{StefanovOptical} proposes using the random choice of a photon between two output signals to generate a random stream of bits, however the theory behind it can be applied to other quantum processes as well.
+{[}@StefanovOptical{]} proposes using the random choice of a photon between two output signals to generate a random stream of bits, however the theory behind it can be applied to other quantum processes as well.
 This particular theorem has been implemented by Wayne et.
 al.
-\cite{Wayne} to create a quantum number generator.
+{[}@Wayne{]} to create a quantum number generator.
 While this article proves the efficacy of OQRNG, it utilizes a slightly different method.
 
-\subsubsection{2.1 Shot noise quantum fluctuations}\label{shot-noise-quantum-fluctuations}
+\hypertarget{shot-noise-quantum-fluctuations}{%
+\subsubsection{2.1 Shot noise quantum fluctuations}\label{shot-noise-quantum-fluctuations}}
 
@@
 al.
-\cite{StefanovOptical}, but instead using shot noise.
-As described by Niemczuk \cite{shotnoise}, shot noise is minor fluctuations in an electrical current, which is inherently random.
+{[}@StefanovOptical{]}, but instead using shot noise.
+As described by Niemczuk {[}@shotnoise{]}, shot noise is minor fluctuations in an electrical current, which is inherently random.
 Reading this property, then, gives us an intrinsically random source from which to generate a random output, which in turn can be processed into a random number.
@@
 al.
-\cite{contender1} presents an implementation using a fairly complex setup, in which a continuous-wave fiber laser is the optical source.
+{[}@contender1{]} presents an implementation using a fairly complex setup, in which a continuous-wave fiber laser is the optical source.
 They conclude that sampling the shot noise is, indeed, suitable for OQRNG.
@@
 al.
-\cite{singh}.
+{[}@singh{]}.
 This particular implementation uses a bespoke circuit board where all components are present on a single board -- e.g., this experimental setup contains an integrated ADC, post-processor, entropy controller and entropy generator.
@@
 al.
-\cite{contender1} and Singh et.
+{[}@contender1{]} and Singh et.
 al.
-\cite{singh} both implement OQRNG through readings of shot noise.
+{[}@singh{]} both implement OQRNG through readings of shot noise.
 However, there are limitations in both of these works.
-Either the system that generates the shot noise is large and complex \cite{contender1} or the system is built on bespoke hardware with limitations in processing power which prevents a fully integrated system \cite{singh}.
+Either the system that generates the shot noise is large and complex {[}@contender1{]} or the system is built on bespoke hardware with limitations in processing power which prevents a fully integrated system {[}@singh{]}.
 Furthermore, to the best of our knowledge, most of the work in this field is from the perspective of physicists, and there appears to be little research on this subject in the domain of computer science.
-Our work aims to bridge this gap by using commercially available hardware (\emph{other than the bespoke shot noise generator \cite{Clason2023}}) and focuses on implementing Toeplitz extraction directly on the microcontroller.
+Our work aims to bridge this gap by using commercially available hardware (\emph{other than the bespoke shot noise generator {[}@Clason2023{]}}) and focuses on implementing Toeplitz extraction directly on the microcontroller.
 Rather than focusing on the intricacies of quantum fluctuations, we will instead approach this problem from a computer science perspective.
 
-\subsection{3 BACKGROUND}\label{background}
+\hypertarget{background}{%
+\subsection{3 BACKGROUND}\label{background}}
 
-Our work is a practical continuation of the work of Clason \cite{Clason2023}.
+Our work is a practical continuation of the work of Clason {[}@Clason2023{]}.
 In this work, quantum shot noise originating from photodiodes was studied, and in so doing a prototype device was constructed.
@@
 
-\subsubsection{3.1 Optical RNG module}\label{optical-rng-module}
+\hypertarget{optical-rng-module}{%
+\subsubsection{3.1 Optical RNG module}\label{optical-rng-module}}
 
-The one bespoke piece of hardware used in this study is the prototype designed by Clason \cite{Clason2023} as a part of his masters thesis.
+The one bespoke piece of hardware used in this study is the prototype designed by Clason {[}@Clason2023{]} as a part of his masters thesis.
 This device produces the optical shot noise which will be the source of randomness in our work.
@@
 
-As described in Clasons work \cite{Clason2023}, the OQRNG-device is an electro-optical system which measures optical shot noise, generating quantum randomness.
+As described in Clasons work {[}@Clason2023{]}, the OQRNG-device is an electro-optical system which measures optical shot noise, generating quantum randomness.
 The device has an LED and a photodiode positioned a few millimeters apart, ensuring efficient light coupling.
@@
 
-Whereas the exact quantum mechanisms that ensure that this system ensures randomness and further details regarding the OQRNG-device is better derived directly from Clasons work \cite{Clason2023}, the primary concern for our study is the inherently random, analog voltage current produced by the system.
+Whereas the exact quantum mechanisms that ensure that this system ensures randomness and further details regarding the OQRNG-device is better derived directly from Clasons work {[}@Clason2023{]}, the primary concern for our study is the inherently random, analog voltage current produced by the system.
 
-\subsubsection{3.2 ADC converter}\label{adc-converter}
+\hypertarget{adc-converter}{%
+\subsubsection{3.2 ADC converter}\label{adc-converter}}
 
@@
 As mentioned in Section 1, the signal needs to pass through an ADC to be converted into raw bits.
-In his thesis, Clason \cite{Clason2023} suggests a discrete ADC chip capable of analyzing frequencies higher than 25 MHz, as this is the highest frequency studied in his work.
+In his thesis, Clason {[}@Clason2023{]} suggests a discrete ADC chip capable of analyzing frequencies higher than 25 MHz, as this is the highest frequency studied in his work.
 However, in the interest of keeping the implementation light and cheap, we will be using ADCs that provide less samples per second and lower frequencies.
@@
 SIMD-controlled DSP architectures, as described by Han et al.
-\cite{simd-dsp}, leverage parallel vectorized computation to accelerate matrix operations -- making them highly effective for Toeplitz matrix-vector multiplications.
+{[}@simd-dsp{]}, leverage parallel vectorized computation to accelerate matrix operations -- making them highly effective for Toeplitz matrix-vector multiplications.
 
@@
 
-\subsubsection{3.4 Toeplitz extraction}\label{toeplitz-extraction}
+\hypertarget{toeplitz-extraction}{%
+\subsubsection{3.4 Toeplitz extraction}\label{toeplitz-extraction}}
 
@@
 A detailed account of the inner workings of Toeplitz extraction can be found in the work of Chouhan et al.
-\cite{toeplitz-desc}.
+{[}@toeplitz-desc{]}.
 This work focuses on implementing Toeplitz extraction on field-programmable gate-arrays (FPGA), but some specific details can be derived from their work.
@@
 To summarize the theoretical working of Toeplitz extraction (\emph{as explained by Chouhan et al.
-\cite{toeplitz-desc}}), a pre-determined seed matrix (\(T\)) is multiplied with the sampled raw bit matrix (\(K\)).
+{[}@toeplitz-desc{]}}), a pre-determined seed matrix (\(T\)) is multiplied with the sampled raw bit matrix (\(K\)).
 The size of the seed is directly dependent on the size of the sampled data, and can be fixed or continually re-sampled as needed.
@@
 
-\subsection{4 RELATED WORKS}\label{related-works}
+\hypertarget{related-works}{%
+\subsection{4 RELATED WORKS}\label{related-works}}
 
@@
 As mentioned in section 3.4, Chouhan et al.
-\cite{toeplitz-desc} utilized FPGA to implement Toeplitz extraction specifically for OQRNG, a work also studied by Zhang et al.
-\cite{toeplitz}.
+{[}@toeplitz-desc{]} utilized FPGA to implement Toeplitz extraction specifically for OQRNG, a work also studied by Zhang et al.
+{[}@toeplitz{]}.
 Furthermore, Zhang et al.~utilized a standardized min-entropy evaluation to ensure true, unbiased randomness in their result.
 Both of these implementations utilize powerful hardware where the matrix multiplication is offloaded to FPGA.
-These implementations provided extraction speeds of between 3.36 Gpbs \cite{toeplitz} to 26.57 Gbps \cite{toeplitz-desc}.
+These implementations provided extraction speeds of between 3.36 Gpbs {[}@toeplitz{]} to 26.57 Gbps {[}@toeplitz-desc{]}.
 Neither of these implementations operate on constrained hardware, instead creating bespoke circuit boards for their works.
@@
 Numerous systems, such as those developed by Chouhan et al.
-\cite{toeplitz-desc} and Zhang et al.
-\cite{toeplitz}, depends on fixed seeds.
+{[}@toeplitz-desc{]} and Zhang et al.
+{[}@toeplitz{]}, depends on fixed seeds.
 Fixed seeds can create security risks over prolonged operation time.
-To tackle this problem, Lin et al \cite{lin} proposed a method for seed-renewable Toeplitz post-processing in QRNG.
+To tackle this problem, Lin et al {[}@lin{]} proposed a method for seed-renewable Toeplitz post-processing in QRNG.
 Their strategy incorporates a dynamic seed pool within the FPGA, where each instance of post-processing picks a new, randomly selected seed.
@@
 Liao et al.
-\cite{liao} showed that this process could be greatly accelerated using Fast Fourier Transform (\emph{FFT}) and its inverse (\emph{IFFT}) -- reducing computational complexity from \(O(n^2)\) to \(O(n \log n)\).
+{[}@liao{]} showed that this process could be greatly accelerated using Fast Fourier Transform (\emph{FFT}) and its inverse (\emph{IFFT}) -- reducing computational complexity from \(O(n^2)\) to \(O(n \log n)\).
 Their implementation on FPGA utilized this approach for deep neural networks, resulting in a 28.7 times decrease in model size while still achieving fast inference speeds.
@@
 
-\subsubsection{5.1 Phase one}\label{phase-one}
+\hypertarget{phase-one}{%
+\subsubsection{5.1 Phase one}\label{phase-one}}
 
@@
 
-\subsubsection{5.2 Phase two}\label{phase-two}
+\hypertarget{phase-two}{%
+\subsubsection{5.2 Phase two}\label{phase-two}}
 
@@
 
-\subsubsection{5.3 Evaluation}\label{evaluation}
+\hypertarget{evaluation}{%
+\subsubsection{5.3 Evaluation}\label{evaluation}}
 
@@
 al.
-\cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements of the execution speed of an algorithm.
+{[}@micromeasurements{]} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements of the execution speed of an algorithm.
 An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (as the process for this may vary greatly between controllers).
@@
 
-\subsubsection{5.4 Limitations}\label{limitations}
+\hypertarget{limitations}{%
+\subsubsection{5.4 Limitations}\label{limitations}}
 
@@
 
-\subsection{6 RESULTS}\label{results}
+\hypertarget{results}{%
+\subsection{6 RESULTS}\label{results}}
 
@@
 
-\subsubsection{6.1 Phase one}\label{phase-one-1}
+\hypertarget{phase-one-1}{%
+\subsubsection{6.1 Phase one}\label{phase-one-1}}
 
@@
 
-\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
-\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
-\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 13.1564 & 106.3914
-\\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580 & 21111.2163 \\ \hline
-\caption{A brief overview of the testscript architecture.} \label{tab:iter1}
-
-\end{tabular}
+\begin{table}[ht] \centering \begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
+\textbf{Teensy ($\mu$s)} & \textbf{Pico ($\mu$s)} \\ \hline 64 & 13.1564 &
+106.3914 \\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580 & 21111.2163 \\
+\hline \end{tabular} \caption{Execution times for different bit sizes on Teensy
+and Pico} \label{tab:iter1} \end{table}
 
@@
 
-\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
+\begin{table}[ht]\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
 \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline 64 & 16.4689 \\ 512 &
 1006.6255 \\ 1024 & 3996.5972 \\ \hline \end{tabular}
-
-\label{tab:iter2}
+\label{tab:iter2}\end{table}
 
@@
 
-\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
+\begin{table}[ht]\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
 \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline 64/64 & 43.0760 \\ 512/512
 & 2663.8194 \\ 1024/1024 & 10513.1767 \\ \hline \end{tabular}
-
-\label{tab:iter3}
+\label{tab:iter3}\end{table}
 
 Table \ref{tab:iter3} presents the average execution time of iteration 3 on the Teensy 4.1 for varying batching and bit sizes.
-Additionally, when performing the same 1024/1024-bit operation, the Raspberry Pi Pico 2 showed average execution time of \texttt{80806.006\ µs} indicating a substantially lower throughput compared to the Teensy.
+Additionally, when performing the same 1024/1024-bit operation, the Raspberry Pi Pico 2 showed average execution time of \texttt{80806.006\ \$\$\textbackslash{}mu\textbackslash{}mathrm\{s\}\$} indicating a substantially lower throughput compared to the Teensy.
 
-\subsubsection{6.2 Phase two}\label{phase-two-1}
+\hypertarget{phase-two-1}{%
+\subsubsection{6.2 Phase two}\label{phase-two-1}}
 
@@
 
-\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
+\begin{table}[ht]\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
 \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
@@
 512 & 551.6358 & 3978.1928 \\ 1024 & 2195.1979 & 15830.8784 \\ \hline
-\end{tabular}
-
-\label{tab:iter4}
+\end{tabular} \label{tab:iter4}\end{table}
 
@@
 
-\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
+\begin{table}[ht]\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
 \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
 \multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 0.0501 & 0.2175 \\
-\hline \end{tabular}
-
-\label{tab:iter5}
+\hline \end{tabular} \label{tab:iter5}\end{table}
 
@@
 
-\begin{tabular}{|c|c|} \hline \textbf{Data structure} &
+\begin{table}[ht]\begin{tabular}{|c|c|} \hline \textbf{Data structure} &
 \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline array & 0.4284 \\
 unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline \end{tabular}
-
-\label{tab:iter6}
+\label{tab:iter6}\end{table}
 
@@
 
-\subsection{7 CONCLUSION}\label{conclusion}
+\hypertarget{conclusion}{%
+\subsection{7 CONCLUSION}\label{conclusion}}
 
-\subsection{CHANGELOG}\label{changelog}
+\hypertarget{changelog}{%
+\subsection{CHANGELOG}\label{changelog}}
 
@@
 2025-05-20: Unfortunately, we've missed adding concrete details about changes made up until this point. Fix later.
-% BALANCE COLUMNS
-\balance{}
-
-% REFERENCES FORMAT
-% References must be the same font size as other body text.
-\bibliographystyle{SIGCHI-Reference-Format}
-\bibliography{sample}
-
-\end{document}
-
-%%% Local Variables:
-%%% mode: latex
-%%% TeX-master: t
-%%% End:

### 2025-05-21 docs/style/article-body.tex
microsec fix

@@
 
-\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
+\begin{table}[ht]\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
 \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline 64/64 & 43.0760 \\ 512/512
 & 2663.8194 \\ 1024/1024 & 10513.1767 \\ \hline \end{tabular}
-
-\label{tab:iter3}
+\label{tab:iter3}\end{table}
 
@@
 
-\subsubsection{6.2 Phase two}\label{phase-two-1}
+\hypertarget{phase-two-1}{%
+\subsubsection{6.2 Phase two}\label{phase-two-1}}
 

### 2025-05-21 docs/style/article-body.tex

--- 2025-05-21 docs/style/article-body.tex
@@
 
-\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
-\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
-\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 13.1564 & 106.3914
-\\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580 & 21111.2163 \\ \hline
+\textbackslash begin\{table\}{[}ht{]}begin\{tabular\}\{\textbar c\textbar c\textbar c\textbar\} \hline \textbf{Bit size} \& \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} \& \multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \textbackslash{} \hline 64 \& 13.1564 \& 106.3914 \textbackslash{} 512 \& 788.3139 \& 5302.4979 \textbackslash{} 1024 \& 3124.0580 \& 21111.2163 \textbackslash{} \hline
+
 \caption{A brief overview of the testscript architecture.} \label{tab:iter1}
 
-\end{tabular}
+\textbackslash end\{tabular\} \label{tab:iter1}/end\{table\}
 

### 2025-05-21 docs/style/article-body.tex
small fix

@@
 
-\textbackslash begin\{table\}{[}ht{]}begin\{tabular\}\{\textbar c\textbar c\textbar c\textbar\} \hline \textbf{Bit size} \& \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} \& \multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \textbackslash{} \hline 64 \& 13.1564 \& 106.3914 \textbackslash{} 512 \& 788.3139 \& 5302.4979 \textbackslash{} 1024 \& 3124.0580 \& 21111.2163 \textbackslash{} \hline
+\begin{table}[ht] \centering \caption{Execution times for different bit sizes on
+Teensy and Pico} \label{tab:iter1} \begin{tabular}{|c|c|c|} \hline \textbf{Bit
+size} & \textbf{Teensy ($\mu$s)} & \textbf{Pico ($\mu$s)} \\ \hline 64 & 13.1564
+& 106.3914 \\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580 & 21111.2163 \\
+\hline \end{tabular} \end{table}
 
-\caption{A brief overview of the testscript architecture.} \label{tab:iter1}
-
-\textbackslash end\{tabular\} \label{tab:iter1}/end\{table\}
-
-Table \ref{tab:iter1} presents the average execution time of iteration 1 on both Teensy 4.1 and Raspberry Pico Pi 2 across thre input sizes.
+Table \ref{tab:iter1} presents the average execution time of iteration 1 on both Teensy 4.1 and Raspberry Pico Pi 2 across the input sizes.
 The Teensy consistently outperforms the Pico, with the gap widening as the bit size increases.
@@
 Additionally, the isolated effect of applying a single bitmask operation \texttt{\&\ 1} was evaluated.
-This resulted in a slight reduction in average execution time, from \texttt{13.1564\ µs} to \texttt{13.1000\ µs}, although this specific result is not included in the tables above.
+This resulted in a slight reduction in average execution time, from \texttt{13.1564\ \$\textbackslash{}mu\ s\$} to \texttt{13.1000\ \$\textbackslash{}mu\ s\$}, although this specific result is not included in the tables above.
 
@@
 Table \ref{tab:iter3} presents the average execution time of iteration 3 on the Teensy 4.1 for varying batching and bit sizes.
-Additionally, when performing the same 1024/1024-bit operation, the Raspberry Pi Pico 2 showed average execution time of \texttt{80806.006\ µs} indicating a substantially lower throughput compared to the Teensy.
+Additionally, when performing the same 1024/1024-bit operation, the Raspberry Pi Pico 2 showed average execution time of \texttt{80806.006\ \$\textbackslash{}mu\ s\$} indicating a substantially lower throughput compared to the Teensy.
 
@@
 
-In addition to the results presented in Table \ref{tab:iter4}, a further tests was conducted on the Teensy using a single-loop unrolled implementation for the 64-bit size, resulting in an average execution time of \texttt{6.6626\ µs}.
-An additional fully unrolled variant, where loops were entirely eliminated, produced a measured execution time of \texttt{0.0491\ µs}.
+In addition to the results presented in Table \ref{tab:iter4}, a further tests was conducted on the Teensy using a single-loop unrolled implementation for the 64-bit size, resulting in an average execution time of \texttt{6.6626\ \$\textbackslash{}mu\ s\$}.
+An additional fully unrolled variant, where loops were entirely eliminated, produced a measured execution time of \texttt{0.0491\ \$\textbackslash{}mu\ s\$}.
 However, this measurement was later determined to be invalid due to packaging error.

### 2025-05-21 docs/style/article-body.tex
table position fix

--- 2025-05-21 docs/style/article-body.tex
@@
 
-\begin{figure*}[ht] \centering
+\begin{figure*}[!t] \centering
 \includegraphics[width=\textwidth]{img/testscript.png} \caption{A brief overview
@@
 
-\textbf{Iteration 1 - Data Structures:}
+\vspace{1em}
 
-\begin{table}[ht] \centering \begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
-\textbf{Teensy ($\mu$s)} & \textbf{Pico ($\mu$s)} \\ \hline 64 & 13.1564 &
-106.3914 \\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580 & 21111.2163 \\
-\hline \end{tabular} \caption{Execution times for different bit sizes on Teensy
-and Pico} \label{tab:iter1} \end{table}
+\begin{table}[ht] \centering \caption{Iteration 1 - Data Structures}
+\label{tab:iter1}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Bit size} & \textbf{Teensy ($\mu$s)} & \textbf{Pico ($\mu$s)} \\
+\hline 64 & 13.1564 & 106.3914 \\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580
+& 21111.2163 \\ \hline \end{tabularx} \end{table}
 
-Table \ref{tab:iter1} presents the average execution time of iteration 1 on both Teensy 4.1 and Raspberry Pico Pi 2 across thre input sizes.
+Table \ref{tab:iter1} presents the average execution time of iteration 1 on both Teensy 4.1 and Raspberry Pico Pi 2 across the input sizes.
 The Teensy consistently outperforms the Pico, with the gap widening as the bit size increases.
 
-\textbf{Iteration 2 - Bitshifting:}
-
-\begin{table}[ht]\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
-\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline 64 & 16.4689 \\ 512 &
-1006.6255 \\ 1024 & 3996.5972 \\ \hline \end{tabular}
-\label{tab:iter2}\end{table}
+\begin{table}[ht] \centring \caption{Iteration 2 - Bitshifting}
+\label{tab:iter2}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline
+64 & 16.4689 \\ 512 & 1006.6255 \\ 1024 & 3996.5972 \\ \hline \end{tabularx}
+\end{table}
 
@@
 Additionally, the isolated effect of applying a single bitmask operation \texttt{\&\ 1} was evaluated.
-This resulted in a slight reduction in average execution time, from \texttt{13.1564\ µs} to \texttt{13.1000\ µs}, although this specific result is not included in the tables above.
+This resulted in a slight reduction in average execution time, from \texttt{13.1564} \(\mu s\) to \texttt{13.1000} \(\mu s\), although this specific result is not included in the tables above.
 
-\textbf{Iteration 3 - Batching and Hardware optimization:}
-
-\begin{table}[ht]\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
-\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline 64/64 & 43.0760 \\ 512/512
-& 2663.8194 \\ 1024/1024 & 10513.1767 \\ \hline \end{tabular}
-\label{tab:iter3}\end{table}
+\begin{table}[H] \centring \caption{Iteration 3 - Batching and Hardware
+optimization} \label{tab:iter3}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline
+64/64 & 43.0760 \\ 512/512 & 2663.8194 \\ 1024/1024 & 10513.1767 \\ \hline
+\end{tabularx} \end{table}
 
 Table \ref{tab:iter3} presents the average execution time of iteration 3 on the Teensy 4.1 for varying batching and bit sizes.
-Additionally, when performing the same 1024/1024-bit operation, the Raspberry Pi Pico 2 showed average execution time of \texttt{80806.006\ µs} indicating a substantially lower throughput compared to the Teensy.
+Additionally, when performing the same 1024/1024-bit operation, the Raspberry Pi Pico 2 showed average execution time of \texttt{80806.006} \(\mu s\) indicating a substantially lower throughput compared to the Teensy.
 
@@
 
-\textbf{Iteration 4 - Loop unrolling:}
+\vspace{1em}
 
-\begin{table}[ht]\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
-\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
+\begin{table}[ht] \centring \caption{Iteration 4 - Loop unrolling}
+\label{tab:iter4}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
 \multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 9.7017 & 70.7402 \\
 512 & 551.6358 & 3978.1928 \\ 1024 & 2195.1979 & 15830.8784 \\ \hline
-\end{tabular} \label{tab:iter4}\end{table}
+\end{tabularx} \end{table}
 
@@
 
-In addition to the results presented in Table \ref{tab:iter4}, a further tests was conducted on the Teensy using a single-loop unrolled implementation for the 64-bit size, resulting in an average execution time of \texttt{6.6626\ µs}.
-An additional fully unrolled variant, where loops were entirely eliminated, produced a measured execution time of \texttt{0.0491\ µs}.
+In addition to the results presented in Table \ref{tab:iter4}, a further tests was conducted on the Teensy using a single-loop unrolled implementation for the 64-bit size, resulting in an average execution time of \texttt{6.6626} \(\mu s\).
+An additional fully unrolled variant, where loops were entirely eliminated, produced a measured execution time of \texttt{0.0491} \(\mu s\).
 However, this measurement was later determined to be invalid due to packaging error.
 
-\textbf{Iteration 5 - Removal of vector usage:}
+\vspace{1em}
 
-\begin{table}[ht]\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
-\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
+\begin{table}[ht] \centring \caption{Iteration 5 - Removal of vector usage}
+\label{tab:iter5}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
 \multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 0.0501 & 0.2175 \\
-\hline \end{tabular} \label{tab:iter5}\end{table}
+\hline \end{tabularx} \end{table}
 
@@
 
-\textbf{Iteration 6 - Data type exploration:}
+\vspace{1em}
 
-\begin{table}[ht]\begin{tabular}{|c|c|} \hline \textbf{Data structure} &
-\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline array & 0.4284 \\
-unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline \end{tabular}
-\label{tab:iter6}\end{table}
+\begin{table}[H] \centring \caption{Iteration 6 - Data type exloration}
+\label{tab:iter6}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Data structure} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\
+\hline array & 0.4284 \\ unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline
+\end{tabularx} \end{table}
 
-Table \ref{tab:iter6} presents the results of iteration 6, which was introduced to address the structural limitations encountered in iteration 5.
+Table \ref{tab:iter6} presents the results of iteration 6 using 64 bit size, which was introduced to address the structural limitations encountered in iteration 5.
 Specifically, the fixed-width output constrains imposed by the removal of vector.
@@
 Vector-based approach from iteration 1 serving as a baseline for comparison.
+
+Additionally, 128 bit size were tested resulting in \texttt{179.8712} \(\mu s\).
 

### 2025-05-21 docs/style/article-final.tex
table position fix

--- 2025-05-21 docs/style/article-final.tex
@@
 \usepackage{algorithmic}
-
+\usepackage{caption}
+\usepackage{tabularx}
+\usepackage{array}
+\usepackage{float}
 % Some optional stuff you might like/need.
@@
 
-\subsubsection{6.1 Phase one}\label{phase-one-1}
+\hypertarget{phase-one-1}{%
+\subsubsection{6.1 Phase one}\label{phase-one-1}}
 
-\textbf{Iteration 1 - Data Structures:}
+\vspace{1em}
 
-\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
-\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
-\multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 13.1564 & 106.3914
-\\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580 & 21111.2163 \\ \hline
-\caption{A brief overview of the testscript architecture.} \label{tab:iter1}
+\begin{table}[ht] \centering \caption{Iteration 1 - Data Structures}
+\label{tab:iter1}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Bit size} & \textbf{Teensy ($\mu$s)} & \textbf{Pico ($\mu$s)} \\
+\hline 64 & 13.1564 & 106.3914 \\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580
+& 21111.2163 \\ \hline \end{tabularx} \end{table}
 
-\end{tabular}
-
-Table \ref{tab:iter1} presents the average execution time of iteration 1 on both Teensy 4.1 and Raspberry Pico Pi 2 across thre input sizes.
+Table \ref{tab:iter1} presents the average execution time of iteration 1 on both Teensy 4.1 and Raspberry Pico Pi 2 across the input sizes.
 The Teensy consistently outperforms the Pico, with the gap widening as the bit size increases.
 
-\textbf{Iteration 2 - Bitshifting:}
-
-\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
-\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline 64 & 16.4689 \\ 512 &
-1006.6255 \\ 1024 & 3996.5972 \\ \hline \end{tabular}
-
+\begin{table}[ht] \centring \caption{Iteration 2 - Bitshifting}
 \label{tab:iter2}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline
+64 & 16.4689 \\ 512 & 1006.6255 \\ 1024 & 3996.5972 \\ \hline \end{tabularx}
+\end{table}
 
@@
 Additionally, the isolated effect of applying a single bitmask operation \texttt{\&\ 1} was evaluated.
-This resulted in a slight reduction in average execution time, from \texttt{13.1564\ µs} to \texttt{13.1000\ µs}, although this specific result is not included in the tables above.
+This resulted in a slight reduction in average execution time, from \texttt{13.1564} \(\mu s\) to \texttt{13.1000} \(\mu s\), although this specific result is not included in the tables above.
 
-\textbf{Iteration 3 - Batching and Hardware optimization:}
-
-\begin{tabular}{|c|c|} \hline \textbf{Bit size} &
-\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline 64/64 & 43.0760 \\ 512/512
-& 2663.8194 \\ 1024/1024 & 10513.1767 \\ \hline \end{tabular}
-
-\label{tab:iter3}
+\begin{table}[H] \centring \caption{Iteration 3 - Batching and Hardware
+optimization} \label{tab:iter3}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline
+64/64 & 43.0760 \\ 512/512 & 2663.8194 \\ 1024/1024 & 10513.1767 \\ \hline
+\end{tabularx} \end{table}
 
 Table \ref{tab:iter3} presents the average execution time of iteration 3 on the Teensy 4.1 for varying batching and bit sizes.
-Additionally, when performing the same 1024/1024-bit operation, the Raspberry Pi Pico 2 showed average execution time of \texttt{80806.006\ µs} indicating a substantially lower throughput compared to the Teensy.
+Additionally, when performing the same 1024/1024-bit operation, the Raspberry Pi Pico 2 showed average execution time of \texttt{80806.006} \(\mu s\) indicating a substantially lower throughput compared to the Teensy.
 
-\subsubsection{6.2 Phase two}\label{phase-two-1}
+\hypertarget{phase-two-1}{%
+\subsubsection{6.2 Phase two}\label{phase-two-1}}
 
-\textbf{Iteration 4 - Loop unrolling:}
+\vspace{1em}
 
-\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
-\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
+\begin{table}[ht] \centring \caption{Iteration 4 - Loop unrolling}
+\label{tab:iter4}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
 \multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 9.7017 & 70.7402 \\
 512 & 551.6358 & 3978.1928 \\ 1024 & 2195.1979 & 15830.8784 \\ \hline
-\end{tabular}
-
-\label{tab:iter4}
+\end{tabularx} \end{table}
 
@@
 
-In addition to the results presented in Table \ref{tab:iter4}, a further tests was conducted on the Teensy using a single-loop unrolled implementation for the 64-bit size, resulting in an average execution time of \texttt{6.6626\ µs}.
-An additional fully unrolled variant, where loops were entirely eliminated, produced a measured execution time of \texttt{0.0491\ µs}.
+In addition to the results presented in Table \ref{tab:iter4}, a further tests was conducted on the Teensy using a single-loop unrolled implementation for the 64-bit size, resulting in an average execution time of \texttt{6.6626} \(\mu s\).
+An additional fully unrolled variant, where loops were entirely eliminated, produced a measured execution time of \texttt{0.0491} \(\mu s\).
 However, this measurement was later determined to be invalid due to packaging error.
 
-\textbf{Iteration 5 - Removal of vector usage:}
+\vspace{1em}
 
-\begin{tabular}{|c|c|c|} \hline \textbf{Bit size} &
-\multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
+\begin{table}[ht] \centring \caption{Iteration 5 - Removal of vector usage}
+\label{tab:iter5}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} &
 \multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 0.0501 & 0.2175 \\
-\hline \end{tabular}
-
-\label{tab:iter5}
+\hline \end{tabularx} \end{table}
 
@@
 
-\textbf{Iteration 6 - Data type exploration:}
+\vspace{1em}
 
-\begin{tabular}{|c|c|} \hline \textbf{Data structure} &
-\multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline array & 0.4284 \\
-unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline \end{tabular}
+\begin{table}[H] \centring \caption{Iteration 6 - Data type exloration}
+\label{tab:iter6}
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Data structure} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\
+\hline array & 0.4284 \\ unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline
+\end{tabularx} \end{table}
 
-\label{tab:iter6}
-
-Table \ref{tab:iter6} presents the results of iteration 6, which was introduced to address the structural limitations encountered in iteration 5.
+Table \ref{tab:iter6} presents the results of iteration 6 using 64 bit size, which was introduced to address the structural limitations encountered in iteration 5.
 Specifically, the fixed-width output constrains imposed by the removal of vector.
@@
 
-\subsection{7 CONCLUSION}\label{conclusion}
+Additionally, 128 bit size were tested resulting in \texttt{179.8712} \(\mu s\).
 
-\subsection{CHANGELOG}\label{changelog}
+\hypertarget{conclusion}{%
+\subsection{7 CONCLUSION}\label{conclusion}}
+
+\hypertarget{changelog}{%
+\subsection{CHANGELOG}\label{changelog}}
 

### 2025-05-21 docs/style/article-body.tex
Perhaps the last one before seminar submission

--- 2025-05-21 docs/style/article-body.tex
@@
 
+Observing the results, we can place them into the context of the limits imposed by the ADC, as discussed in Section 3.2.
+In (3), we calculate the average execution speed required to be \(2.667 \mu\) for the ADC that requires soldering to the microcontroller.
+From Tables \ref{tab:iter5} and \ref{tab:iter6} we demonstrate execution speeds well below that, with iteration 5 as well as iteration 6 -- specifically with bitsets -- being the most prominent results to discuss.
+Here, our implementation achieved speeds of \(0.0501 \mu s\) and \(0.0474 \mu\) respectively which within the bounds of the ADC listed in Section 3.2.
+The built in ADC on Teensy has an effective sample rate of 1 MSPS (\emph{e.g.
+half of MAX11102AUB}), which would require an average execution speed of \textasciitilde{}\(2.667/2 \mu s \approx 1.334 \mu s\), making our implementation suitable even for that limited conversion speed.
+
+The main culprit that led other iterations to not be viable was the inclusion of \texttt{std::vector}, which was used to ensure we could deliver variable lengths of our return data -- e.g.~either 32, 64 or 128 bits.
+The overhead introduced from initializing and populating the vector led to significant delays in execution speed, leading to all other attempts at optimization to be superfluous.
+In fact, any other complexities led to significantly worse results than the naive, original implementation.
+
+However, it is important to note that both of these implementations has a hard limit of working on 64 bits input.
+Whereas larger bitstrings still yield results with high entropy, the execution speed increases far above the hardware limitations, as evident in \ref{tab:iter6}.
+Iteration 5 is hard capped at using only 64 bits input, as we need to store the bits in fixed-size integers before running them through the algorithm, and on embedded machines we have no larger integer types available.
+
+One issue still remains concerning the size of the input string and subsequent output string.
+The larger the size of the matrix provided to the Toeplitz function (\emph{e.g.~the larger our input}), the higher the potential overall entropy becomes.
+As our implementation is essentially capped at 64 bits input at this point in time (\emph{without being limited by the hardware}), this may or may not prove to not be secure enough for use in real world applications.
+This, however, is out of scope for this particular thesis, but a consideration nonetheless.
+
 \subsection{CHANGELOG}\label{changelog}

### 2025-05-21 docs/style/article-body.tex

--- 2025-05-21 docs/style/article-body.tex
@@
-\hypertarget{introduction}{%
-\section{1 INTRODUCTION}\label{introduction}}
+\section{1 INTRODUCTION}\label{introduction}
 
@@
 
-\hypertarget{theory}{%
-\subsection{2 THEORY}\label{theory}}
+\subsection{2 THEORY}\label{theory}
 
@@
 
-\hypertarget{shot-noise-quantum-fluctuations}{%
-\subsubsection{2.1 Shot noise quantum fluctuations}\label{shot-noise-quantum-fluctuations}}
+\subsubsection{2.1 Shot noise quantum fluctuations}\label{shot-noise-quantum-fluctuations}
 
@@
 
-\hypertarget{background}{%
-\subsection{3 BACKGROUND}\label{background}}
+\subsection{3 BACKGROUND}\label{background}
 
@@
 
-\hypertarget{optical-rng-module}{%
-\subsubsection{3.1 Optical RNG module}\label{optical-rng-module}}
+\subsubsection{3.1 Optical RNG module}\label{optical-rng-module}
 
@@
 
-\hypertarget{adc-converter}{%
-\subsubsection{3.2 ADC converter}\label{adc-converter}}
+\subsubsection{3.2 ADC converter}\label{adc-converter}
 
@@
 
-\hypertarget{microcontroller}{%
-\subsubsection{3.3 Microcontroller}\label{microcontroller}}
+\subsubsection{3.3 Microcontroller}\label{microcontroller}
 
@@
 
-\hypertarget{toeplitz-extraction}{%
-\subsubsection{3.4 Toeplitz extraction}\label{toeplitz-extraction}}
+\subsubsection{3.4 Toeplitz extraction}\label{toeplitz-extraction}
 
@@
 
-\hypertarget{summary}{%
-\subsubsection{3.5 Summary}\label{summary}}
+\subsubsection{3.5 Summary}\label{summary}
 
@@
 
-\hypertarget{related-works}{%
-\subsection{4 RELATED WORKS}\label{related-works}}
+\subsection{4 RELATED WORKS}\label{related-works}
 
@@
 
-\hypertarget{methodology}{%
-\subsection{5 METHODOLOGY}\label{methodology}}
+\subsection{5 METHODOLOGY}\label{methodology}
 
@@
 
-\hypertarget{phase-one}{%
-\subsubsection{5.1 Phase one}\label{phase-one}}
+\subsubsection{5.1 Phase one}\label{phase-one}
 
@@
 
-\hypertarget{phase-two}{%
-\subsubsection{5.2 Phase two}\label{phase-two}}
+\subsubsection{5.2 Phase two}\label{phase-two}
 
@@
 
-\hypertarget{evaluation}{%
-\subsubsection{5.3 Evaluation}\label{evaluation}}
+\subsubsection{5.3 Evaluation}\label{evaluation}
 
@@
 
-\hypertarget{limitations}{%
-\subsubsection{5.4 Limitations}\label{limitations}}
+\subsubsection{5.4 Limitations}\label{limitations}
 
@@
 
-\hypertarget{results}{%
-\subsection{6 RESULTS}\label{results}}
+\subsection{6 RESULTS}\label{results}
 
@@
 
-\hypertarget{phase-one-1}{%
-\subsubsection{6.1 Phase one}\label{phase-one-1}}
+\subsubsection{6.1 Phase one}\label{phase-one-1}
 
@@
 
-\hypertarget{phase-two-1}{%
-\subsubsection{6.2 Phase two}\label{phase-two-1}}
+\subsubsection{6.2 Phase two}\label{phase-two-1}
 
@@
 
-\hypertarget{conclusion}{%
-\subsection{7 CONCLUSION}\label{conclusion}}
+\subsection{7 CONCLUSION}\label{conclusion}
 
-\hypertarget{changelog}{%
-\subsection{CHANGELOG}\label{changelog}}
+Observing the results, we can place them into the context of the limits imposed by the ADC, as discussed in Section 3.2.
+In (3), we calculate the average execution speed required to be \(2.667 \mu\) for the ADC that requires soldering to the microcontroller.
+From Tables \ref{tab:iter5} and \ref{tab:iter6} we demonstrate execution speeds well below that, with iteration 5 as well as iteration 6 -- specifically with bitsets -- being the most prominent results to discuss.
+Here, our implementation achieved speeds of \(0.0501 \mu s\) and \(0.0474 \mu\) respectively which within the bounds of the ADC listed in Section 3.2.
+The built in ADC on Teensy has an effective sample rate of 1 MSPS (\emph{e.g.
+half of MAX11102AUB}), which would require an average execution speed of \textasciitilde{}\(2.667/2 \mu s \approx 1.334 \mu s\), making our implementation suitable even for that limited conversion speed.
+
+The main culprit that led other iterations to not be viable was the inclusion of \texttt{std::vector}, which was used to ensure we could deliver variable lengths of our return data -- e.g.~either 32, 64 or 128 bits.
+The overhead introduced from initializing and populating the vector led to significant delays in execution speed, leading to all other attempts at optimization to be superfluous.
+In fact, any other complexities led to significantly worse results than the naive, original implementation.
+
+However, it is important to note that both of these implementations has a hard limit of working on 64 bits input.
+Whereas larger bitstrings still yield results with high entropy, the execution speed increases far above the hardware limitations, as evident in \ref{tab:iter6}.
+Iteration 5 is hard capped at using only 64 bits input, as we need to store the bits in fixed-size integers before running them through the algorithm, and on embedded machines we have no larger integer types available.
+
+One issue still remains concerning the size of the input string and subsequent output string.
+The larger the size of the matrix provided to the Toeplitz function (\emph{e.g.~the larger our input}), the higher the potential overall entropy becomes.
+As our implementation is essentially capped at 64 bits input at this point in time (\emph{without being limited by the hardware}), this may or may not prove to not be secure enough for use in real world applications.
+This, however, is out of scope for this particular thesis, but a consideration nonetheless.
+
+\subsection{CHANGELOG}\label{changelog}
 

### 2025-05-21 docs/style/article-final.tex

--- 2025-05-21 docs/style/article-final.tex
@@
 
-\hypertarget{introduction}{%
-\section{1 INTRODUCTION}\label{introduction}}
+\section{1 INTRODUCTION}\label{introduction}
 

### 2025-06-02 docs/style/article-body.tex
tabel fix, added MHz for MCU and some small changes

--- 2025-05-21 docs/style/article-body.tex
@@
 
-In order to evaluate how efficent our implementation can become, our aim is to try our implementation on other MCUs with varying levels of power and hardware support.
-Whereas Teensy 4.1 is our primary development platform which we will evaluate closely, we aim to run our implementations on Raspberry Pi Pico 2\footnote{\href{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}{Raspberry Pi Pico 2 documentation, accessed 2025-03-13}} as well as ESP32-S3\footnote{\href{https://www.espressif.com/sites/default/files/documentation/esp32-s3_datasheet_en.pdf}{ESP32-S3 documentation, accessed 2025-03-13}}.
-Due to the lower computational power of these MCUs, there may be significant issues in utilizing these weaker models, yet they are significantly cheaper and easier to access.
-Testing of these will consist solely of running the implementation on these controllers and measuring execution speed and correctness of the output.
+In order to evaluate how efficent our implementation can become, our aim is to try our implementation on other MCU with different level of power and hardware support.
+Whereas Teensy 4.1, operating at \texttt{600\ MHz}, is our primary development platform which we will evaluate closely, we aim to run our implementations on the Raspberry Pi Pico 2\footnote{\href{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}{Raspberry Pi Pico 2 documentation, accessed 2025-03-13}}, operating at \texttt{150\ MHz}.
+Due to the lower computational power of this MCU, there may be significant issues in utilizing this weaker model, yet it is significantly cheaper and easier to access.
+Testing will consist solely of running the implementation on these controllers and measuring execution speed and correctness of the output.
 
@@
 An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (as the process for this may vary greatly between controllers).
+
+In our implementation, execution time for each extraction was measured using Arduino's built-in \texttt{micros()}\footnote{\href{https://docs.arduino.cc/language-reference/en/functions/time/micros/}{Documentation for \texttt{micros()}, accessed 2025-05-02.}} function.
+This function offers a convenient and platform-independent method for recording elapsed time with microsecond precision, enabling consistent performance evaluation across different iterations.
+
 As we will use fixed-size bitstrings for evaluation, we can then derive the throughput of the algorithm in \(Mbit/s\) as follows:
@@
 
-\begin{table}[ht] \centering \caption{Iteration 1 - Data Structures}
-\label{tab:iter1}
+\begin{table}[ht] \centring
 \begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
@@
 \hline 64 & 13.1564 & 106.3914 \\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580
-& 21111.2163 \\ \hline \end{tabularx} \end{table}
+& 21111.2163 \\ \hline \end{tabularx} \caption{Iteration 1 - Data Structures}
+\label{tab:iter1} \end{table}
 
@@
 
-\begin{table}[ht] \centring \caption{Iteration 2 - Bitshifting}
-\label{tab:iter2}
+\begin{table}[ht] \centring
 \begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
-\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline
-64 & 16.4689 \\ 512 & 1006.6255 \\ 1024 & 3996.5972 \\ \hline \end{tabularx}
-\end{table}
+\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} \\
+\hline 64 & 16.4689 \\ 512 & 1006.6255 \\ 1024 & 3996.5972 \\ \hline
+\end{tabularx} \caption{Iteration 2 - Bitshifting} \label{tab:iter2} \end{table}
 
@@
 
-\begin{table}[H] \centring \caption{Iteration 3 - Batching and Hardware
-optimization} \label{tab:iter3}
+\begin{table}[H] \centring
 \begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
-\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\ \hline
-64/64 & 43.0760 \\ 512/512 & 2663.8194 \\ 1024/1024 & 10513.1767 \\ \hline
-\end{tabularx} \end{table}
+\hline \textbf{Bit size} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}} \\
+\hline 64/64 & 43.0760 \\ 512/512 & 2663.8194 \\ 1024/1024 & 10513.1767 \\
+\hline \end{tabularx} \caption{Iteration 3 - Batching and Hardware optimization}
+\label{tab:iter3} \end{table}
 
@@
 
-\begin{table}[ht] \centring \caption{Iteration 4 - Loop unrolling}
-\label{tab:iter4}
+\begin{table}[ht] \centring
 \begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
@@
 512 & 551.6358 & 3978.1928 \\ 1024 & 2195.1979 & 15830.8784 \\ \hline
-\end{tabularx} \end{table}
+\end{tabularx} \caption{Iteration 4 - Loop unrolling} \label{tab:iter4}
+\end{table}
 
@@
 
-\begin{table}[ht] \centring \caption{Iteration 5 - Removal of vector usage}
-\label{tab:iter5}
+\begin{table}[ht] \centring
 \begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
@@
 \multicolumn{1}{c|}{\textbf{Pico ($\mu s$)}} \\ \hline 64 & 0.0501 & 0.2175 \\
-\hline \end{tabularx} \end{table}
+\hline \end{tabularx} \caption{Iteration 5 - Removal of vector usage}
+\label{tab:iter5} \end{table}
 
@@
 
-\begin{table}[H] \centring \caption{Iteration 6 - Data type exloration}
-\label{tab:iter6}
+\begin{table}[H] \centring
 \begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
-\hline \textbf{Data structure} & \multicolumn{1}{c|}{\textbf{Avg ($\mu s$)}} \\
-\hline array & 0.4284 \\ unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline
-\end{tabularx} \end{table}
+\hline \textbf{Data structure} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}}
+\\ \hline array & 0.4284 \\ unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline
+\end{tabularx} \caption{Iteration 6 - Data type exloration} \label{tab:iter6}
+\end{table}
 
@@
 \subsection{7 CONCLUSION}\label{conclusion}}
+
+Observing the results, we can place them into the context of the limits imposed by the ADC, as discussed in Section 3.2.
+In (3), we calculate the average execution speed required to be \(2.667 \mu\) for the ADC that requires soldering to the microcontroller.
+From Tables \ref{tab:iter5} and \ref{tab:iter6} we demonstrate execution speeds well below that, with iteration 5 as well as iteration 6 -- specifically with bitsets -- being the most prominent results to discuss.
+Here, our implementation achieved speeds of \(0.0501 \mu s\) and \(0.0474 \mu\) respectively which within the bounds of the ADC listed in Section 3.2.
+The built in ADC on Teensy has an effective sample rate of 1 MSPS (\emph{e.g.
+half of MAX11102AUB}), which would require an average execution speed of \textasciitilde{}\(2.667/2 \mu s \approx 1.334 \mu s\), making our implementation suitable even for that limited conversion speed.
+
+The main culprit that led other iterations to not be viable was the inclusion of \texttt{std::vector}, which was used to ensure we could deliver variable lengths of our return data -- e.g.~either 32, 64 or 128 bits.
+The overhead introduced from initializing and populating the vector led to significant delays in execution speed, leading to all other attempts at optimization to be superfluous.
+In fact, any other complexities led to significantly worse results than the naive, original implementation.
+
+However, it is important to note that both of these implementations has a hard limit of working on 64 bits input.
+Whereas larger bitstrings still yield results with high entropy, the execution speed increases far above the hardware limitations, as evident in \ref{tab:iter6}.
+Iteration 5 is hard capped at using only 64 bits input, as we need to store the bits in fixed-size integers before running them through the algorithm, and on embedded machines we have no larger integer types available.
+
+One issue still remains concerning the size of the input string and subsequent output string.
+The larger the size of the matrix provided to the Toeplitz function (\emph{e.g.~the larger our input}), the higher the potential overall entropy becomes.
+As our implementation is essentially capped at 64 bits input at this point in time (\emph{without being limited by the hardware}), this may or may not prove to not be secure enough for use in real world applications.
+This, however, is out of scope for this particular thesis, but a consideration nonetheless.
 

### 2025-06-02 docs/style/article-body.tex
added ent, conclusion

@@
 To verify that the algorithm successfully removed the patterns it should, we verified the measured entropy score with the command line utility \texttt{ent}\footnote{\href{https://manpages.ubuntu.com/manpages/trusty/man1/ent.1.html}{Manual page for \texttt{ent}, accessed 2025-04-23.}}.
+This tool verifies the statistical randomness of any given file, and does several calculations to provide an estimate of how random any given binary file is.
+
+The way \texttt{ent} calculates the entropy of a given binary file is by checking the frequency of certain byte values.
+If some byte values are more frequent, the binary file has a low entropy -- essentially singifying that there are predictable patterns.
+Conversely, if all byte values appear roughly equally, the binary file has a high entropy -- meaning that it is more random, and less predictable.
+The maximum entropy of any given file is \(\log_2(256) = 8 \text{ bits per byte}\), meaning that a perfectly random binary file has an entropy score of 8.
+For the purposes of our work, we consider values greater than \(7.99\) to be acceptable.
+
+Furthermore, \texttt{ent} also provides a measurement of the possible compression ratio of the file, e.g.~how much more the file can be theoretically compressed.
+This is calculated as \(1 - \frac{\text{entropy}}{\text{maximum entropy}}\), signifying that the file cannot be compressed further -- that the data is already as compressed as its random distribution allows.
+Essentially, this informs us that the file has no more detectable statistical patterns or compressable regularities remaining.
+We consider any file where \texttt{ent} reports a compression ratio of \(0\%\) to be acceptable.
 
@@
 \\ \hline array & 0.4284 \\ unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline
-\end{tabularx} \caption{Iteration 6 - Data type exloration} \label{tab:iter6}
+\end{tabularx} \caption{Iteration 6 - Data type exploration} \label{tab:iter6}
 \end{table}
@@
 
-\hypertarget{conclusion}{%
-\subsection{7 CONCLUSION}\label{conclusion}}
+\subsection{7 DISCUSSION}\label{discussion}
 
 Observing the results, we can place them into the context of the limits imposed by the ADC, as discussed in Section 3.2.
-In (3), we calculate the average execution speed required to be \(2.667 \mu\) for the ADC that requires soldering to the microcontroller.
+In algorithm (3) (\emph{as seen in Section 3.3}), we calculate the average execution speed required to be \(2.667 \mu\) for the ADC that requires soldering to the microcontroller.
 From Tables \ref{tab:iter5} and \ref{tab:iter6} we demonstrate execution speeds well below that, with iteration 5 as well as iteration 6 -- specifically with bitsets -- being the most prominent results to discuss.
 Here, our implementation achieved speeds of \(0.0501 \mu s\) and \(0.0474 \mu\) respectively which within the bounds of the ADC listed in Section 3.2.
-The built in ADC on Teensy has an effective sample rate of 1 MSPS (\emph{e.g.
-half of MAX11102AUB}), which would require an average execution speed of \textasciitilde{}\(2.667/2 \mu s \approx 1.334 \mu s\), making our implementation suitable even for that limited conversion speed.
+The built in ADC on Teensy has an effective sample rate of 1 MSPS (\emph{e.g.~half of MAX11102AUB}), which would require an average execution speed of \textasciitilde{}\(2.667/2 \mu s \approx 1.334 \mu s\), making our implementation suitable even for that limited conversion speed.
 
@@
 
-\hypertarget{changelog}{%
-\subsection{CHANGELOG}\label{changelog}}
+Any requirement of output integers larger than 32 bits may prove problematic.
+Whereas the bits processed by the algorithm does show an entropy measure quite close to 8, e.g.~reliably random, this does not necessarily mean that two sets of 32 bits can be combined on the host machine to produce a 64-bit integer.
+Combining these two sets may very well result in a less random result -- in essence, an entropy score that is not entirely random.
+As quantum computing evolves, the process of bruteforcing algorithms with all possible numbers becomes less of an impossible task, which is why this process of generating random numbers is required.
+The purpose of using quantum flucuation for random number generation is to provide greater assurances of security not possible from traditional, pseudo-random RNG, and being so severely limited by the size of the output is a major flaw in this implementation worth considering.
 
-2025-02-14: Added background section, smaller reviews to introduction.
+\subsection{8 CONCLUSION}\label{conclusion}
 
-2025-02-28: Template adjusted, added methodology.
-Started review of background and theory to add stronger correlation to computer science.
-Not yet finished due to review of articles as well as some additional information required from the project owner.
-The update to theory and background should be considered a heavy work in progress at this stage.
+This work set out to investigate the feasibility of implementing Toeplitz extraction on resource constrained hardware.
+Furthermore, we aimed to conclude whether the extraction algorithm could be executed fast enough to not be the limiting factor, e.g.~that the speed of the extraction did not exceed the speed of the ADC conversion.
+Ostensibly, these investigations were both successful, considering that we managed to create an implementation that executes in roughly one half microsecond on average, far exceeding the limits imposed by the ADC.
+The optimization with the algorithm revolved around removing as much complexity as possible, avoiding anything that is not the simplest form of data structure for maximum performance gains.
 
-2025-03-10: Moved evaluation down in the methodology in order to provide a better flow.
-Elaborated further on Toeplitz extraction and ADC converters, as well as motivating the selection of these.
-Some additional information added in introduction as motivation for the work.
+As mentioned in Section 7, this comes at a price.
+The implementation that did indeed perform above expectation could only output a 32-bit integer, as we cannot take any larger input than 64 bits without any more complex data structure.
+How much of a flaw this is for a system utilizing this method of OQRNG is beyond the scope of this thesis to investigate.
+However, it is likely that a larger output is required for any application revolving around tight security.
+For less critical systems and less essential functionality, this method might, however, prove just secure enough.
+For instance, an IoT-device that requires a random number output for symmetric cryptography when connecting to a user device may be an ideal candidate for this method of OQRNG.
 
-2025-03-13: Elaborated on background, as well as adding more details regarding hardware.
-Note that the hardware selected is subject to change over time.
-Further elaborated on related works in optimizing Toeplitz extraction.
+\subsection{9 FUTURE WORK}\label{future-work}
 
-2025-04-22: Begun including details regarding initial experimentation, updating details regarding experiments that had to change (e.g.~no baseline on separate hardware).
-
-2025-04-27: Added first table with test data.
-
-2025-05-20: Unfortunately, we've missed adding concrete details about changes made up until this point. Fix later.
+The main limitation for producing output with a larger size is the lack of a suitable data type in the limited standard library available for MCUs.
+Furthermore, it appears as if generic containers (\emph{such as \texttt{std::vector}}) that have already been severely optimized still impose a far greater performance penalty for this purpose.
+Future work may focus on implementing custom data types for larger bit sizes, for instance 128-bit and 256-bit integers.
+It may be possible that external libraries exist that alredy provide implementations of these integer sizes, but even then a slimmed down and customized implementation may prove the better alternative for this extreme performance constraint.

### 2025-06-02 docs/style/article-final.tex

@@
 
-\hypertarget{changelog}{%
-\subsection{CHANGELOG}\label{changelog}}
+Any requirement of output integers larger than 32 bits may prove problematic.
+Whereas the bits processed by the algorithm does show an entropy measure quite close to 8, e.g.~reliably random, this does not necessarily mean that two sets of 32 bits can be combined on the host machine to produce a 64-bit integer.
+Combining these two sets may very well result in a less random result -- in essence, an entropy score that is not entirely random.
+As quantum computing evolves, the process of bruteforcing algorithms with all possible numbers becomes less of an impossible task, which is why this process of generating random numbers is required.
+The purpose of using quantum flucuation for random number generation is to provide greater assurances of security not possible from traditional, pseudo-random RNG, and being so severely limited by the size of the output is a major flaw in this implementation worth considering.
 
-2025-02-14: Added background section, smaller reviews to introduction.
+\subsection{8 CONCLUSION}\label{conclusion}
 
-2025-02-28: Template adjusted, added methodology.
-Started review of background and theory to add stronger correlation to computer science.
-Not yet finished due to review of articles as well as some additional information required from the project owner.
-The update to theory and background should be considered a heavy work in progress at this stage.
+This work set out to investigate the feasibility of implementing Toeplitz extraction on resource constrained hardware.
+Furthermore, we aimed to conclude whether the extraction algorithm could be executed fast enough to not be the limiting factor, e.g.~that the speed of the extraction did not exceed the speed of the ADC conversion.
+Ostensibly, these investigations were both successful, considering that we managed to create an implementation that executes in roughly one half microsecond on average, far exceeding the limits imposed by the ADC.
+The optimization with the algorithm revolved around removing as much complexity as possible, avoiding anything that is not the simplest form of data structure for maximum performance gains.
 
-2025-03-10: Moved evaluation down in the methodology in order to provide a better flow.
-Elaborated further on Toeplitz extraction and ADC converters, as well as motivating the selection of these.
-Some additional information added in introduction as motivation for the work.
+As mentioned in Section 7, this comes at a price.
+The implementation that did indeed perform above expectation could only output a 32-bit integer, as we cannot take any larger input than 64 bits without any more complex data structure.
+How much of a flaw this is for a system utilizing this method of OQRNG is beyond the scope of this thesis to investigate.
+However, it is likely that a larger output is required for any application revolving around tight security.
+For less critical systems and less essential functionality, this method might, however, prove just secure enough.
+For instance, an IoT-device that requires a random number output for symmetric cryptography when connecting to a user device may be an ideal candidate for this method of OQRNG.
 
-2025-03-13: Elaborated on background, as well as adding more details regarding hardware.
-Note that the hardware selected is subject to change over time.
-Further elaborated on related works in optimizing Toeplitz extraction.
+\subsection{9 FUTURE WORK}\label{future-work}
 
-2025-04-22: Begun including details regarding initial experimentation, updating details regarding experiments that had to change (e.g.~no baseline on separate hardware).
-
-2025-04-27: Added first table with test data.
-
-2025-05-20: Unfortunately, we've missed adding concrete details about changes made up until this point. Fix later.
+The main limitation for producing output with a larger size is the lack of a suitable data type in the limited standard library available for MCUs.
+Furthermore, it appears as if generic containers (\emph{such as \texttt{std::vector}}) that have already been severely optimized still impose a far greater performance penalty for this purpose.
+Future work may focus on implementing custom data types for larger bit sizes, for instance 128-bit and 256-bit integers.
+It may be possible that external libraries exist that alredy provide implementations of these integer sizes, but even then a slimmed down and customized implementation may prove the better alternative for this extreme performance constraint.
 % BALANCE COLUMNS

### 2025-06-04 docs/style/article-body.tex
up to 10 pages

@@
 These systems require some input that is neither replicable nor reproducible.
-One method that can be realistically used is the inherently random movement of lava lamps \cite{lavarand}, which is used as a backup source of randomness for Cloudflare\footnote{\href{https://blog.cloudflare.com/randomness-101-lavarand-in-production/}{Cloudflare.com, accessed 2025-03-10}}.
+One method that can be realistically used is the inherently random movement of lava lamps \cite{lavarand}, which is used as a backup source of randomness for Cloudflare\footnote{\url{https://blog.cloudflare.com/randomness-101-lavarand-in-production/}}.
 Another proposed solution for this is quantum random number generation (\emph{QRNG}) \cite{QRNG}.
@@
 
-\textbf{Research area 2 (\emph{RA2})}: Can we ensure that the output of random numbers is not primarily limited primarily by our implementation, but rather limited by the processing power or the USB transfer speed of the microcontroller, alternatively by the ADC?
+\textbf{Research area 2 (\emph{RA2})}: Can we ensure that the output of random numbers is not primarily limited by our implementation, but rather limited by the processing power or the USB transfer speed of the microcontroller, alternatively by the ADC?
 
@@
 
-Many microcontrollers furthermore come equipped with internal ADCs that can be utilized, and while these provide a lower sample size (\emph{often around 1 MSPS}), the ease of development may be prudent to utilize for this proof-of-concept.
+Many modern microcontrollers (\emph{including both MCUs used in this study}) furthermore come equipped with internal ADCs that can be utilized.
+While these provide a lower sample size (\emph{often around 1 MSPS}), the ease of development may be prudent to utilize for this proof-of-concept.
 While our initial ADC has a fairly low throughput, this can always be upgraded if it ends up becoming too limiting.
 
-Should these internal ADCs prove too limiting, we propose utilizing MAX11102AUB\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} with an effective sample rate of 2 million samples per second (\emph{MSPS}).
+Should these internal ADCs prove too limiting, we propose utilizing MAX11102AUB\footnote{\url{https://www.farnell.com/datasheets/1913106.pdf}} with an effective sample rate of two million samples per second (\emph{MSPS}).
 This ADC provides a 12 bit sample size, providing roughly 24 Mbit/s of sampled data per second, derived by the following calculation.
@@
 High-performance MCUs enhance memory usage in various ways.
-Some of them use Tightly Coupled Memory\footnote{\href{https://www.farnell.com/datasheets/1913106.pdf}{Technical specification for MAX11102AUB, accessed 2025-03-13}} (\emph{TCM}), which gives fast SRAM with specific access routes for important data, avoiding cache misses and guaranteeing consistent performance.
+Some of them use Tightly Coupled Memory\footnote{\url{https://www.farnell.com/datasheets/1913106.pdf}} (\emph{TCM}), which gives fast SRAM with specific access routes for important data, avoiding cache misses and guaranteeing consistent performance.
 Moreover, instruction and data caching techniques, including instruction pre-fetching and branch prediction, help minimize execution delays in computationally intensive real-time applications.
-Another important aspect is Direct Memory Access\footnote{\href{https://www.sciencedirect.com/topics/computer-science/direct-memory-access}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (DMA), which enables data transfer between peripherals such as the ADC and RAM, without CPU intervention.
+Another important aspect is Direct Memory Access\footnote{\url{https://www.sciencedirect.com/topics/computer-science/direct-memory-access}} (DMA), which enables data transfer between peripherals such as the ADC and RAM, without CPU intervention.
 This offloading reduces processing overhead, allowing the MCU to manage fast data transfers effectively.
@@
 
-In our work, we intend to use Teensy 4.1\footnote{\href{https://www.pjrc.com/store/teensy41.html}{Teensy developer documentation, accessed 2025-02-27}}, based on the ARM Cortex-M7.
+\begin{figure}
+\centering
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio,alt={Teensy 4.1 on the left, compared to Raspberry Pi Pico 2 on the right.}]{img/mcu.jpeg}
+\caption{Teensy 4.1 on the left, compared to Raspberry Pi Pico 2 on the right.}\label{fig:mcu}
+\end{figure}
+
+In our work, we intend to use Teensy 4.1\footnote{\url{https://www.pjrc.com/store/teensy41.html}} based on the ARM Cortex-M7 pictured to the right in Figure \cite{fig:mcu}.
 This MCU is especially suitable for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and Digital Signal Processing (\emph{DSP}) capabilities.
-The floating-point unit\footnote{\href{https://www.sciencedirect.com/topics/computer-science/floating-point-unit}{ScienceDirect Journals \& Books, accessed 2025-03-13}} (\emph{FPU}) and Single Instruction, Multiple Data (\emph{SIMD}) style DSP instructions improve how quickly it can perform bitwise and arithmetic tasks, which is crucial for quick Toeplitz extraction.
+The floating-point unit\footnote{\url{https://www.sciencedirect.com/topics/computer-science/floating-point-unit}} (\emph{FPU}) and Single Instruction, Multiple Data (\emph{SIMD}) style DSP instructions improve how quickly it can perform bitwise and arithmetic tasks, which is crucial for quick Toeplitz extraction.
 SIMD-controlled DSP architectures, as described by Han et al.
@@
 In order to evaluate how efficent our implementation can become, our aim is to try our implementation on other MCU with different level of power and hardware support.
-Whereas Teensy 4.1, operating at \texttt{600\ MHz}, is our primary development platform which we will evaluate closely, we aim to run our implementations on the Raspberry Pi Pico 2\footnote{\href{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}{Raspberry Pi Pico 2 documentation, accessed 2025-03-13}}, operating at \texttt{150\ MHz}.
+Whereas Teensy 4.1, operating at \texttt{600\ MHz}, is our primary development platform which we will evaluate closely, we aim to run our implementations on the Raspberry Pi Pico 2\footnote{\url{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}}, operating at \texttt{150\ MHz} (\emph{pictured to the right in Figure \cite{fig:mcu}}).
 Due to the lower computational power of this MCU, there may be significant issues in utilizing this weaker model, yet it is significantly cheaper and easier to access.
@@
 
-The initial implementation followed the pseudocode described in Algorithm \ref{alg:bit-conv}, using matrix multiplication over raw input and seed data.
-It relied on \texttt{std::vector\textless{}int\textgreater{}} for storage and used nested loops to compute each output bit.
-From this point, this implementation was improved over the coming iterations, and the new implementation verified in the same manner as the initial version.
+The initial implementation followed the pseudocode described in Algorithm \ref{alg:bit-conv}, using matrix multiplication over the raw input and seed data.
+It relied on \texttt{std::vector\textless{}int\textgreater{}} for storing the raw data, and utilized nested loops to compute each output bit.
+From this point, this implementation was adapted over the coming iterations, and each new version was then verified in the same manner as the initial version as outlined in Section 5.3.
 
@@
 
-\textbf{Iteration 2 - Bitshifting:} Basic bitwise operation was introduced to replace arithmetic whenever possible.
-Multiplication was replaced with logical \texttt{AND\ \&} and modulo operations with bit masking \texttt{\&\ 1}.
+\textbf{Iteration 2 - Bitshifting:} Basic bitwise operation was introduced to replace arithmetic operations whenever possible.
+Multiplication was replaced with logical \texttt{AND\ \&}, and modulo operations with bit masking \texttt{\&\ 1}.
 The goal was to reduce the number of instructions and improve per-bit processing speed.
@@
 This was followed by isolated use of ARM instructions such as \texttt{\_\_builtin\_popcountll()} (\emph{which counts the number of set bits in an unsigned integer}).
-After establishing their individual effect we combined both techniques, multiple batch sizes were tested to determine their impact.
+After establishing their individual effect, we combined both techniques, Multiple batch sizes were tested to determine their impact.
 More details and benchmarks for each configuration can be found in the section 6.
@@
 
-Phase two focused on addressing inefficiencies and design issues that were unintentionally introduced during earlier iterations.
+Phase two focused on addressing inefficiencies and design issues that were unintentionally introduced or not addressed during earlier iterations.
 Rather than continuing with new algorithmic ideas, this phase aimed to identify and fix structural problems.
@@
 The goal was to decrease overhead created by loops.
-Whereas this operation is commonly done by compiler optimization, manually performing this guarantees that we unroll the loops rather than leaving it to the compiler.
+Whereas this operation is commonly done by compiler optimization, manually unrolling the loops guarantees the optimization rather than leaving it to the compiler.
 
@@
 
-To evaluate the correctness of each implementation, a baseline was generated as discussed in section 5.2.
+To evaluate the correctness of each implementation, a baseline was generated as discussed in Section 5.
 Using the naive, initial implementation to process bits and saving for later evaluation gave us a source of truth against which to compare following iterations.
-To verify that the algorithm successfully removed the patterns it should, we verified the measured entropy score with the command line utility \texttt{ent}\footnote{\href{https://manpages.ubuntu.com/manpages/trusty/man1/ent.1.html}{Manual page for \texttt{ent}, accessed 2025-04-23.}}.
+To verify that the algorithm successfully removed the patterns it should, we verified the measured entropy score with the command line utility \texttt{ent}\footnote{\url{https://manpages.ubuntu.com/manpages/trusty/man1/ent.1.html}}.
 This tool verifies the statistical randomness of any given file, and does several calculations to provide an estimate of how random any given binary file is.
@@
 
-Hyncica et.
+In regards to execution time, Hyncica et.
 al.
@@
 
-In our implementation, execution time for each extraction was measured using Arduino's built-in \texttt{micros()}\footnote{\href{https://docs.arduino.cc/language-reference/en/functions/time/micros/}{Documentation for \texttt{micros()}, accessed 2025-05-02.}} function.
+In our implementation, execution time for each extraction was measured using Arduino's built-in \texttt{micros()}\footnote{\url{https://docs.arduino.cc/language-reference/en/functions/time/micros/}} function.
 This function offers a convenient and platform-independent method for recording elapsed time with microsecond precision, enabling consistent performance evaluation across different iterations.
+
+To isolate the execution speed of the algorithm as precisely as possible, we only intend to measure the execution speed of the algorithm.
+Whereas there is some other code present in each iteration (\emph{primarily packing the output data to be sent to the host computer in a workable format}), the brunt of the work should reasonably be performed by the algorithm.
+As such, we expect the time spent on executing other code to be negligible.
+Furthermore, this code performs significantly less work per cycle, and does not exponentially increase the larger the input and output requirements increase.
 
@@
 
+Whereas our implementation can be controlled at the algorithmic and structural level, compiler optimization are managed outside of our influence.
+The Arduino framework automatically sets compiler flags and optimization levels specifically chosen for the hardware -- for instance, specific flags are selected when compiling code for Teensy, applying aggressive settings for performance.
+However, we cannot determine exactly how these flags influence our iterations.
+Consequently, this makes it impossible to isolate the impact of compiler optimizations from the performance gains of our implementations.
+Therefore, any performance differences observed in our experiments are considered to stem solely from our implementation, as the compiler behavior remains opaque and unchangeable in this context.
+
 \subsection{6 RESULTS}\label{results}
@@
 
-\begin{table}[ht] \centring
+\begin{table}[H] \centring
 \begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
@@
 \hline 64 & 13.1564 & 106.3914 \\ 512 & 788.3139 & 5302.4979 \\ 1024 & 3124.0580
-& 21111.2163 \\ \hline \end{tabularx} \caption{Iteration 1 - Data Structures}
-\label{tab:iter1} \end{table}
+& 21111.2163 \\ \hline \end{tabularx} \caption{Iteration 0 - Naive
+implementation} \label{tab:iter0} \end{table}
 
-Table \ref{tab:iter1} presents the average execution time of iteration 1 on both Teensy 4.1 and Raspberry Pico Pi 2 across the input sizes.
+Table \ref{tab:iter0} presents the average execution time of iteration 1 on both Teensy 4.1 and Raspberry Pico Pi 2 across the input sizes.
 The Teensy consistently outperforms the Pico, with the gap widening as the bit size increases.
+These results will be used as baseline for future comparison.
+
+\begin{table}[ht] \centring
+\begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
+\hline \textbf{Bit size} & \textbf{Teensy ($\mu$s)} \\ \hline pointers & 13.7999
+\\ hash map & 68.8099 \\ \hline \end{tabularx} \caption{Iteration 1 - Data
+Structures} \label{tab:iter1} \end{table}
+
+Table \ref{tab:iter1} presents the average execution time of iteration 1 on Teensy 4.1.
+This iteration introduced additional complexity without yielding significant gains and in some cases led to worse performance.
+Although this iteration did not achieve the desired performance gains, it provided valuable insight into the limitations of certain data structure choices within the context of this study.
+
+Notably, as the execution speed of the more powerful Teensy 4.1 increased with this iteration, we omit the limited results of executing the same code on Pico 2.
+During the remainder of this phase as well as the next, test results of any given iteration of Pico 2 are only reported if there are significant enough improvements in execution speeds for Teensy to potentially lead to gains for the weaker MCU.
 
@@
 
-\vspace{1em}
-
-\begin{table}[ht] \centring
+\vspace{1em} \begin{table}[ht] \centring
 \begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
@@
 This iteration led to an extremely efficient implementation that approaches the physical execution limits of the Teensy 4.1.
+Due to the removal of \texttt{std::vector}, this implementation is limited by the available integer sizes of the MCU, and could as such only be tested with at most 64 bits.
+The ramifications of this will be discussed in more detail in Section 7.
 
@@
 \hline \textbf{Data structure} & \multicolumn{1}{c|}{\textbf{Teensy ($\mu s$)}}
-\\ \hline array & 0.4284 \\ unordered_map & 31.5090 \\ bitset & 0.0474 \\ \hline
-\end{tabularx} \caption{Iteration 6 - Data type exploration} \label{tab:iter6}
-\end{table}
+\\ \hline array & 0.4284 \\ unordered\_{map} & 31.5090 \\ bitset & 0.0474 \\
+\hline \end{tabularx} \caption{Iteration 6 - Data type exploration}
+\label{tab:iter6} \end{table}
 
-Table \ref{tab:iter6} presents the results of iteration 6 using 64 bit size, which was introduced to address the structural limitations encountered in iteration 5.
-Specifically, the fixed-width output constrains imposed by the removal of vector.
-In this iteration, alternative data structures were evaluated.
-Vector-based approach from iteration 1 serving as a baseline for comparison.
+Table \ref{tab:iter6} presents the results of iteration 6 using 64 bits input size, which was introduced to address the structural limitations encountered in iteration 5 (\emph{specifically, the fixed-width output constrains imposed by the removal of \texttt{std::vector}}).
+In this iteration, another set of alternative data structures were evaluated, the vector-based approach from iteration 1 serving as a baseline for comparison.
 
-Additionally, 128 bit size were tested resulting in \texttt{179.8712} \(\mu s\).
+As the bitset appeared to provide gains in execution speed, we attempted to increase the input bit size to 128 bits in order to investigate whether this data structure could allow for larger outputs.
+However, this substantially increased the execution time to \texttt{179.8712} \(\mu s\).
 
@@
 Observing the results, we can place them into the context of the limits imposed by the ADC, as discussed in Section 3.2.
-In algorithm (3) (\emph{as seen in Section 3.3}), we calculate the average execution speed required to be \(2.667 \mu\) for the ADC that requires soldering to the microcontroller.
+In algorithm (3) (\emph{as seen in Section 3.3}), we calculate the average execution speed required to be \(2.667 \mu s\) for the ADC that requires soldering to the microcontroller.
 From Tables \ref{tab:iter5} and \ref{tab:iter6} we demonstrate execution speeds well below that, with iteration 5 as well as iteration 6 -- specifically with bitsets -- being the most prominent results to discuss.
-Here, our implementation achieved speeds of \(0.0501 \mu s\) and \(0.0474 \mu\) respectively which within the bounds of the ADC listed in Section 3.2.
+Here, our implementation achieved speeds of \(0.0501 \mu s\) and \(0.0474 \mu s\) respectively which within the bounds of the ADC listed in Section 3.2.
 The built in ADC on Teensy has an effective sample rate of 1 MSPS (\emph{e.g.~half of MAX11102AUB}), which would require an average execution speed of \textasciitilde{}\(2.667/2 \mu s \approx 1.334 \mu s\), making our implementation suitable even for that limited conversion speed.
@@
 It may be possible that external libraries exist that alredy provide implementations of these integer sizes, but even then a slimmed down and customized implementation may prove the better alternative for this extreme performance constraint.
+
+Finally, this proof-of-concept has not been tested when connected to the OQRNG-device itself, and has focused on a theoretical exploration of the capabilities of the MCUs as well as the limitations of Toeplitz extraction.
+Sampling raw, live data from the OQRNG-device and processing it utilizing the firmware developed in this work may reveal flaws not evident from our tests, as well as demonstrating the efficacy of our implementation in a real world scenario.

### 2025-06-04 docs/style/article-final.tex
up to 10 pages

@@
 It may be possible that external libraries exist that alredy provide implementations of these integer sizes, but even then a slimmed down and customized implementation may prove the better alternative for this extreme performance constraint.
+
+Finally, this proof-of-concept has not been tested when connected to the OQRNG-device itself, and has focused on a theoretical exploration of the capabilities of the MCUs as well as the limitations of Toeplitz extraction.
+Sampling raw, live data from the OQRNG-device and processing it utilizing the firmware developed in this work may reveal flaws not evident from our tests, as well as demonstrating the efficacy of our implementation in a real world scenario.
 % BALANCE COLUMNS

### 2025-06-04 docs/style/article-body.tex

@@
+\section{ABSTRACT}\label{abstract}
+
+True random number generation is indispensable for a wide range of cryptographic and distributed computing applications, yet traditional pseudo-random number generators remain deterministic, and thus vulnerable to state‐recovery attacks.
+Quantum random number generation (QRNG) offers a robust alternative by harnessing inherent quantum fluctuations -- in particular, optical QRNG (OQRNG) based on shot noise provides an accessible source of entropy from which true random numbers can be generated.
+However, existing OQRNG solutions typically rely on expensive or bespoke hardware and defer post‐processing to host computers, limiting portability and integration.
+In this work, we investigate the feasibility of performing Toeplitz randomness extraction directly on resource constrained microcontrollers, thus creating a fully self‐contained QRNG system.
+We address two primary research areas: (RA1) developing an implementation of Toeplitz extraction that minimizes processing time on embedded hardware, and (RA2) ensuring that the extraction algorithm does not become the throughput bottleneck compared to ADC conversion or USB transfer speeds.
+
+Our results demonstrate that, by eliminating dynamic allocations and leveraging bit‐parallel techniques, Toeplitz extraction on 64‐bit inputs can be executed in approximately \(0.05 \mu s\) on the Teensy 4.1 -- well below the \(2.667 \mu s\) threshold imposed by a 12‐bit, 2 MSPS ADC.
+Even accounting for variance in individual measurements, these speeds confirm that the algorithmic implementation does not constrain overall throughput.
+On the Pico 2, the same implementations yield sub‐microsecond performance, albeit with a wider variance.
+However, our most efficient implementations are limited to 64‐bit inputs -- extending to larger input bit sizes incurs significant overhead due to the absence of wider integer types in typical MCU toolchains.
+
+In conclusion, Toeplitz extraction can be performed in real time on low‐power microcontrollers without becoming the system's performance bottleneck, validating the viability of a portable QRNG thumbstick.
+Future work should explore custom data types or lightweight libraries for 128- and 256-bit size extraction, as well as integration with a live OQRNG device to assess end‐to‐end randomness quality and throughput.
+
 \section{1 INTRODUCTION}\label{introduction}
@@
 
+To estimate the theoretical lower bound for execution time on the Teensy 4.1, the duration of a single CPU cycle can be calculated using the following expression:
+
+\[
+\text{Cycle time} = \frac{1}{600\, \mathrm{MHz}} \approx 1.67\, \mathrm{ns} \phantom{123}(2)
+\]
+
+Given a clock frequency of \texttt{600\ MHz}, each cycle lasts approximately 1.67 nanoseconds.
+Assuming a minimal implementation requires 30 instruction cycles to complete a single extraction, the theoretical execution time becomes:
+
+\[
+t_{\text{min}} = 30 \times 1.67\,\mathrm{ns} = 50.1\,\mathrm{ns} \approx 0.05\,\mu\mathrm{s} \phantom{123}(3)
+\]
+
+These theoretical limits will serve as a baseline to assess how closely our implementations approach the maximum capabilities of the Teensy 4.1 microcontroller.
+
 \subsubsection{3.4 Toeplitz extraction}\label{toeplitz-extraction}
@@
 \times 10^{-3}
-\phantom{12}(2)
+\phantom{12}(4)
 \]
@@
 =2.667\ \mu\mathrm{s}.
-\phantom{12}(3)
+\phantom{12}(5)
 \]
@@
 
-Table \ref{tab:iter0} presents the average execution time of iteration 1 on both Teensy 4.1 and Raspberry Pico Pi 2 across the input sizes.
-The Teensy consistently outperforms the Pico, with the gap widening as the bit size increases.
-These results will be used as baseline for future comparison.
+Table \ref{tab:iter0} presents the average execution time of the naive implementation (iteration 0) across different input sizes for both the Teensy 4.1 and the Raspberry Pi Pico 2.
+The results reveal that the Teensy consistently achieves lower execution times compared to the Pico, with performance differences becoming more pronounced as thje bit size increases.
+This behaivor aligns with the expected difference in hardware specifications, such as CPU clock frequency and memory bandwidth.
+
+These measurements serve as the baseline for all subsequent iterations.
+They provide a reference point for assessing performance improvements introduced in later implementations and allow for cross-platform comparisons.
 
@@
 In addition to the results presented in Table \ref{tab:iter4}, a further tests was conducted on the Teensy using a single-loop unrolled implementation for the 64-bit size, resulting in an average execution time of \texttt{6.6626} \(\mu s\).
+
 An additional fully unrolled variant, where loops were entirely eliminated, produced a measured execution time of \texttt{0.0491} \(\mu s\).
-However, this measurement was later determined to be invalid due to packaging error.
+However, this result was later deemed invalid due to packaging error in the implementation.
+Specifically, the algorithm was incorrectly designed to transmit only a single output bit instead of full 32-bit expected output.
+This lead to an underestimation of the true execution time, as the cost of correctly formatting and sending the full result was omitted from measurement.
 
@@
 The ramifications of this will be discussed in more detail in Section 7.
+However, the results remain inconsistent---while the average is low, individual measurements vary, indicating that the implementation does not yet achieve stable execution at the architectural limit.
 
-\vspace{1em}
-
-\begin{table}[H] \centring
+\vspace{1em} \begin{table}[H] \centring
 \begin{tabularx}{\columnwidth}{|>{\centering\arraybackslash}X|>{\centering\arraybackslash}X|}
@@
 
-Table \ref{tab:iter6} presents the results of iteration 6 using 64 bits input size, which was introduced to address the structural limitations encountered in iteration 5 (\emph{specifically, the fixed-width output constrains imposed by the removal of \texttt{std::vector}}).
-In this iteration, another set of alternative data structures were evaluated, the vector-based approach from iteration 1 serving as a baseline for comparison.
+Table \ref{tab:iter6} presents the results of iteration 6 using 64 bit size, which was introduced to address the structural limitations encountered in iteration 5.
+Specifically, the fixed-width output constrains imposed by the removal of vector.
+Several alternative data structures were evaluated, with the original vecotr-based implementation from iteration 1 serving as a baseline.
 
-As the bitset appeared to provide gains in execution speed, we attempted to increase the input bit size to 128 bits in order to investigate whether this data structure could allow for larger outputs.
-However, this substantially increased the execution time to \texttt{179.8712} \(\mu s\).
+Notably, while the bitset-based implementation achieved an average execution time of 0.0474\,\(\mu s\), this value represents an average and not a consistently stable result.
+Individual measurements showed significant variance.
+This suggests that, despite nearing the theoretical limits outlined in Section 3.3, the implementation has not yet reached a consistently minimal execution time.
+Additional tests at 128-bit input size yielded 179.8712\,\(\mu s\), confirming that the system still operates within the physical capabilities of the Teensy 4.1.
 
@@
 Observing the results, we can place them into the context of the limits imposed by the ADC, as discussed in Section 3.2.
-In algorithm (3) (\emph{as seen in Section 3.3}), we calculate the average execution speed required to be \(2.667 \mu s\) for the ADC that requires soldering to the microcontroller.
+In equation (5) (\emph{as seen in Section 5.3}), we calculate the average execution speed required to be \(2.667 \mu s\) for the ADC that requires soldering to the microcontroller.
 From Tables \ref{tab:iter5} and \ref{tab:iter6} we demonstrate execution speeds well below that, with iteration 5 as well as iteration 6 -- specifically with bitsets -- being the most prominent results to discuss.

### 2025-06-04 docs/style/article-final.tex

@@
 \maketitle
+
+\section{ABSTRACT}\label{abstract}
+
+True random number generation is indispensable for a wide range of cryptographic and distributed computing applications, yet traditional pseudo-random number generators remain deterministic, and thus vulnerable to state‐recovery attacks.
+Quantum random number generation (QRNG) offers a robust alternative by harnessing inherent quantum fluctuations -- in particular, optical QRNG (OQRNG) based on shot noise provides an accessible source of entropy from which true random numbers can be generated.
+However, existing OQRNG solutions typically rely on expensive or bespoke hardware and defer post‐processing to host computers, limiting portability and integration.
+In this work, we investigate the feasibility of performing Toeplitz randomness extraction directly on resource constrained microcontrollers, thus creating a fully self‐contained QRNG system.
+We address two primary research areas: (RA1) developing an implementation of Toeplitz extraction that minimizes processing time on embedded hardware, and (RA2) ensuring that the extraction algorithm does not become the throughput bottleneck compared to ADC conversion or USB transfer speeds.
+
+Our results demonstrate that, by eliminating dynamic allocations and leveraging bit‐parallel techniques, Toeplitz extraction on 64‐bit inputs can be executed in approximately \(0.05 \mu s\) on the Teensy 4.1 -- well below the \(2.667 \mu s\) threshold imposed by a 12‐bit, 2 MSPS ADC.
+Even accounting for variance in individual measurements, these speeds confirm that the algorithmic implementation does not constrain overall throughput.
+On the Pico 2, the same implementations yield sub‐microsecond performance, albeit with a wider variance.
+However, our most efficient implementations are limited to 64‐bit inputs -- extending to larger input bit sizes incurs significant overhead due to the absence of wider integer types in typical MCU toolchains.
+
+In conclusion, Toeplitz extraction can be performed in real time on low‐power microcontrollers without becoming the system's performance bottleneck, validating the viability of a portable QRNG thumbstick.
+Future work should explore custom data types or lightweight libraries for 128- and 256-bit size extraction, as well as integration with a live OQRNG device to assess end‐to‐end randomness quality and throughput.
 

### 2025-06-04 docs/style/article-body.tex
final version?

@@
 True random number generation is indispensable for a wide range of cryptographic and distributed computing applications, yet traditional pseudo-random number generators remain deterministic, and thus vulnerable to state‐recovery attacks.
-Quantum random number generation (QRNG) offers a robust alternative by harnessing inherent quantum fluctuations -- in particular, optical QRNG (OQRNG) based on shot noise provides an accessible source of entropy from which true random numbers can be generated.
+Quantum random number generation (\emph{QRNG}) offers a robust alternative by harnessing inherent quantum fluctuations -- in particular, optical QRNG (\emph{OQRNG}) based on shot noise provides an accessible source of entropy from which true random numbers can be generated.
 However, existing OQRNG solutions typically rely on expensive or bespoke hardware and defer post‐processing to host computers, limiting portability and integration.
 In this work, we investigate the feasibility of performing Toeplitz randomness extraction directly on resource constrained microcontrollers, thus creating a fully self‐contained QRNG system.
-We address two primary research areas: (RA1) developing an implementation of Toeplitz extraction that minimizes processing time on embedded hardware, and (RA2) ensuring that the extraction algorithm does not become the throughput bottleneck compared to ADC conversion or USB transfer speeds.
+We address two primary research areas: RA1 developing an implementation of Toeplitz extraction that minimizes processing time on embedded hardware, and RA2 ensuring that the extraction algorithm does not become the throughput bottleneck compared to ADC conversion or USB transfer speeds.
 
-Our results demonstrate that, by eliminating dynamic allocations and leveraging bit‐parallel techniques, Toeplitz extraction on 64‐bit inputs can be executed in approximately \(0.05 \mu s\) on the Teensy 4.1 -- well below the \(2.667 \mu s\) threshold imposed by a 12‐bit, 2 MSPS ADC.
+Our results demonstrate that, by eliminating dynamic allocations and leveraging bit parallel techniques, Toeplitz extraction on 64-bit inputs can be executed in approximately \(0.05 \mu s\) on the Teensy 4.1 -- well below the \(2.667 \mu s\) threshold imposed by a 12-bit, 2 MSPS ADC.
 Even accounting for variance in individual measurements, these speeds confirm that the algorithmic implementation does not constrain overall throughput.
@@
 The post-processing finally yields a random number.
-This relationship can be seen in Figure \ref{fig:linear-system}.
+This relationship can be seen in fig.~\ref{fig:linear-system}.
 
@@
 
-In our work, we intend to use Teensy 4.1\footnote{\url{https://www.pjrc.com/store/teensy41.html}} based on the ARM Cortex-M7 pictured to the right in Figure \cite{fig:mcu}.
+In our work, we intend to use Teensy 4.1\footnote{\url{https://www.pjrc.com/store/teensy41.html}} based on the ARM Cortex-M7 pictured to the right in fig.~\ref{fig:mcu}.
 This MCU is especially suitable for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and Digital Signal Processing (\emph{DSP}) capabilities.
@@
 
-In order to evaluate how efficent our implementation can become, our aim is to try our implementation on other MCU with different level of power and hardware support.
-Whereas Teensy 4.1, operating at \texttt{600\ MHz}, is our primary development platform which we will evaluate closely, we aim to run our implementations on the Raspberry Pi Pico 2\footnote{\url{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}}, operating at \texttt{150\ MHz} (\emph{pictured to the right in Figure \cite{fig:mcu}}).
+In order to evaluate how efficient our implementation can become, our aim is to try our implementation on other MCU with different level of power and hardware support.
+Whereas Teensy 4.1, operating at \texttt{600\ MHz}, is our primary development platform which we will evaluate closely, we aim to run our implementations on the Raspberry Pi Pico 2\footnote{\url{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}}, operating at \texttt{150\ MHz} (\_pictured to the right in fig.~\ref{fig:mcu_}).
 Due to the lower computational power of this MCU, there may be significant issues in utilizing this weaker model, yet it is significantly cheaper and easier to access.
@@
 \cite{toeplitz-desc}.
-This work focuses on implementing Toeplitz extraction on field-programmable gate-arrays (FPGA), but some specific details can be derived from their work.
+This work focuses on implementing Toeplitz extraction on field-programmable gate-arrays (\emph{FPGA}), but some specific details can be derived from their work.
 As these authors describe, Toeplitz extraction is a strong contender for our work due to a lower computational complexity than other alternatives, as well as a relatively easy algorithm to use.
@@
 \begin{algorithm}[ht] \caption{Toeplitz extraction}\label{alg:bit-conv}
-\begin{algorithmic}[1] \REQUIRE \(x[0\,..\,n-1]\) \Comment{input bit array of
+\begin{algorithmic}[1] \REQUIRE \(k[0\,..\,n-1]\) \Comment{input bit array of
 length \(n\)} \REQUIRE \(t[0\,..\,n+m-2]\) \Comment{seed matrix of length
@@
 \FOR{\(i = 0\) to \(m - 1\)} \STATE sum = 0 \FOR{\(j = 0\) to \(n - 1\)} \STATE
-sum = sum + \(x[j] \* t[i + j]\) \ENDFOR \STATE \(y[i] =
+sum = sum + \(k[j] $\times$ t[i + j]\) \ENDFOR \STATE \(y[i] =
 \mathrm{sum}\;\bmod\;2\) \ENDFOR \RETURN \(y\) \end{algorithmic} \end{algorithm}
@@
 
-Table \ref{tab:iter0} presents the average execution time of the naive implementation (iteration 0) across different input sizes for both the Teensy 4.1 and the Raspberry Pi Pico 2.
+Table \ref{tab:iter0} presents the average execution time of the naive implementation (\emph{iteration 0}) across different input sizes for both the Teensy 4.1 and the Raspberry Pi Pico 2.
 The results reveal that the Teensy consistently achieves lower execution times compared to the Pico, with performance differences becoming more pronounced as thje bit size increases.

### 2025-06-04 docs/style/article-body.tex
the draft sent in for preliminary supervision

@@
-\section{ABSTRACT}\label{abstract}
-
-True random number generation is indispensable for a wide range of cryptographic and distributed computing applications, yet traditional pseudo-random number generators remain deterministic, and thus vulnerable to state‐recovery attacks.
-Quantum random number generation (\emph{QRNG}) offers a robust alternative by harnessing inherent quantum fluctuations -- in particular, optical QRNG (\emph{OQRNG}) based on shot noise provides an accessible source of entropy from which true random numbers can be generated.
-However, existing OQRNG solutions typically rely on expensive or bespoke hardware and defer post‐processing to host computers, limiting portability and integration.
-In this work, we investigate the feasibility of performing Toeplitz randomness extraction directly on resource constrained microcontrollers, thus creating a fully self‐contained QRNG system.
-We address two primary research areas: RA1 developing an implementation of Toeplitz extraction that minimizes processing time on embedded hardware, and RA2 ensuring that the extraction algorithm does not become the throughput bottleneck compared to ADC conversion or USB transfer speeds.
-
-Our results demonstrate that, by eliminating dynamic allocations and leveraging bit parallel techniques, Toeplitz extraction on 64-bit inputs can be executed in approximately \(0.05 \mu s\) on the Teensy 4.1 -- well below the \(2.667 \mu s\) threshold imposed by a 12-bit, 2 MSPS ADC.
-Even accounting for variance in individual measurements, these speeds confirm that the algorithmic implementation does not constrain overall throughput.
-On the Pico 2, the same implementations yield sub‐microsecond performance, albeit with a wider variance.
-However, our most efficient implementations are limited to 64‐bit inputs -- extending to larger input bit sizes incurs significant overhead due to the absence of wider integer types in typical MCU toolchains.
-
-In conclusion, Toeplitz extraction can be performed in real time on low‐power microcontrollers without becoming the system's performance bottleneck, validating the viability of a portable QRNG thumbstick.
-Future work should explore custom data types or lightweight libraries for 128- and 256-bit size extraction, as well as integration with a live OQRNG device to assess end‐to‐end randomness quality and throughput.
-
 \section{1 INTRODUCTION}\label{introduction}
@@
 For instance, the speed at which the ADC can process the optical signal into raw bits as well as the speed that the USB output can transfer processed random number to the host computer will be limiting factors.
-Further details on the limitations of the ADC will be outlined in Section 3.
+Further details on the limitations of the ADC will be outlined in Section 3.2.
 The slowest of these bottlenecks will inevitably be the limiting factor for any implementation.
@@
 Section 5 will present our methodology and implementation strategy, as well as some limitations imposed on our work.
-Finally, section 7 will present the results of our experimentation.
+Finally, Section 7 will present the results of our experimentation, Section 8 presents our conclusions and Section 9 presents future work.
 
@@
 In this work, quantum shot noise originating from photodiodes was studied, and in so doing a prototype device was constructed.
-This prototype utilized an LED soldered millimeters apart from a photodiode, which in turn outputs a variable analog signal from the device.
 In this section, we introduce the remaining hardware used for our implementations as well as the considerations taken in order to shift the focus from physics to computer science.
@@
 The photodiode detects light from the LED, and converts the light into a current signal, which is sent to a transimpedance amplifier to convert it into measurable voltage.
-In order to minimize disruptions by other external lights, the system is enclosed in a shielded measurment box.
+In order to minimize disruptions by other external lights, the system is enclosed in a shielded measurement box.
 
@@
 Microcontrollers (\emph{MCUs}) are compact and low-power computing devices designed for embedded systems and real-time operations, and suitable as a processing unit for the purposes of this work.
-Unlike general CPUs, an MCU integrates a processor, memory and peripherals (\emph{such as an ADC}) into a single chip.
+Unlike general CPUs, an MCU integrates a processor, memory and certain peripherals (\emph{such as an ADC}) into a single chip.
 Furthermore, modern MCUs often feature advanced microarchitectural elements to enhance processing capabilities on single threads (\emph{such as dual-issue superscalar architectures, allowing the MCU to run several instructions per CPU cycle}), making them suitable candidates for the post-processing required for OQRNG-data.
@@
 Moreover, instruction and data caching techniques, including instruction pre-fetching and branch prediction, help minimize execution delays in computationally intensive real-time applications.
-Another important aspect is Direct Memory Access\footnote{\url{https://www.sciencedirect.com/topics/computer-science/direct-memory-access}} (DMA), which enables data transfer between peripherals such as the ADC and RAM, without CPU intervention.
+Another important aspect is Direct Memory Access\footnote{\url{https://www.sciencedirect.com/topics/computer-science/direct-memory-access}} (\emph{DMA}), which enables data transfer between peripherals such as the ADC and RAM, without CPU intervention.
 This offloading reduces processing overhead, allowing the MCU to manage fast data transfers effectively.
@@
 
-In our work, we intend to use Teensy 4.1\footnote{\url{https://www.pjrc.com/store/teensy41.html}} based on the ARM Cortex-M7 pictured to the right in fig.~\ref{fig:mcu}.
+In our work, we intend to use Teensy 4.1\footnote{\url{https://www.pjrc.com/store/teensy41.html}} based on the ARM Cortex-M7 pictured to the left in fig.~\ref{fig:mcu}.
 This MCU is especially suitable for computationally demanding tasks involving randomness extraction due to its dual-issue superscalar architecture and Digital Signal Processing (\emph{DSP}) capabilities.
@@
 In order to evaluate how efficient our implementation can become, our aim is to try our implementation on other MCU with different level of power and hardware support.
-Whereas Teensy 4.1, operating at \texttt{600\ MHz}, is our primary development platform which we will evaluate closely, we aim to run our implementations on the Raspberry Pi Pico 2\footnote{\url{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}}, operating at \texttt{150\ MHz} (\_pictured to the right in fig.~\ref{fig:mcu_}).
+Whereas Teensy 4.1, operating at \texttt{600\ MHz}, is our primary development platform which we will evaluate closely, we aim to run our implementations on the Raspberry Pi Pico 2\footnote{\url{https://datasheets.raspberrypi.com/pico/pico-2-product-brief.pdf}}, pictured to the right in fig.~\ref{fig:mcu}, operating at \texttt{150\ MHz}.
 Due to the lower computational power of this MCU, there may be significant issues in utilizing this weaker model, yet it is significantly cheaper and easier to access.
@@
 The raw bits from the ADC can potentially have some deterministic patterns, and as such have to be processed somehow in order to remove these patterns.
-Several methods exist for this purpose, and for our work, we will perform this pre-processing via Toeplitz extraction.
+Several methods exist for this purpose, and for our work, we will perform this post-processing via Toeplitz extraction.
 The main focus of this study is to implement this extraction algorithm as effectively as possible on resource constrained hardware.
@@
 
-The main focus of this work is implementing this algorithm as efficiently as possible on our MCUs, and as such, several optimization efforts need to be taken into account during our experimentation.
-
 \subsubsection{3.5 Summary}\label{summary}
@@
 Considering the maximum conversion speed from the ADC and the USB-output from the MCU, we have a clear bound over 24 Mbit/s (\emph{imposed by the ADC}) in which Toeplitz extraction needs to be executed.
-Any speeds over 2.86 MB/s allows us to upgrade the ADC iteratively to continue increasing the output speed.
+Any speeds over 24 Mbit/s allows us to upgrade the ADC iteratively to continue increasing the output speed.
 Any implementation of Toeplitz extraction must then execute fast enough on any given microcontroller feasible for the proposed quantum RNG-thumbstick as to not be the decisive limiting factor.
@@
 To tackle this problem, Lin et al \cite{lin} proposed a method for seed-renewable Toeplitz post-processing in QRNG.
-Their strategy incorporates a dynamic seed pool within the FPGA, where each instance of post-processing picks a new, randomly selected seed.
-Thus, minimizing temporal correlations between extractions.
+Their strategy incorporates a dynamic seed pool within the FPGA, where each instance of post-processing picks a new, randomly selected seed, minimizing temporal correlations between extractions.
 Furthermore, an external seed updating mechanism via Peripheral Component Interconnect Express (\emph{PCIe}) ensures that seeds are refreshed whenever a certain security limit is reached.
@@
 
-Efficient Toeplitz matrix-vector multiplication (\emph{TMVM}) is critical for optimizing randomness extraction that relies on Toeplitz, particulary in constrained hardware environments.
+Efficient Toeplitz matrix-vector multiplication (\emph{TMVM}) is critical for optimizing randomness extraction that relies on Toeplitz, particularly in constrained hardware environments.
 Liao et al.
@@
 By using FFT and IFFT acceleration, Toeplitz post-processing for randomness extraction could achieve higher throughput.
-Thus, potentially could improve performance and reduce latency.
+This could potentially improve performance and reduce latency.
 
@@
 After establishing their individual effect, we combined both techniques, Multiple batch sizes were tested to determine their impact.
-More details and benchmarks for each configuration can be found in the section 6.
+More details and benchmarks for each configuration can be found in Section 6.
 
@@
 Phase two focused on addressing inefficiencies and design issues that were unintentionally introduced or not addressed during earlier iterations.
-Rather than continuing with new algorithmic ideas, this phase aimed to identify and fix structural problems.
+Rather than continuing with new algorithmic ideas, this phase aimed to identify and fix structural problems identified during the previous phase.
 Several assumptions from phase one were re-evaluated -- such as the benefits of certain data structures or abstractions.
@@
 The way \texttt{ent} calculates the entropy of a given binary file is by checking the frequency of certain byte values.
-If some byte values are more frequent, the binary file has a low entropy -- essentially singifying that there are predictable patterns.
+If some byte values are more frequent, the binary file has a low entropy -- essentially signifying that there are predictable patterns.
 Conversely, if all byte values appear roughly equally, the binary file has a high entropy -- meaning that it is more random, and less predictable.
@@
 This is calculated as \(1 - \frac{\text{entropy}}{\text{maximum entropy}}\), signifying that the file cannot be compressed further -- that the data is already as compressed as its random distribution allows.
-Essentially, this informs us that the file has no more detectable statistical patterns or compressable regularities remaining.
+Essentially, this informs us that the file has no more detectable statistical patterns or compressible regularities remaining.
 We consider any file where \texttt{ent} reports a compression ratio of \(0\%\) to be acceptable.
@@
 al.
-\cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (while subtracting the interrupt overhead) provides adequate measurements of the execution speed of an algorithm.
-An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (as the process for this may vary greatly between controllers).
+\cite{micromeasurements} propose that measuring execution time of algorithms directly via the microcontrollers internal timers (\emph{while subtracting the interrupt overhead}) provides adequate measurements of the execution speed of an algorithm.
+An additional advantage is that the same code can be used to measure execution speed on several different microcontrollers, rather than relying on counting CPU cycles (\emph{as the process for this may vary greatly between controllers}).
 
@@
 As such, we expect the time spent on executing other code to be negligible.
-Furthermore, this code performs significantly less work per cycle, and does not exponentially increase the larger the input and output requirements increase.
+Furthermore, this code executes far less often and does less work than the algorithm, and does not exponentially increase with larger the input and output requirements.
 
@@
 
-In section 6, this calculation will be used to derive the execution speed of the various iterations.
-
 \subsubsection{5.4 Limitations}\label{limitations}
@@
 
-We created a script to facilitate easier testing, which is attached to this paper.
+We created a script to facilitate easier testing, which is attached as an appendix to this paper.
 Using this script, all bitstrings used for testing can be evaluated against the baseline, ensuring that the output from the new iteration matches the baseline exactly.
@@
 Table \ref{tab:iter0} presents the average execution time of the naive implementation (\emph{iteration 0}) across different input sizes for both the Teensy 4.1 and the Raspberry Pi Pico 2.
-The results reveal that the Teensy consistently achieves lower execution times compared to the Pico, with performance differences becoming more pronounced as thje bit size increases.
-This behaivor aligns with the expected difference in hardware specifications, such as CPU clock frequency and memory bandwidth.
+The results reveal that the Teensy consistently achieves lower execution times compared to the Pico, with performance differences becoming more pronounced as the bit size increases.
+This behavior aligns with the expected difference in hardware specifications, such as CPU clock frequency and memory bandwidth.
 
@@
 
-Additionally, the isolated effect of applying a single bitmask operation \texttt{\&\ 1} was evaluated.
+Additionally, the isolated effect of applying a single bitmask operation \texttt{\&\ 1} was evaluated in the naive implementation.
 This resulted in a slight reduction in average execution time, from \texttt{13.1564} \(\mu s\) to \texttt{13.1000} \(\mu s\), although this specific result is not included in the tables above.
@@
 Table \ref{tab:iter3} presents the average execution time of iteration 3 on the Teensy 4.1 for varying batching and bit sizes.
-Additionally, when performing the same 1024/1024-bit operation, the Raspberry Pi Pico 2 showed average execution time of \texttt{80806.006} \(\mu s\) indicating a substantially lower throughput compared to the Teensy.
+The notation \texttt{64/64}, \texttt{512/512} and \texttt{1024/1024} refers to the input bit size and batching size.
+
+To evaluate the effect of batching granularity, additional tests were conducted using batch sizes of 32, 256 and 512 for various input sizes.
+The observed difference in execution time between low and high batching configurations ranged approximately from 10 to 50 microseconds, depending on the input size.
+
+For comparative analysis, the same 1024/1024-bit configuration was executed on the Raspberry Pi Pico 2.
+The average execution time was measured at \texttt{80806.006} \(\mu s\) indicating a substantially lower throughput compared to the Teensy.
 
@@
 Table \ref{tab:iter4} presents the execution time of iteration 4, which utilizes a 4x unrolled bit-processing loop.
-Compared to erlier iterations, this approach yields a substantial performance improvments on both the Teensy and the Raspberry Pi Pico Mcu's.
+Compared to earlier iterations, this approach yields a substantial performance improvements on both the Teensy and the Raspberry Pi Pico MCUs.
 
@@
 Specifically, the algorithm was incorrectly designed to transmit only a single output bit instead of full 32-bit expected output.
-This lead to an underestimation of the true execution time, as the cost of correctly formatting and sending the full result was omitted from measurement.
+This led to an underestimation of the true execution time, as the cost of correctly formatting and sending the full result was omitted from measurement.
 
@@
 
-Table \ref{tab:iter5} presents the results of iteration 5, in which vector structure were removed in favor of fixed-size integer types \texttt{uint32\_t} and \texttt{uint64\_t}.
+Table \ref{tab:iter5} presents the results of iteration 5, in which the original vector structure was removed in favor of utilizing fixed-size integer types \texttt{uint32\_t} and \texttt{uint64\_t}.
 This iteration led to an extremely efficient implementation that approaches the physical execution limits of the Teensy 4.1.
@@
 The ramifications of this will be discussed in more detail in Section 7.
-However, the results remain inconsistent---while the average is low, individual measurements vary, indicating that the implementation does not yet achieve stable execution at the architectural limit.
+However, the results remain inconsistent -- while the average execution speed is low, individual measurements vary, indicating that the implementation does not yet achieve stable execution at the architectural limit.
 
@@
 Specifically, the fixed-width output constrains imposed by the removal of vector.
-Several alternative data structures were evaluated, with the original vecotr-based implementation from iteration 1 serving as a baseline.
+Several alternative data structures were evaluated, with the original vector-based implementation from iteration 1 serving as a baseline.
 
@@
 This suggests that, despite nearing the theoretical limits outlined in Section 3.3, the implementation has not yet reached a consistently minimal execution time.
-Additional tests at 128-bit input size yielded 179.8712\,\(\mu s\), confirming that the system still operates within the physical capabilities of the Teensy 4.1.
+Additional tests at 128-bit input size yielded 179.8712\,\(\mu s\), confirming that this approach is not suitable for increasing the input and output bit sizes.
 
@@
 Observing the results, we can place them into the context of the limits imposed by the ADC, as discussed in Section 3.2.
-In equation (5) (\emph{as seen in Section 5.3}), we calculate the average execution speed required to be \(2.667 \mu s\) for the ADC that requires soldering to the microcontroller.
-From Tables \ref{tab:iter5} and \ref{tab:iter6} we demonstrate execution speeds well below that, with iteration 5 as well as iteration 6 -- specifically with bitsets -- being the most prominent results to discuss.
+In equation (5) (\emph{as seen in Section 5.3}), we calculate the average execution speed required to be \(2.667 \mu s\) for the ADC suggested in Section 3.2.
+From Tables \ref{tab:iter5} and \ref{tab:iter6} we can see execution speeds well below that, with iteration 5 as well as iteration 6 -- specifically with bitsets -- being the most prominent results to discuss.
 Here, our implementation achieved speeds of \(0.0501 \mu s\) and \(0.0474 \mu s\) respectively which within the bounds of the ADC listed in Section 3.2.
-The built in ADC on Teensy has an effective sample rate of 1 MSPS (\emph{e.g.~half of MAX11102AUB}), which would require an average execution speed of \textasciitilde{}\(2.667/2 \mu s \approx 1.334 \mu s\), making our implementation suitable even for that limited conversion speed.
+The built in ADC on Teensy has an effective sample rate of 1 MSPS (\emph{e.g.~half of MAX11102AUB}), which would require an average execution speed of \textasciitilde{}\(2.667 \mu s / 2 \approx 1.334 \mu s\), making our implementation suitable even for that limited conversion speed.
 
@@
 However, it is important to note that both of these implementations has a hard limit of working on 64 bits input.
-Whereas larger bitstrings still yield results with high entropy, the execution speed increases far above the hardware limitations, as evident in \ref{tab:iter6}.
-Iteration 5 is hard capped at using only 64 bits input, as we need to store the bits in fixed-size integers before running them through the algorithm, and on embedded machines we have no larger integer types available.
+Whereas larger bitstrings still yield results with high entropy, the execution speed increases far above the hardware limitations, as evident in Table \ref{tab:iter6}.
 
@@
 The larger the size of the matrix provided to the Toeplitz function (\emph{e.g.~the larger our input}), the higher the potential overall entropy becomes.
-As our implementation is essentially capped at 64 bits input at this point in time (\emph{without being limited by the hardware}), this may or may not prove to not be secure enough for use in real world applications.
+Iteration 5 is hard capped at using only 64 bits input, as we need to store the bits in fixed-size integers before running them through the algorithm, and on embedded machines we have no larger integer types available, which may or may not prove to not be secure enough for use in real world applications.
 This, however, is out of scope for this particular thesis, but a consideration nonetheless.
@@
 
-As mentioned in Section 7, this comes at a price.
+As discussed in Section 7, this comes at a price.
 The implementation that did indeed perform above expectation could only output a 32-bit integer, as we cannot take any larger input than 64 bits without any more complex data structure.
@@
 For less critical systems and less essential functionality, this method might, however, prove just secure enough.
-For instance, an IoT-device that requires a random number output for symmetric cryptography when connecting to a user device may be an ideal candidate for this method of OQRNG.
+For instance, an IoT-device that requires a random number output for short-term symmetric cryptography when connecting to a user device may be an ideal candidate for this method of OQRNG.
 

### 2025-06-04 docs/style/article-final.tex
the draft sent in for preliminary supervision

@@
 \usepackage{todonotes}
+
+\def\UrlBreaks{\do\/\do-}
 
@@
 
-\section{ABSTRACT}\label{abstract}
-
-True random number generation is indispensable for a wide range of cryptographic and distributed computing applications, yet traditional pseudo-random number generators remain deterministic, and thus vulnerable to state‐recovery attacks.
-Quantum random number generation (\emph{QRNG}) offers a robust alternative by harnessing inherent quantum fluctuations -- in particular, optical QRNG (\emph{OQRNG}) based on shot noise provides an accessible source of entropy from which true random numbers can be generated.
-However, existing OQRNG solutions typically rely on expensive or bespoke hardware and defer post‐processing to host computers, limiting portability and integration.
-In this work, we investigate the feasibility of performing Toeplitz randomness extraction directly on resource constrained microcontrollers, thus creating a fully self‐contained QRNG system.
-We address two primary research areas: RA1 developing an implementation of Toeplitz extraction that minimizes processing time on embedded hardware, and RA2 ensuring that the extraction algorithm does not become the throughput bottleneck compared to ADC conversion or USB transfer speeds.
-
-Our results demonstrate that, by eliminating dynamic allocations and leveraging bit parallel techniques, Toeplitz extraction on 64-bit inputs can be executed in approximately \(0.05 \mu s\) on the Teensy 4.1 -- well below the \(2.667 \mu s\) threshold imposed by a 12-bit, 2 MSPS ADC.
-Even accounting for variance in individual measurements, these speeds confirm that the algorithmic implementation does not constrain overall throughput.
-On the Pico 2, the same implementations yield sub‐microsecond performance, albeit with a wider variance.
-However, our most efficient implementations are limited to 64‐bit inputs -- extending to larger input bit sizes incurs significant overhead due to the absence of wider integer types in typical MCU toolchains.
-
-In conclusion, Toeplitz extraction can be performed in real time on low‐power microcontrollers without becoming the system's performance bottleneck, validating the viability of a portable QRNG thumbstick.
-Future work should explore custom data types or lightweight libraries for 128- and 256-bit size extraction, as well as integration with a live OQRNG device to assess end‐to‐end randomness quality and throughput.
-
 \section{1 INTRODUCTION}\label{introduction}

### 2025-06-05 pres/opp/stack.rb
clean up

@@ -1,351 +1,0 @@
-#!/usr/bin/env ruby
-
-# This file is called rdparse.rb because it implements a Recursive
-# Descent Parser. Read more about the theory on e.g.
-# http://en.wikipedia.org/wiki/Recursive_descent_parser
-
-# 2010-02-11 New version of this file for the 2010 instance of TDP007
-#   which handles false return values during parsing, and has an easy way
-#   of turning on and off debug messages.
-# 2014-02-16 New version that handles { false } blocks and :empty tokens.
-
-require 'logger'
-
-class Rule
-
-  # A rule is created through the rule method of the Parser class, like this:
-  #   rule :term do
-  #     match(:term, '*', :dice) {|a, _, b| a * b }
-  #     match(:term, '/', :dice) {|a, _, b| a / b }
-  #     match(:dice)
-  #   end
-  
-  Match = Struct.new :pattern, :block
-  
-  def initialize(name, parser)
-    @logger = parser.logger
-    # The name of the expressions this rule matches
-    @name = name
-    # We need the parser to recursively parse sub-expressions occurring 
-    # within the pattern of the match objects associated with this rule
-    @parser = parser
-    @matches = []
-    # Left-recursive matches
-    @lrmatches = []
-  end
-  
-  # Add a matching expression to this rule, as in this example:
-  #   match(:term, '*', :dice) {|a, _, b| a * b }
-  # The arguments to 'match' describe the constituents of this expression.
-  def match(*pattern, &block)
-    match = Match.new(pattern, block)
-    # If the pattern is left-recursive, then add it to the left-recursive set
-    if pattern[0] == @name
-      pattern.shift
-      @lrmatches << match
-    else
-      @matches << match
-    end
-  end
-  
-  def parse
-    # Try non-left-recursive matches first, to avoid infinite recursion
-    match_result = try_matches(@matches)
-    return nil if match_result.nil?
-    loop do
-      result = try_matches(@lrmatches, match_result)
-      return match_result if result.nil?
-      match_result = result
-    end
-  end
-
-  private
-  
-  # Try out all matching patterns of this rule
-  def try_matches(matches, pre_result = nil)
-    match_result = nil
-    # Begin at the current position in the input string of the parser
-    start = @parser.pos
-    matches.each do |match|
-      # pre_result is a previously available result from evaluating expressions
-      result = pre_result.nil? ? [] : [pre_result]
-
[…]

### 2025-06-05 pres/opp/stacksolution.rb

@@ -1,354 +1,0 @@
-#!/usr/bin/env ruby
-
-# This file is called rdparse.rb because it implements a Recursive
-# Descent Parser. Read more about the theory on e.g.
-# http://en.wikipedia.org/wiki/Recursive_descent_parser
-
-# 2010-02-11 New version of this file for the 2010 instance of TDP007
-#   which handles false return values during parsing, and has an easy way
-#   of turning on and off debug messages.
-# 2014-02-16 New version that handles { false } blocks and :empty tokens.
-
-require 'logger'
-
-class Rule
-
-  # A rule is created through the rule method of the Parser class, like this:
-  #   rule :term do
-  #     match(:term, '*', :dice) {|a, _, b| a * b }
-  #     match(:term, '/', :dice) {|a, _, b| a / b }
-  #     match(:dice)
-  #   end
-  
-  Match = Struct.new :pattern, :block
-  
-  def initialize(name, parser)
-    @logger = parser.logger
-    # The name of the expressions this rule matches
-    @name = name
-    # We need the parser to recursively parse sub-expressions occurring 
-    # within the pattern of the match objects associated with this rule
-    @parser = parser
-    @matches = []
-    # Left-recursive matches
-    @lrmatches = []
-  end
-  
-  # Add a matching expression to this rule, as in this example:
-  #   match(:term, '*', :dice) {|a, _, b| a * b }
-  # The arguments to 'match' describe the constituents of this expression.
-  def match(*pattern, &block)
-    match = Match.new(pattern, block)
-    # If the pattern is left-recursive, then add it to the left-recursive set
-    if pattern[0] == @name
-      pattern.shift
-      @lrmatches << match
-    else
-      @matches << match
-    end
-  end
-  
-  def parse
-    # Try non-left-recursive matches first, to avoid infinite recursion
-    match_result = try_matches(@matches)
-    return nil if match_result.nil?
-    loop do
-      result = try_matches(@lrmatches, match_result)
-      return match_result if result.nil?
-      match_result = result
-    end
-  end
-
-  private
-  
-  # Try out all matching patterns of this rule
-  def try_matches(matches, pre_result = nil)
-    match_result = nil
-    # Begin at the current position in the input string of the parser
-    start = @parser.pos
-    matches.each do |match|
-      # pre_result is a previously available result from evaluating expressions
[…]

### 2025-06-05 project/deprecated/arduino/connect.sh

@@ -1,32 +1,0 @@
-#!/bin/bash
-
-echo "Detecting platform..."
-
-UNAME_OUT="$(uname -s)"
-case "${UNAME_OUT}" in
-  Linux*)   PORT_GLOB="/dev/ttyACM* /dev/ttyUSB*";;
-  Darwin*)  PORT_GLOB="/dev/tty.usbmodem* /dev/tty.usbserial* /dev/cu.usbmodem*";;
-  *)        echo "Unsupported OS: ${UNAME_OUT}"; exit 1;;
-esac
-
-echo "Searching for available serial ports..."
-PORTS=$(ls $PORT_GLOB 2>/dev/null)
-
-if [ -z "$PORTS" ]; then
-  echo "No serial devices found. Please connect a microcontroller and try again."
-  exit 1
-fi
-
-echo "Available ports:"
-echo "$PORTS"
-
-echo "Select a port to connect to:"
-select PORT in $PORTS; do
-  if [ -n "$PORT" ]; then
-    echo "Connecting to $PORT..."
-    minicom -D "$PORT" -b 115200
-    break
-  else
-    echo "Invalid selection, please try again."
-  fi
-done

### 2025-06-05 project/deprecated/firmware_singlecore/src/main.rs

@@ -1,61 +1,0 @@
-#![no_std]
-#![no_main]
-
-use bsp::board;
-use bsp::hal::usbd::{BusAdapter, EndpointMemory, EndpointState};
-use teensy4_bsp as bsp;
-use teensy4_panic as _;
-use usb_device::bus::UsbBusAllocator;
-use usb_device::prelude::*;
-use usbd_serial::{SerialPort, USB_CLASS_CDC};
-
-#[bsp::rt::entry]
-fn main() -> ! {
-    let board::Resources {
-        usb,
-        mut gpio2,
-        pins,
-        ..
-    } = bsp::board::t41(bsp::board::instances());
-
-    usb.enable();
-
-    // ✅ Set up LED for debugging
-    let mut led = bsp::board::led(&mut gpio2, pins.p13);
-
-    // ✅ Allocate USB memory & state
-    static mut EP_MEMORY: EndpointMemory<1024> = EndpointMemory::new();
-    static mut EP_STATE: EndpointState = EndpointState::max_endpoints();
-
-    // ✅ Set up the USB Bus Adapter
-    let usb_adapter = unsafe { BusAdapter::new(usb, &mut EP_MEMORY, &mut EP_STATE) };
-
-    // ✅ Allocate USB Bus
-    let usb_alloc = UsbBusAllocator::new(usb_adapter);
-
-    // ✅ Create USB Serial device
-    let mut serial = SerialPort::new(&usb_alloc);
-
-    // ✅ Set up the USB Device
-    let mut usb_dev = UsbDeviceBuilder::new(&usb_alloc, UsbVidPid(0x16c0, 0x0483))
-        .product("Teensy USB Serial")
-        .device_class(USB_CLASS_CDC)
-        .build();
-
-    led.toggle();
-
-    loop {
-        // ✅ Poll USB to keep it alive
-        if usb_dev.poll(&mut [&mut serial]) {
-            let mut buf = [0u8; 64];
-
-            // ✅ Read user input from `/dev/ttyACM1`
-            if let Ok(count) = serial.read(&mut buf) {
-                if count > 0 {
-                    led.toggle(); // ✅ Blink LED on input
-                    serial.write(&buf[..count]).ok(); // ✅ Echo input back
-                }
-            }
-        }
-    }
-}

### 2025-06-05 project/deprecated/utils/pipe.py

@@ -1,53 +1,0 @@
-# import sys
-#
-# if len(sys.argv) != 3:
-#   print("Usage: python3 pipe.py input.bin /dev/tty.usbmodemXXXX")
-#   sys.exit(1)
-#
-# input_file = sys.argv[1]
-# serial_port = sys.argv[2]
-#
-# with open(input_file, "rb") as f, open(serial_port, "wb", buffering=0) as out:
-#   for byte in f.read():
-#     bits = format(byte, "08b")
-#     out.write(bits.encode())
-
-import sys
-import serial
-import time
-
-if len(sys.argv) != 4:
-  print("Usage: python3 pipe.py input.bin /dev/tty.usbmodemXXXX output.bin")
-  sys.exit(1)
-
-input_path = sys.argv[1]
-serial_path = sys.argv[2]
-output_path = sys.argv[3]
-
-with serial.Serial(serial_path, 115200, timeout=1) as ser:
-  time.sleep(2)
-  ser.reset_input_buffer()
-
-  with open(input_path, "rb") as f:
-    for byte in f.read():
-      bits = format(byte, "08b")
-      ser.write(bits.encode())
-      time.sleep(0.0005)
-
-  extracted_bits = ""
-  while True:
-    if ser.in_waiting:
-      line = ser.readline().decode(errors='ignore').strip()
-      if line.startswith("out:"):
-        extracted_bits += line[4:]
-    else:
-      break
-
-byte_chunks = [extracted_bits[i:i+8] for i in range(0, len(extracted_bits), 8)]
-with open(output_path, "wb") as f:
-  for chunk in byte_chunks:
-    if len(chunk) < 8:
-      chunk = chunk.ljust(8, '0')
-    f.write(int(chunk, 2).to_bytes(1, byteorder='big'))
-
-print(f"Wrote {len(byte_chunks)} bytes of extracted output to {output_path}")
